### 2016-04-09 paper.tex
Initial

\documentclass[titlepage]{article}

\usepackage{setspace}

\usepackage[utf8]{inputenc}
\usepackage{amsmath}   
\usepackage{mathtools}
\usepackage{graphicx}
\usepackage{subcaption}
\usepackage{float}
\usepackage{hyperref}
\usepackage{listings}
\usepackage[labelfont=bf]{caption}
\usepackage[margin=3.5cm]{geometry}

%\usepackage{qtree}
%\usepackage{tikz}
%\usepackage{tikz-qtree}

%\usepackage[]{algorithm2e}

\lstset{
	basicstyle=\footnotesize\ttfamily,
}

\title{
\vspace{80pt}\\
\LARGE \bfseries SKEDGE
\\\vspace{10pt}\Large Smarter course scheduling for our\\University of Rochester
}

\author{
	Dan Hassin\\
    \vspace{5pt}\\
    Supervised by Professor Philip Guo\\
    \vspace{5pt}\\
    Department of Computer Science\\
    University of Rochester\\
    Rochester, New York\\
}

\date{April 12, 2016\\
    \vspace{150pt}\\
    submitted in partial fulfillment of\\
    the requirements for the degree\\
    honors bachelor of science\\
}

\begin{document}

\maketitle

\doublespacing

% Direct search for specific course
% Discovery, browsing, exploring
% behavioral patterns

% STAT: Find major of users, find how many non-major courses they found

\begin{abstract}

In this paper I present Skedge, a web application for students to comfortably and effectively engage with the University's course catalog. Skedge matches and surpasses the capabilities of the existing University tool for this purpose, ``Course Description / Course Schedule'' (CDCS) and presents its information in a more visually pleasing way. As a result, Skedge boasts strong user-retention rates, long session durations, and high student adoption despite having virtually no advertisement. Through collected usage data, I demonstrate that a) Skedge's differences from and additions to CDCS are usable and have real need, b) the two major use-cases associated with course browsing---direct search and exploratory search---are effectively accommodated by Skedge, and c) Skedge's search mechanism is user-friendly and self-teaches to users over time.

\end{abstract}

\begin{figure}[ht]
    \centering
    \begin{tabular}{c c}
        \begin{subfigure}[h]{6.5cm}
            \centering
            \includegraphics[width=1.00\textwidth]{cdcs}
        \end{subfigure}
        \hspace{1em}
        \begin{subfigure}[h]{7.5cm}
            \centering
            \includegraphics[width=1.00\textwidth]{skedge}
        \end{subfigure}
    \end{tabular}
    \caption{CDCS (left) and Skedge (right) for the search query {\tt csc}.}
\end{figure}

\section{Introduction}

This paper will begin by discussing Markov chains and how we set out to use them for our goal of creating an assistive program editor. We will then describe in detail how we implemented a Markov chain to house program syntax trees, how we were able to generate code using it, and finally we'll discuss our results and address the limitations of our product.

\subsection{Markov chain}

A Markov chain is a mathematical system that undergoes a stochastic process whereby
the current state and a set of probabilities are used to determine what state
to transition to next. In the most simple case of a Markov chain, the next action is based
only upon the current state and all other previous actions are forgotten.
Thus, this process is essentially ``memoryless'' because all past states can be
forgotten once a new state is reached \cite{markov}. In this way, Markov chains can be likened
to deterministic finite automata. So while they cannot remember a limitless amount of their history,
calling them ``memoryless'' may be misleading, as they can indeed be implemented to remember a fixed
size of their state history.

Markov chains find application in many different fields. Already familiar with Markov chains used for generating random dialogue (by training a model based on bigrams and trigrams of common speech), we were interested in more novel domains. We chose to use a Markov chain to drive code generation, because, like text, we feel that code can be personal and expressive, and the idea of generating code possibilities from a training set seemed like exciting research. It was also challenging. Since randomly generating code tokens would most likely produce syntactically invalid programs, we looked at the abstract syntax trees of the programs and had to fit them into a Markov chain model. And as our research developed, we found ourselves with a new vision of an exciting project.

\subsection{Emerging goal}

For many novice programmers, programming can seem like a confusing and
tedious activity. It can be unclear for many new programmers how they should
begin and what code should come next. They may conceptually understand
what the next step should be, and yet be unable to express it with code. Or, they
could be unsure of what to do next, and perhaps giving code suggestions could inspire them
and help them to understand it conceptually.

Even as one becomes more
adept at programming, he or she may notice typing the same or
similar lines of code over and over again. This repetitiveness slows down
the user and can make programming more tiresome a task.

So, the prospect of using a Markov chain to generate code snippets in an assistive, predictive integrated development environment (IDE) for JavaScript seemed to be a natural progression from our research. Our final product, a Probabilistic Programming
web application, allows a user to code in JavaScript by way of selecting from
a list of suggestions for the user's next line of code. We imagine that this won't only be helpful for new programmers,
but that this system would allow for expert programmers to write code with more ease and efficiency.

% overview of markov chain

\section{Tree-based Markov chain}

\subsection{Abstract syntax trees}

An abstract syntax tree (AST) is a tree-based, rather than text-based, representation of a program.
Such a tree is generated using the grammar rules of a programming language and text input. In the tree, \emph{nodes}
are language control structures such as {\tt for} loops, {\tt if} statements, variable and function
declarations, and so on. The \emph{edges} connecting a parent node to its children represent the different \emph{features}
of that node. For instance, a {\tt for} loop would have the features \emph{init} (e.g. {\tt var i = 0;}),
\emph{test} (e.g. {\tt i < array.length;}), \emph{update} (e.g. {\tt i++}), and a \emph{body} (a list
of statements that should be executed per loop iteration). To implement lists, we include the additional
\emph{follow} edge as a feature to all statements, which would connect a statement node to the next statement
in the list. If the statement is last in its container list, we connect its \emph{follow} edge to a special
{\tt {\_end}} node. (See figure~\ref{fig:sample-ast} for an example of a AST.)

% \begin{center}

% \begin{tikzpicture}[level distance=1.5cm,
% level 1/.style={sibling distance=3.5cm},
% level 2/.style={sibling distance=1cm}]

% \node (Root) {Program}
% child {
%     node {for} edge from parent node[draw=none] {\emph{body}}
%     child { node {VD} edge from parent node[draw=none] {\emph{init}} }
% };

% \end{tikzpicture}

% \end{center}

Our system deals with programs at the level of AST's as they provide more structure and modularity than text.
We use the syntax tree format of Mozilla's Parser API \cite{parser api}, and \emph{esprima} \cite{esprima}, an ECMAScript
(the language standard of which JavaScript is an implementation)
parser, to transform program source code into an AST of this format. We chose JavaScript because it is (a) dynamic and interpreted, (b) runnable in a web browser, and (c) because JavaScript AST tools already existed.

\subsection{Our Markov chain}

To build our Markov chain structure, we parse a corpus of JavaScript files into their
respective AST's, and then collect statistics from these trees to establish probability distributions of nodes for a given AST context.

An \emph{AST path} is a list of items, alternating between node types and features, that encodes the unique position of a 
specific node in the AST. The \emph{probability distribution} for an AST path is the set of possible node types or values that would appear in the
syntax tree following the given path, and the respective probabilities of their
occurrence at that point. The model maps AST paths to the probability distributions of node types (or values, as we'll see later) for that given
path.

More formally, given a depth $d$ (indicating how many nodes we should backtrace) for a given
${node}_d$ and feature ${edge}_d$, the generation of ${node}_d$'s child node given by
 ${edge}_d$ can be expressed by the following function $f$: \begin{align*}
f({node}_1, {edge}_1, ..., {node}_{d}, {edge}_{d}) = \begin{dcases*}
		{newnode_1},	&	$P({newnode_1})$ \\
		{newnode_2},	&	$P({newnode_2})$ \\
		...\\
		{newnode_n},    &   $P({newnode_n})$ \\
	\end{dcases*},
\end{align*}

where \begin{align*}
\sum\limits_{i = 0}^n P({newnode_i}) = 1,
\end{align*}

\noindent and ${Path}({node}_1, {edge}_1) = {node}_2$, ${Path}({node}_2, {edge}_2) = {node}_3$, ..., ${Path}({node}_d, {edge}_d) = {newnode}_k$, where ${newnode}_k$ is randomly selected from ${newnode}_1, ..., {newnode}_n$ given the probabilities $P({newnode_1}), ..., P({newnode_n})$. From this it can be seen that generating programs is simply a repetitive application of this function, where for the chosen ${newnode}_k$, we can generate each child of ${newnode}_k$ with the following formula: \begin{align*} &\forall {edge} \in {Features}({newnode}_k), \\
& \hspace{3em} {Path}({newnode}_k, {edge}) = f({node}_2, {edge}_2, ..., {node}_{d}, {edge}_{d}, {newnode}_k, {edge}).
\end{align*}

This process must eventually terminate as the probability of generating {\tt \_end} nodes for a \emph{follow} edge will always be non-zero (since the programs we parse into our corpus are surely finite) and such nodes have no features to be ``expanded.''

\subsection{Implementation}

\subsubsection{Data structure}

Building the Markov chain model means being able to provide the probability distribution of new nodes for arbitrary input,
as described by the function in Section 2.2. To do this, we construct a data structure using JavaScript Object Notation (JSON),
a structural format useful for ease of specification and integration with JavaScript programs.

The data structure is a hash table (also commonly called a map, dictionary, or associative array), that, on a given input, returns another hash table for paths that begin with that input. This behavior is repeated until the number of inputs reaches the desired depth, at which point a final hash table, mapping the node type to its probability distribution, is returned. This, in effect, is modularizing the function $f$ described above, so that we can reach the probability distribution with minimal code and data duplication. This is demonstrated by following example, given  ${depth} = 2$. \begin{align*}
lookup(lookup(lookup(lookup({Model}, {node}_1), {edge}_1), {node}_2), {edge}_2) = \textbf{P}({newnodes}).
\end{align*}

\subsubsection{Building the model}

We first obtain collection of JavaScript programs in AST form. We then run the following algorithm to construct our model:

\begin{verbatim}
procedure addToModel(node, path):
    model = CorpusModel
    for each elem in path[path.length-depth*2 .. path.length]:
        model = model[elem]
    model[node.type] += 1
    model[_total] += 1
    for each feature of node:
        temp_path = path + [node.type, feature]
        if node.feature is list of nodes:
            for each item in node.feature:
                addToModel(item, temp_path)
                temp_path += [item.type, "FOLLOW"]
            addToModel(END_NODE, temp_path)
        else:
            addToModel(node.feature, temp_path)

for each parsed AST:
    addToModel(AST.root, [])
\end{verbatim}

For each AST, we walk through every node in the tree by recursively exploring each feature\footnote{While we use \emph{follow} as if it were a feature edge, programs encoded in the AST's we parse use arrays of nodes (rather than linked-lists as we represent them) for a collection of statements. The given pseudocode demonstrates how we can transform an array into a linear sequence of edges.} of each node. In each call to explore a node's feature, we maintain the path up until that node by appending to it the node type and the name of the feature to reach the wanted child. We trim the beginning of the path off so that the length of the path is appropriate with respect to the chosen depth (i.e. so that $|path| = 2*d$).\footnote{Note that this requires that the path is at least of length $2*depth$. If the path is shorter than this, it represents not having at least $d$ levels of context. This can occur, say, for the first statement in the program, where there are no preceding nodes in its AST path. For these cases, we add the ``{\tt \_null}'' feature until we reach the desired depth (omitted in the pseudocode.)} We traverse through the layers of hash tables, creating the path elements if they don't already exist, add the node type to the probability distribution if it isn't already included, and then increment a counter associated with that node type. We also increment a ``{\tt \_total}'' counter for the convenience of the generator program.

When this process is complete, the JSON object is saved to a file and can be augmented by more code samples thereafter. (An example of such a model is shown in figure~\ref{fig:sample-model} in the Appendix. In this model we see, for instance, that the path
[\emph{Program, body, VariableDeclaration, declarations}] leads to
\emph{VariableDeclarator} with probability 1. This indicates that a \emph{Program}
node whose \emph{body} is a \emph{VariableDeclaration} always has a statement that is of
type \emph{VariableDeclarator} as its first statement.)

\subsection{Generating code}

Code can be generated rather easily using the approach explained in Section 2.2. The program tree is built up one node at a time, each one generated based on its unique AST path. The following algorithm demonstrates this process.

\clearpage

\begin{verbatim}
procedure generateNode(model, path):
    for each elem in path[path.length-depth*2 .. path.length]:
        model = model[elem]
    nodeType = pickRandomNodeType(model)
    node = {`type':nodeType}
    for each feature of nodeType:
        new_path = path + [node.type, feature]
        if feature is a list feature:
            nodes = []
            while (newNode = generateNode(model, new_path) != END_NODE:
                nodes += [nodeNode]
                new_path += [newNode.type, "FOLLOW"]
            node.feature = nodes
        else:
            node.feature = generateNode(model, new_path)
    return node

generateNode(CorpusModel, [])
\end{verbatim}

The algorithm looks very similar to the parse algorithm, as it is almost the same process, simply inverted. Rather than examine nodes, determine a path by building it up, and then increment a count in the model, this algorithm first finds the counts in the model for a \emph{given} path, generates a node, and generates nodes for all of that node's features recursively. (Since the final hash table is not a proper probability distribution, and is rather a tally of each node type, here the {\tt \_total} is useful for determining ratios.) The result is a generated AST for our program, which is then fed into \emph{escodegen} \cite{escodegen}, an ECMAScript code generator to output program source code that we can display and run.

To generate random programs, we always begin with a ``Program'' node, and then add nodes to its \emph{body} feature (an array.) However, we quickly found that while entertaining, generating random programs had little potential for real application. Instead, we turned to generating code snippets (to be placed in specific contexts of the syntax tree.) Since our algorithm was general enough for this purpose, no more had to be done besides knowing the path for the node we wanted to generate.

\subsubsection{Identifiers and literals}

Naturally, identifiers and literals are a very large part of programming. Without them, code would be entirely unspecific, and would only consist of syntactic symbols and keywords. Yet they pose an issue with code generation, as we found generating identifiers or literals that are likely to be what the user is expecting an especially difficult problem to solve, let alone model. Struggling to come up with a simple, intelligent, and robust solution, and short on time, we fell back to using our existing Markov approach.

In the current implementation of this system, following the paths [\emph{Identifier}, \emph{Name}], or [\emph{Literal}, \emph{Raw}] result not in node types, but rather in the text of the identifier names or literal values. This way, generating identifiers and literals is simply a probabilistic process like the rest of the code generation. Of course, this means that sometimes identifiers are added that are not in scope, that have mismatched types, or that simply make no semantic sense. (But adjusting $depth$ did seem to make programs more meaningful, as expected. This can be seen experimentally by modifying depth on our web IDE, which is discussed in Section 4.1.)

%\lstinputlisting[caption=Scheduler, style=json]{../rand.js}

\section{Applications}

\subsection{\emph{vim} autocomplete script}

We integrated our model into the text editor vim to offer autocompletion
suggestions of programs upon the user's request, using vim's
\texttt{completefunc} feature. When the user is editing a JavaScript file in vim with
this feature, they may press a keyboard command to get a list of program
snippets to choose from. The snippets are generated in the context of the
program path at cursor's position. The user may pick one of the snippet
suggestions to have it inserted into their code at their cursor. The autocomplete snippets are generated as follows:
\begin{enumerate}
	\item Parse the contents of the current file into an AST.
	\item Find the node path corresponding to the cursor by traversing the AST.
	\item Generate programs starting from the node path of the cursor.
	\item Present the programs to the user as code completion snippets.
\end{enumerate}

\begin{figure}[ht]
	\centering
	\begin{tabular}{c c}
		\begin{subfigure}[h]{6.25cm}
			\centering
			\includegraphics[width=1.00\textwidth]{vim1}
			\caption{Inserting a {\tt for} loop} \label{fig:vim1}
		\end{subfigure}
		\begin{subfigure}[h]{6cm}
			\centering
			\includegraphics[width=1.00\textwidth]{vim2}
			\caption{Inserting an object property} \label{fig:vim2}
		\end{subfigure}
	\end{tabular}
	\caption{Screenshots of vim using autocomplete of programs}
\end{figure}

The vim autocomplete feature is limited to inserting program snippets on a single
line, and requires the user to explicitly trigger the \texttt{completefunc}
feature to be prompted with code snippets. To further explore using our model
for editing, we created a web-based editor that more deeply integrates Probabilistic
Programming.

\subsection{Guided IDE}

The Guided IDE offers users an environment for building a program entirely using
our model, using program snippets generated at the cursor's context as building
blocks. The editor presents a window for the program text, and a list of
suggestion code snippets. The user can move a cursor through the code window at
positions corresponding to boundaries of program nodes. As the user moves the
cursor, the snippet suggestions are regenerated using the cursor's node path, to
continuously provide suggestions relevant to the current context of the program. When the
user selects a suggestion snippet, its AST is inserted into the program's AST at
the cursor's position in the tree, and the code window is redrawn. The user can
move the cursor within newly inserted blocks, so that the program can grow
deeper. The initial corpus for the IDE's model consists of the JavaScript files
that comprise our project.

\begin{figure}[h]
  \centering
  \includegraphics[width=1.00\textwidth]{ide}
  \caption{Screenshot of the web program editor GUI. The header along the top indicates the path (trimmed to the selected depth) to the current position of the cursor. Code snippets on the left can be navigated using the arrow keys.} \label{fig:screenshot}
\end{figure}

\section{Results}

\subsection{Using the IDE}

We can use the Guided IDE to generate programs. The programs look like actual
programs but do not actually function unless the user is very careful about
which snippets they pick. Crucially, the variable and identifier names that our
system generates are not in general meaningful in the scope of the generated
program.  Whereas with the vim autocompletions the identifiers would often be
valid because because the corpus was the current file, with the larger corpus of
the Guided IDE and writing a program from scratch, the user is less likely to
find program snippets that use valid identifiers.

\subsection{Limitations of the model}

It is a limitation of the current model that identifier names are chosen
probabilistically based on their position in the code, without regard to whether
they are in scope and without refering to anything in another part of the
program. Because of this limitation, the Guided IDE is not useful for practical
purposes. We consider some alternative approaches for the Guided IDE using a
similar model.

\subsection{Alternative approaches for the IDE}

\subsubsection{Prompt for identifiers}

Rather than picking identifier names from the model, use placeholder identifiers
and prompt the user to enter actual identifiers, with suggestions based on the
current variables in scope and previously used identifiers. This would give the
user the opportunity to create semantically valid programs, but the entering of
identifier names could be tedious.

\subsubsection{Stubs}

Instead of having a movable cursor in the code window, consider stubs in the
program. Each stub would be a marker in place of an AST node. The user would
select a stub and be given choices of snippets to replace it. Selecting a
snippet would move the current snippet into the program in place of the selected
stub.  The generated snippets could themselves have stubs in them, so that the
user may continue to complete stubs and expand their program.  At the level of
variable identifiers, the snippet suggestions would be taken from the computed
list of in-scope variables at the position of the selected stub, unless the
completion was for a variable declaration, in which case the user would be
prompted to name a new variable.

This approach, while offering less places to insert code then the current
approach, would allow for programs to more closely be generated from our model,
since the stub completions would always be at leaf nodes of the AST, whereas in
the current approach the user may choose to insert a snippet in the middle of a
list, which means that the transition from the inserted snippet to the following
node in the list does not in general follow from the model.

While the stub approach would be more restrictive for the user as it requires
them to build the program in a certain way, with proper handling of identifiers,
it may provide for an interesting or useful paradigm for interacting with a program
source.

\subsection{Intelligent handling of identifiers and literals}

We explored other approaches to solve the issue of generating meaningful identifiers and literals described before.

\subsubsection{Secondary models}

When generating identifiers, we were interested in generating meaningful ones rather than random ones (albeit from a weighted distribution.) We wanted to design a model that could keep track of when the same variables were used in the same (or similar) place.

One idea was to, during the parse, keep track of the path at which each variable identifier was used, grouped by the identifier name. When an identifier needs to be generated, we can find that path in our secondary model, and if so, bind to a random (but weighted) identifier that has been used at that path, with the weight being based on the frequency of that identifier appearing at that path. For subsequent statements that need identifiers, the same process could occur, and, if the same variable group was chosen, it could generate the variable name that was bound to the first statement.

Another idea also involved constructing a secondary model during the parse, that kept track of statements that used the same identifier. When such statements are found, it would compute the path between those statements and increment a counter associated that path. This would represent the likelihood of two nodes connected by that path using the same identifier. Then, during code generation, whenever an identifier is needed, the generator would make a list of all variables in scope at that point, and compute the path from the variable declaration (and/or last use) to the node being generated. It would look up each path in the secondary model and assign each of the in-scope variables a weight. A variable could then be chosen based on this probability.

% We suspect that this may affect the complexity of our algorithms significantly. If what the system is doing now is analogous to the behavior of a context-free grammar (i.e. by recursively generating nodes), then detection of use of the same identifiers across nodes would involve pattern-matching and perhaps even analogy-solving. This could analogously be considered more like the action of a Turing machine, which would mean that the complexity of the process could be unbounded.

Unfortunately, we were unable to implement these ideas or explore them further due to time constraints.

\subsubsection{Using a higher-order model}

Just as our approach of modeling a program with a tree-based Markov chain is a
higher order model than a linear Markov chain of the program source code tokens,
there may be applications of probabilistic modeling to higher-order
representations of programs than the AST, such as a \emph{higher-order abstract syntax}, which promises to ``incorporate name binding information in a uniform and language generic way'' \cite{hosa}. Such an application may model the relationship between different
parts of an AST, and handle scoping of variables to model identifiers
abstractly.

With a more sophisticated model of identifiers that could generate programs with
identifiers in meaningful relation to each other, Probabilistic Programming could
be used to generate semantically valid programs, for greater assistive,
expressive, and programmatic power.

\section{Conclusion}

We consider the Probabilistic Programming web IDE a strong proof of concept for context-based code snippet generation. We are aware of some of its serious structural limitations, but we speculate that related concepts, such as genetic algorithms, might enjoy more success. Specifically, we are interested in the possibility of creating genetic algorithms that take advantage of this platform to build and ``evolve'' code, swapping parts of AST's in a Markovian-like process. Perhaps the evolution could even be given a goal; for instance, to satisfy a set of given test cases. Regardless, we believe that our research opens new possibilities with program generation.

%Other applications of the model: genetic algorithms

\subsection*{Source code \& demo}

The source code for Probabilistic Programming is available online under an open
source license:
\url{https://github.com/RocHack/jschain}.\\

\noindent The web IDE can be found at
\url{https://csug.rochester.edu/u/clehner/jschain/web/}.

\clearpage

\begin{thebibliography}{1}

	\bibitem{esprima} Ariya Hidayat {\em Esprima}
		\\\url{http://esprima.org/}

	\bibitem{markov} Takis Konstantopoulos {\em Introductory lecture notes on
		Markov Chains and Random Walks.} Uppsala University,
		\\\url{http://www2.math.uu.se/~takis/L/McRw/mcrw.pdf}

	\bibitem{hosa} Frank Pfenning, Conal Elliott
		{\em Higher-Order Abstract Syntax.}
		ACM SIGPLAN Notices. Vol. 23. No. 7. ACM, 1988.
		\\\url{http://www.cs.cmu.edu/~fp/papers/pldi88.pdf}

	\bibitem{escodegen} Yusuke Suzuki {\em Escodegen}
		\\\url{https://github.com/Constellation/escodegen}

	\bibitem{parser api} Mozilla Developer Network {\em SpiderMonkey Parser API}
		\\\url{https://developer.mozilla.org/en-US/docs/SpiderMonkey/Parser_API}

\end{thebibliography}

\clearpage
\section*{Appendix}

\begin{figure}[h!]
	\caption{JSON AST sample}
	\label{fig:sample-ast}
	\centering
    \vspace{-8pt}
	An AST generated from the program ``{\tt var x = 34;}"

	\lstinputlisting{ast-sample.json}
\end{figure}

\begin{figure}[h!]
	\caption{JSON Markov model sample}
	\label{fig:sample-model}
	\centering
    \vspace{-8pt}
	A Markov chain model generated from the program ``{\tt var x = 34;}"

	\lstinputlisting{model-sample.json}
\end{figure}

\end{document}

### 2016-04-09 paper.tex
Remove old stuff

@@
             \centering
-            \includegraphics[width=1.00\textwidth]{cdcs}
+            \includegraphics[width=1.00\textwidth]{images/cdcs}
         \end{subfigure}
@@
             \centering
-            \includegraphics[width=1.00\textwidth]{skedge}
+            \includegraphics[width=1.00\textwidth]{images/skedge}
         \end{subfigure}
@@
 
-This paper will begin by discussing Markov chains and how we set out to use them for our goal of creating an assistive program editor.
+This paper\cite{escodegen} will begin by discussing Markov chains and how we set out to use them for our goal of creating an assistive program editor.
 We will then describe in detail how we implemented a Markov chain to house program syntax trees, how we were able to generate code using it, and finally we'll discuss our results and address the limitations of our product.
@@
 A Markov chain is a mathematical system that undergoes a stochastic process whereby
-the current state and a set of probabilities are used to determine what state
-to transition to next. In the most simple case of a Markov chain, the next action is based
-only upon the current state and all other previous actions are forgotten.
-Thus, this process is essentially ``memoryless'' because all past states can be
-forgotten once a new state is reached \cite{markov}. In this way, Markov chains can be likened
-to deterministic finite automata. So while they cannot remember a limitless amount of their history,
-calling them ``memoryless'' may be misleading, as they can indeed be implemented to remember a fixed
-size of their state history.
 
-Markov chains find application in many different fields.
-Already familiar with Markov chains used for generating random dialogue (by training a model based on bigrams and trigrams of common speech), we were interested in more novel domains.
-We chose to use a Markov chain to drive code generation, because, like text, we feel that code can be personal and expressive, and the idea of generating code possibilities from a training set seemed like exciting research.
-It was also challenging.
-Since randomly generating code tokens would most likely produce syntactically invalid programs, we looked at the abstract syntax trees of the programs and had to fit them into a Markov chain model.
-And as our research developed, we found ourselves with a new vision of an exciting project.
+\subsection*{Source code \& live site}
 
-\subsection{Emerging goal}
+The source code for Skedge is available online under an open
+source license:
+\url{https://github.com/RocHack/skedge}.\\
 
-For many novice programmers, programming can seem like a confusing and
-tedious activity. It can be unclear for many new programmers how they should
-begin and what code should come next. They may conceptually understand
-what the next step should be, and yet be unable to express it with code. Or, they
-could be unsure of what to do next, and perhaps giving code suggestions could inspire them
-and help them to understand it conceptually.
-
-Even as one becomes more
-adept at programming, he or she may notice typing the same or
-similar lines of code over and over again. This repetitiveness slows down
-the user and can make programming more tiresome a task.
-
-So, the prospect of using a Markov chain to generate code snippets in an assistive, predictive integrated development environment (IDE) for JavaScript seemed to be a natural progression from our research.
-Our final product, a Probabilistic Programming
-web application, allows a user to code in JavaScript by way of selecting from
-a list of suggestions for the user's next line of code. We imagine that this won't only be helpful for new programmers,
-but that this system would allow for expert programmers to write code with more ease and efficiency.
-
-% overview of markov chain
-
-\section{Tree-based Markov chain}
-
-\subsection{Abstract syntax trees}
-
-An abstract syntax tree (AST) is a tree-based, rather than text-based, representation of a program.
-Such a tree is generated using the grammar rules of a programming language and text input. In the tree, \emph{nodes}
-are language control structures such as {\tt for} loops, {\tt if} statements, variable and function
-declarations, and so on. The \emph{edges} connecting a parent node to its children represent the different \emph{features}
-of that node. For instance, a {\tt for} loop would have the features \emph{init} (e.g. {\tt var i = 0;}),
-\emph{test} (e.g. {\tt i < array.length;}), \emph{update} (e.g. {\tt i++}), and a \emph{body} (a list
-of statements that should be executed per loop iteration). To implement lists, we include the additional
-\emph{follow} edge as a feature to all statements, which would connect a statement node to the next statement
-in the list. If the statement is last in its container list, we connect its \emph{follow} edge to a special
-{\tt {\_end}} node. (See figure~\ref{fig:sample-ast} for an example of a AST.)
-
-% \begin{center}
-
-% \begin{tikzpicture}[level distance=1.5cm,
-% level 1/.style={sibling distance=3.5cm},
-% level 2/.style={sibling distance=1cm}]
-
-% \node (Root) {Program}
-% child {
-%     node {for} edge from parent node[draw=none] {\emph{body}}
-%     child { node {VD} edge from parent node[draw=none] {\emph{init}} }
-% };
-
-% \end{tikzpicture}
-
-% \end{center}
-
-Our system deals with programs at the level of AST's as they provide more structure and modularity than text.
-We use the syntax tree format of Mozilla's Parser API \cite{parser api}, and \emph{esprima} \cite{esprima}, an ECMAScript
-(the language standard of which JavaScript is an implementation)
-parser, to transform program source code into an AST of this format.
-We chose JavaScript because it is (a) dynamic and interpreted, (b) runnable in a web browser, and (c) because JavaScript AST tools already existed.
-
-\subsection{Our Markov chain}
-
-To build our Markov chain structure, we parse a corpus of JavaScript files into their
-respective AST's, and then collect statistics from these trees to establish probability distributions of nodes for a given AST context.
-
-An \emph{AST path} is a list of items, alternating between node types and features, that encodes the unique position of a 
-specific node in the AST. The \emph{probability distribution} for an AST path is the set of possible node types or values that would appear in the
-syntax tree following the given path, and the respective probabilities of their
-occurrence at that point. The model maps AST paths to the probability distributions of node types (or values, as we'll see later) for that given
-path.
-
-More formally, given a depth $d$ (indicating how many nodes we should backtrace) for a given
-${node}_d$ and feature ${edge}_d$, the generation of ${node}_d$'s child node given by
- ${edge}_d$ can be expressed by the following function $f$: \begin{align*}
-f({node}_1, {edge}_1, ..., {node}_{d}, {edge}_{d}) = \begin{dcases*}
-		{newnode_1},	&	$P({newnode_1})$ \\
-		{newnode_2},	&	$P({newnode_2})$ \\
-		...\\
-		{newnode_n},    &   $P({newnode_n})$ \\
-	\end{dcases*},
-\end{align*}
-
-where \begin{align*}
-\sum\limits_{i = 0}^n P({newnode_i}) = 1,
-\end{align*}
-
-\noindent and ${Path}({node}_1, {edge}_1) = {node}_2$, ${Path}({node}_2, {edge}_2) = {node}_3$, ..., ${Path}({node}_d, {edge}_d) = {newnode}_k$, where ${newnode}_k$ is randomly selected from ${newnode}_1, ..., {newnode}_n$ given the probabilities $P({newnode_1}), ..., P({newnode_n})$.
-From this it can be seen that generating programs is simply a repetitive application of this function, where for the chosen ${newnode}_k$, we can generate each child of ${newnode}_k$ with the following formula: \begin{align*} &\forall {edge} \in {Features}({newnode}_k), \\
-& \hspace{3em} {Path}({newnode}_k, {edge}) = f({node}_2, {edge}_2, ..., {node}_{d}, {edge}_{d}, {newnode}_k, {edge}).
-\end{align*}
-
-This process must eventually terminate as the probability of generating {\tt \_end} nodes for a \emph{follow} edge will always be non-zero (since the programs we parse into our corpus are surely finite) and such nodes have no features to be ``expanded.''
-
-\subsection{Implementation}
-
-\subsubsection{Data structure}
-
-Building the Markov chain model means being able to provide the probability distribution of new nodes for arbitrary input,
-as described by the function in Section 2.2. To do this, we construct a data structure using JavaScript Object Notation (JSON),
-a structural format useful for ease of specification and integration with JavaScript programs.
-
-The data structure is a hash table (also commonly called a map, dictionary, or associative array), that, on a given input, returns another hash table for paths that begin with that input.
-This behavior is repeated until the number of inputs reaches the desired depth, at which point a final hash table, mapping the node type to its probability distribution, is returned.
-This, in effect, is modularizing the function $f$ described above, so that we can reach the probability distribution with minimal code and data duplication.
-This is demonstrated by following example, given  ${depth} = 2$.
-\begin{align*}
-lookup(lookup(lookup(lookup({Model}, {node}_1), {edge}_1), {node}_2), {edge}_2) = \textbf{P}({newnodes}).
-\end{align*}
-
-\subsubsection{Building the model}
-
-We first obtain collection of JavaScript programs in AST form. We then run the following algorithm to construct our model:
-
-\begin{verbatim}
-procedure addToModel(node, path):
-    model = CorpusModel
-    for each elem in path[path.length-depth*2 .. path.length]:
-        model = model[elem]
-    model[node.type] += 1
-    model[_total] += 1
-    for each feature of node:
-        temp_path = path + [node.type, feature]
-        if node.feature is list of nodes:
-            for each item in node.feature:
-                addToModel(item, temp_path)
-                temp_path += [item.type, "FOLLOW"]
-            addToModel(END_NODE, temp_path)
-        else:
-            addToModel(node.feature, temp_path)
-
-for each parsed AST:
-    addToModel(AST.root, [])
-\end{verbatim}
-
-For each AST, we walk through every node in the tree by recursively exploring each feature\footnote{While we use \emph{follow} as if it were a feature edge, programs encoded in the AST's we parse use arrays of nodes (rather than linked-lists as we represent them) for a collection of statements.
-The given pseudocode demonstrates how we can transform an array into a linear sequence of edges.} of each node.
-In each call to explore a node's feature, we maintain the path up until that node by appending to it the node type and the name of the feature to reach the wanted child.
-We trim the beginning of the path off so that the length of the path is appropriate with respect to the chosen depth (i.e.
-so that $|path| = 2*d$).\footnote{Note that this requires that the path is at least of length $2*depth$.
-If the path is shorter than this, it represents not having at least $d$ levels of context.
-This can occur, say, for the first statement in the program, where there are no preceding nodes in its AST path.
-For these cases, we add the ``{\tt \_null}'' feature until we reach the desired depth (omitted in the pseudocode.)} We traverse through the layers of hash tables, creating the path elements if they don't already exist, add the node type to the probability distribution if it isn't already included, and then increment a counter associated with that node type.
-We also increment a ``{\tt \_total}'' counter for the convenience of the generator program.
-
-When this process is complete, the JSON object is saved to a file and can be augmented by more code samples thereafter.
-(An example of such a model is shown in figure~\ref{fig:sample-model} in the Appendix.
-In this model we see, for instance, that the path
-[\emph{Program, body, VariableDeclaration, declarations}] leads to
-\emph{VariableDeclarator} with probability 1. This indicates that a \emph{Program}
-node whose \emph{body} is a \emph{VariableDeclaration} always has a statement that is of
-type \emph{VariableDeclarator} as its first statement.)
-
-\subsection{Generating code}
-
-Code can be generated rather easily using the approach explained in Section 2.2.
-The program tree is built up one node at a time, each one generated based on its unique AST path.
-The following algorithm demonstrates this process.
-
-\clearpage
-
-\begin{verbatim}
-procedure generateNode(model, path):
-    for each elem in path[path.length-depth*2 .. path.length]:
-        model = model[elem]
-    nodeType = pickRandomNodeType(model)
-    node = {`type':nodeType}
-    for each feature of nodeType:
-        new_path = path + [node.type, feature]
-        if feature is a list feature:
-            nodes = []
-            while (newNode = generateNode(model, new_path) != END_NODE:
-                nodes += [nodeNode]
-                new_path += [newNode.type, "FOLLOW"]
-            node.feature = nodes
-        else:
-            node.feature = generateNode(model, new_path)
-    return node
-
-generateNode(CorpusModel, [])
-\end{verbatim}
-
-The algorithm looks very similar to the parse algorithm, as it is almost the same process, simply inverted.
-Rather than examine nodes, determine a path by building it up, and then increment a count in the model, this algorithm first finds the counts in the model for a \emph{given} path, generates a node, and generates nodes for all of that node's features recursively.
-(Since the final hash table is not a proper probability distribution, and is rather a tally of each node type, here the {\tt \_total} is useful for determining ratios.) The result is a generated AST for our program, which is then fed into \emph{escodegen} \cite{escodegen}, an ECMAScript code generator to output program source code that we can display and run.
-
-To generate random programs, we always begin with a ``Program'' node, and then add nodes to its \emph{body} feature (an array.) However, we quickly found that while entertaining, generating random programs had little potential for real application.
-Instead, we turned to generating code snippets (to be placed in specific contexts of the syntax tree.) Since our algorithm was general enough for this purpose, no more had to be done besides knowing the path for the node we wanted to generate.
-
-\subsubsection{Identifiers and literals}
-
-Naturally, identifiers and literals are a very large part of programming.
-Without them, code would be entirely unspecific, and would only consist of syntactic symbols and keywords.
-Yet they pose an issue with code generation, as we found generating identifiers or literals that are likely to be what the user is expecting an especially difficult problem to solve, let alone model.
-Struggling to come up with a simple, intelligent, and robust solution, and short on time, we fell back to using our existing Markov approach.
-
-In the current implementation of this system, following the paths [\emph{Identifier}, \emph{Name}], or [\emph{Literal}, \emph{Raw}] result not in node types, but rather in the text of the identifier names or literal values.
-This way, generating identifiers and literals is simply a probabilistic process like the rest of the code generation.
-Of course, this means that sometimes identifiers are added that are not in scope, that have mismatched types, or that simply make no semantic sense.
-(But adjusting $depth$ did seem to make programs more meaningful, as expected.
-This can be seen experimentally by modifying depth on our web IDE, which is discussed in Section 4.1.)
-
-%\lstinputlisting[caption=Scheduler, style=json]{../rand.js}
-
-\section{Applications}
-
-\subsection{\emph{vim} autocomplete script}
-
-We integrated our model into the text editor vim to offer autocompletion
-suggestions of programs upon the user's request, using vim's
-\texttt{completefunc} feature. When the user is editing a JavaScript file in vim with
-this feature, they may press a keyboard command to get a list of program
-snippets to choose from. The snippets are generated in the context of the
-program path at cursor's position. The user may pick one of the snippet
-suggestions to have it inserted into their code at their cursor. The autocomplete snippets are generated as follows:
-\begin{enumerate}
-	\item Parse the contents of the current file into an AST.
-	\item Find the node path corresponding to the cursor by traversing the AST.
-	\item Generate programs starting from the node path of the cursor.
-	\item Present the programs to the user as code completion snippets.
-\end{enumerate}
-
-\begin{figure}[ht]
-	\centering
-	\begin{tabular}{c c}
-		\begin{subfigure}[h]{6.25cm}
-			\centering
-			\includegraphics[width=1.00\textwidth]{vim1}
-			\caption{Inserting a {\tt for} loop} \label{fig:vim1}
-		\end{subfigure}
-		\begin{subfigure}[h]{6cm}
-			\centering
-			\includegraphics[width=1.00\textwidth]{vim2}
-			\caption{Inserting an object property} \label{fig:vim2}
-		\end{subfigure}
-	\end{tabular}
-	\caption{Screenshots of vim using autocomplete of programs}
-\end{figure}
-
-The vim autocomplete feature is limited to inserting program snippets on a single
-line, and requires the user to explicitly trigger the \texttt{completefunc}
-feature to be prompted with code snippets. To further explore using our model
-for editing, we created a web-based editor that more deeply integrates Probabilistic
-Programming.
-
-\subsection{Guided IDE}
-
-The Guided IDE offers users an environment for building a program entirely using
-our model, using program snippets generated at the cursor's context as building
-blocks. The editor presents a window for the program text, and a list of
-suggestion code snippets. The user can move a cursor through the code window at
-positions corresponding to boundaries of program nodes. As the user moves the
-cursor, the snippet suggestions are regenerated using the cursor's node path, to
-continuously provide suggestions relevant to the current context of the program. When the
-user selects a suggestion snippet, its AST is inserted into the program's AST at
-the cursor's position in the tree, and the code window is redrawn. The user can
-move the cursor within newly inserted blocks, so that the program can grow
-deeper. The initial corpus for the IDE's model consists of the JavaScript files
-that comprise our project.
-
-\begin{figure}[h]
-  \centering
-  \includegraphics[width=1.00\textwidth]{ide}
-  \caption{Screenshot of the web program editor GUI.
-The header along the top indicates the path (trimmed to the selected depth) to the current position of the cursor.
-Code snippets on the left can be navigated using the arrow keys.} \label{fig:screenshot}
-\end{figure}
-
-\section{Results}
-
-\subsection{Using the IDE}
-
-We can use the Guided IDE to generate programs. The programs look like actual
-programs but do not actually function unless the user is very careful about
-which snippets they pick. Crucially, the variable and identifier names that our
-system generates are not in general meaningful in the scope of the generated
-program.  Whereas with the vim autocompletions the identifiers would often be
-valid because because the corpus was the current file, with the larger corpus of
-the Guided IDE and writing a program from scratch, the user is less likely to
-find program snippets that use valid identifiers.
-
-\subsection{Limitations of the model}
-
-It is a limitation of the current model that identifier names are chosen
-probabilistically based on their position in the code, without regard to whether
-they are in scope and without refering to anything in another part of the
-program. Because of this limitation, the Guided IDE is not useful for practical
-purposes. We consider some alternative approaches for the Guided IDE using a
-similar model.
-
-\subsection{Alternative approaches for the IDE}
-
-\subsubsection{Prompt for identifiers}
-
-Rather than picking identifier names from the model, use placeholder identifiers
-and prompt the user to enter actual identifiers, with suggestions based on the
-current variables in scope and previously used identifiers. This would give the
-user the opportunity to create semantically valid programs, but the entering of
-identifier names could be tedious.
-
-\subsubsection{Stubs}
-
-Instead of having a movable cursor in the code window, consider stubs in the
-program. Each stub would be a marker in place of an AST node. The user would
-select a stub and be given choices of snippets to replace it. Selecting a
-snippet would move the current snippet into the program in place of the selected
-stub.  The generated snippets could themselves have stubs in them, so that the
-user may continue to complete stubs and expand their program.  At the level of
-variable identifiers, the snippet suggestions would be taken from the computed
-list of in-scope variables at the position of the selected stub, unless the
-completion was for a variable declaration, in which case the user would be
-prompted to name a new variable.
-
-This approach, while offering less places to insert code then the current
-approach, would allow for programs to more closely be generated from our model,
-since the stub completions would always be at leaf nodes of the AST, whereas in
-the current approach the user may choose to insert a snippet in the middle of a
-list, which means that the transition from the inserted snippet to the following
-node in the list does not in general follow from the model.
-
-While the stub approach would be more restrictive for the user as it requires
-them to build the program in a certain way, with proper handling of identifiers,
-it may provide for an interesting or useful paradigm for interacting with a program
-source.
-
-\subsection{Intelligent handling of identifiers and literals}
-
-We explored other approaches to solve the issue of generating meaningful identifiers and literals described before.
-
-\subsubsection{Secondary models}
-
-When generating identifiers, we were interested in generating meaningful ones rather than random ones (albeit from a weighted distribution.) We wanted to design a model that could keep track of when the same variables were used in the same (or similar) place.
-
-One idea was to, during the parse, keep track of the path at which each variable identifier was used, grouped by the identifier name.
-When an identifier needs to be generated, we can find that path in our secondary model, and if so, bind to a random (but weighted) identifier that has been used at that path, with the weight being based on the frequency of that identifier appearing at that path.
-For subsequent statements that need identifiers, the same process could occur, and, if the same variable group was chosen, it could generate the variable name that was bound to the first statement.
-
-Another idea also involved constructing a secondary model during the parse, that kept track of statements that used the same identifier.
-When such statements are found, it would compute the path between those statements and increment a counter associated that path.
-This would represent the likelihood of two nodes connected by that path using the same identifier.
-Then, during code generation, whenever an identifier is needed, the generator would make a list of all variables in scope at that point, and compute the path from the variable declaration (and/or last use) to the node being generated.
-It would look up each path in the secondary model and assign each of the in-scope variables a weight.
-A variable could then be chosen based on this probability.
-
-% We suspect that this may affect the complexity of our algorithms significantly.
-If what the system is doing now is analogous to the behavior of a context-free grammar (i.e.
-by recursively generating nodes), then detection of use of the same identifiers across nodes would involve pattern-matching and perhaps even analogy-solving.
-This could analogously be considered more like the action of a Turing machine, which would mean that the complexity of the process could be unbounded.
-
-Unfortunately, we were unable to implement these ideas or explore them further due to time constraints.
-
-\subsubsection{Using a higher-order model}
-
-Just as our approach of modeling a program with a tree-based Markov chain is a
-higher order model than a linear Markov chain of the program source code tokens,
-there may be applications of probabilistic modeling to higher-order
-representations of programs than the AST, such as a \emph{higher-order abstract syntax}, which promises to ``incorporate name binding information in a uniform and language generic way'' \cite{hosa}.
-Such an application may model the relationship between different
-parts of an AST, and handle scoping of variables to model identifiers
-abstractly.
-
-With a more sophisticated model of identifiers that could generate programs with
-identifiers in meaningful relation to each other, Probabilistic Programming could
-be used to generate semantically valid programs, for greater assistive,
-expressive, and programmatic power.
-
-\section{Conclusion}
-
-We consider the Probabilistic Programming web IDE a strong proof of concept for context-based code snippet generation.
-We are aware of some of its serious structural limitations, but we speculate that related concepts, such as genetic algorithms, might enjoy more success.
-Specifically, we are interested in the possibility of creating genetic algorithms that take advantage of this platform to build and ``evolve'' code, swapping parts of AST's in a Markovian-like process.
-Perhaps the evolution could even be given a goal; for instance, to satisfy a set of given test cases.
-Regardless, we believe that our research opens new possibilities with program generation.
-
-%Other applications of the model: genetic algorithms
-
-\subsection*{Source code \& demo}
-
-The source code for Probabilistic Programming is available online under an open
-source license:
-\url{https://github.com/RocHack/jschain}.\\
-
-\noindent The web IDE can be found at
-\url{https://csug.rochester.edu/u/clehner/jschain/web/}.
+\noindent The site can be found at
+\url{http://skedgeur.com}.
 
@@
 
-\begin{figure}[h!]
-	\caption{JSON AST sample}
-	\label{fig:sample-ast}
-	\centering
-    \vspace{-8pt}
-	An AST generated from the program ``{\tt var x = 34;}"
-
-	\lstinputlisting{ast-sample.json}
-\end{figure}
-
-\begin{figure}[h!]
-	\caption{JSON Markov model sample}
-	\label{fig:sample-model}
-	\centering
-    \vspace{-8pt}
-	A Markov chain model generated from the program ``{\tt var x = 34;}"
-
-	\lstinputlisting{model-sample.json}
-\end{figure}
-
 \end{document}

### 2016-04-09 01-introduction.tex
more formatting

--- 2016-04-09 paper.tex
@@
-\documentclass[titlepage]{article}
+\chapter{Introduction}
 
-\usepackage{setspace}
-
-\usepackage[utf8]{inputenc}
-\usepackage{amsmath}   
-\usepackage{mathtools}
-\usepackage{graphicx}
-\usepackage{subcaption}
-\usepackage{float}
-\usepackage{hyperref}
-\usepackage{listings}
-\usepackage[labelfont=bf]{caption}
-\usepackage[margin=3.5cm]{geometry}
-
-%\usepackage{qtree}
-%\usepackage{tikz}
-%\usepackage{tikz-qtree}
-
-%\usepackage[]{algorithm2e}
-
-\lstset{
-	basicstyle=\footnotesize\ttfamily,
-}
-
-\title{
-\vspace{80pt}\\
-\LARGE \bfseries SKEDGE
-\\\vspace{10pt}\Large Smarter course scheduling for our\\University of Rochester
-}
-
-\author{
-	Dan Hassin\\
-    \vspace{5pt}\\
-    Supervised by Professor Philip Guo\\
-    \vspace{5pt}\\
-    Department of Computer Science\\
-    University of Rochester\\
-    Rochester, New York\\
-}
-
-\date{April 12, 2016\\
-    \vspace{150pt}\\
-    submitted in partial fulfillment of\\
-    the requirements for the degree\\
-    honors bachelor of science\\
-}
-
-\begin{document}
-
-\maketitle
-
-\doublespacing
-
-% Direct search for specific course
-% Discovery, browsing, exploring
-% behavioral patterns
-
-% STAT: Find major of users, find how many non-major courses they found
-
-\begin{abstract}
-
-In this paper I present Skedge, a web application for students to comfortably and effectively engage with the University's course catalog.
-Skedge matches and surpasses the capabilities of the existing University tool for this purpose, ``Course Description / Course Schedule'' (CDCS) and presents its information in a more visually pleasing way.
-As a result, Skedge boasts strong user-retention rates, long session durations, and high student adoption despite having virtually no advertisement.
-Through collected usage data, I demonstrate that a) Skedge's differences from and additions to CDCS are usable and have real need, b) the two major use-cases associated with course browsing---direct search and exploratory search---are effectively accommodated by Skedge, and c) Skedge's search mechanism is user-friendly and self-teaches to users over time.
-
-\end{abstract}
-
-\begin{figure}[ht]
-    \centering
-    \begin{tabular}{c c}
-        \begin{subfigure}[h]{6.5cm}
-            \centering
-            \includegraphics[width=1.00\textwidth]{images/cdcs}
-        \end{subfigure}
-        \hspace{1em}
-        \begin{subfigure}[h]{7.5cm}
-            \centering
-            \includegraphics[width=1.00\textwidth]{images/skedge}
-        \end{subfigure}
-    \end{tabular}
-    \caption{CDCS (left) and Skedge (right) for the search query {\tt csc}.}
-\end{figure}
+This paper will begin by discussing Markov chains and how we set out to use them for our goal of creating an assistive program editor.
+We will then describe in detail how we implemented a Markov chain to house program syntax trees, how we were able to generate code using it, and finally we'll discuss our results and address the limitations of our product.
 
 \section{Introduction}
-
-This paper\cite{escodegen} will begin by discussing Markov chains and how we set out to use them for our goal of creating an assistive program editor.
-We will then describe in detail how we implemented a Markov chain to house program syntax trees, how we were able to generate code using it, and finally we'll discuss our results and address the limitations of our product.
 
@@
 
-\subsection*{Source code \& live site}
-
-The source code for Skedge is available online under an open
-source license:
-\url{https://github.com/RocHack/skedge}.\\
-
-\noindent The site can be found at
-\url{http://skedgeur.com}.
-
-\clearpage
-
-\begin{thebibliography}{1}
-
-	\bibitem{esprima} Ariya Hidayat {\em Esprima}
-		\\\url{http://esprima.org/}
-
-	\bibitem{markov} Takis Konstantopoulos {\em Introductory lecture notes on
-		Markov Chains and Random Walks.} Uppsala University,
-		\\\url{http://www2.math.uu.se/~takis/L/McRw/mcrw.pdf}
-
-	\bibitem{hosa} Frank Pfenning, Conal Elliott
-		{\em Higher-Order Abstract Syntax.}
-		ACM SIGPLAN Notices. Vol. 23. No. 7. ACM, 1988.
-		\\\url{http://www.cs.cmu.edu/~fp/papers/pldi88.pdf}
-
-	\bibitem{escodegen} Yusuke Suzuki {\em Escodegen}
-		\\\url{https://github.com/Constellation/escodegen}
-
-	\bibitem{parser api} Mozilla Developer Network {\em SpiderMonkey Parser API}
-		\\\url{https://developer.mozilla.org/en-US/docs/SpiderMonkey/Parser_API}
-
-\end{thebibliography}
-
-\clearpage
-\section*{Appendix}
-
-\end{document}
+\begin{figure}[ht]
+    \centering
+        \begin{subfigure}[h]{14cm}
+            \centering
+            \includegraphics[width=1.00\textwidth]{images/cdcs}
+        \end{subfigure}\\
+        \vspace{20pt}\\
+        \begin{subfigure}[h]{15cm}
+            \centering
+            \includegraphics[width=1.00\textwidth]{images/skedge}
+        \end{subfigure}
+        \vspace{20pt}\\
+    \caption{CDCS (top) and Skedge (bottom) for the search query {\tt csc}.}
+\end{figure}

### 2016-04-09 paper.tex
more formatting

@@
-\documentclass[titlepage]{article}
+\documentclass[titlepage]{report}
 
@@
 
-\begin{figure}[ht]
-    \centering
-    \begin{tabular}{c c}
-        \begin{subfigure}[h]{6.5cm}
-            \centering
-            \includegraphics[width=1.00\textwidth]{images/cdcs}
-        \end{subfigure}
-        \hspace{1em}
-        \begin{subfigure}[h]{7.5cm}
-            \centering
-            \includegraphics[width=1.00\textwidth]{images/skedge}
-        \end{subfigure}
-    \end{tabular}
-    \caption{CDCS (left) and Skedge (right) for the search query {\tt csc}.}
-\end{figure}
+\tableofcontents
 
-\section{Introduction}
+\include{01-introduction}
 
-This paper\cite{escodegen} will begin by discussing Markov chains and how we set out to use them for our goal of creating an assistive program editor.
-We will then describe in detail how we implemented a Markov chain to house program syntax trees, how we were able to generate code using it, and finally we'll discuss our results and address the limitations of our product.
+\clearpage
 
-\subsection{Markov chain}
-
-A Markov chain is a mathematical system that undergoes a stochastic process whereby
-
-\subsection*{Source code \& live site}
-
-The source code for Skedge is available online under an open
-source license:
-\url{https://github.com/RocHack/skedge}.\\
-
-\noindent The site can be found at
-\url{http://skedgeur.com}.
+\include{04-conclusion}
 

### 2016-04-09 01-grievances.tex
Chapters and sections placeholders

\chapter{Design as a reaction to CDCS}

Improvements were made by USING CDCS, (bottom-up, not top-down)!

%%%

\section{Modernity}

CDCS is an old system.

\subsection{GET requests vs. AJAX}

- Can use back button
- Can send a link to a course or search

\subsection{Built-in scheduler vs. browser extension}

- Better UX
- Data is centralized

\subsection{Mobile}

- Important nowadays
- Mobile traffic stat or smth

\subsection{Public API}

- Important nowadays, extends student possibility
- JSON
- Brief demo of API

%%%

\section{Usability}

\subsection{Data quality}

- Courses don't shout
- Typos in comments
- 12-hour time

\subsection{Section display}

- Grouped course sections
- Embedded labs (A/B too), workshops, \& recitations

\subsection{Course reference}

- Clickable/hoverable course links, professor searches

\subsection{Multiple schedule support}

- Old CDCS+betterCDCS system can't keep track of this, have conflicts when adding stuff

\subsection{Exporting to GCal, .ics, image}

- Mobile sync support
- Security: BetterCDCS export gcal is currently broken and sends netID in PLAINTEXT over http(!!!)

\subsection{Search}

Most important usability concern is finding courses.

%%%

\section{Search}

Use cases, natural language.

\subsection{Course selection criteria}

Narrowed it down to three criteria. Keep in mind that \emph{none} of the things listed below are supported by CDCS, and they are all supported by Skedge.

  \subsubsection{Requirements}

  - Finding crosslists
  - Clusters

  \subsubsection{Browsing}

  - ``New'' courses
  - ``Autofit'' search
  - Random
  - Sorts

  \subsubsection{Friends}

  - ``What are my friends taking?'' (``what are you taking this semester'' = probably most common smalltalk phrase uttered on campus)
  - ``What do my friends recommend?'' - ``have you taken this class, and if so, what did you think of it?''

\subsection{Natural language search}

%Figure

See figure.

  \subsubsection{Advantages}

  - 15 fields reduced to 1

  vs form entry:
  - Faster
  - More intuitive
  - More easily extendable

  \subsubsection{Disadvantages}

  Having to know the DSL, grammar ambiguities (can be solved with a `did you mean')

\subsection{Multipurpose}

Used by other links (instructors, course references) around the site

\subsection{Added features}

- CRN (!)
- Crosslist
- Class size

%%%

\section{Social}

\subsection{The issue}

  \subsubsection{Static image vs. live site}

  - Edits don’t update
  - Referencing courses

  \subsubsection{Finding common courses}

  - requires your friends to share their schedules on FB publicly and you to see their post
  - is schedule-first, not search-first
  - typically only occurs for the current semester

%figure

\subsection{Skedge Social}

% Walkthru

  \subsubsection{Friends' course enrollments}

  Mini-feed

  \subsubsection{Friends' course likes}

  \subsubsection{Likes \& enrollments embedded in results}

  \subsubsection{Personal schedule synchronization}

  \subsubsection{Privacy}

  \subsubsection{Notifications}

  %figs of the 2 types

### 2016-04-09 03-analytics.tex
Chapters and sections placeholders

\chapter{Data Analytics}

Hypotheses:

1. Skedge's differences from and additions to CDCS are usable and have real need

2. Skedge’s navigations-per-add and other metrics demonstrate effectiveness of the use cases
a) direct searching, and
b) course browsing

3. Skedge’s DSL is user-friendly; users learn more advanced search types over time by using it

\section{Usage}

\subsection{General}

Since November 3rd 2015 (137 days)
3,768 unique users
4,500 schedules
Average 90 sessions/day
Average 4.92 pages/session
Average 5:31 minutes/session
28\% of sessions are from new users

MOBILE RESULT

\subsection{Search}

% figs

  \subsubsection{Empty searches}

  Can learn from these
  Some funny ones

\subsection{Course blocks}

40\% of sessions have at least one block-click
Average of 4.94 block-clicks per session

\subsection{Social}

90 users have linked Skedge to Facebook
Since March 1st,
4,000+ visits (200 visits/day)
~60\% of visits to /social were returning visitors
90 overlays onto friends’ schedules
10 clicks to Facebook profiles :(
- get stats from the fb dashboard

\subsection{Conclusion}

Success! Considering skedge is OPTIONAL.
+ course blocks (obv usecase, can't click)
+ exports (not supported by thing)
+ mobile

\section{Navigations-per-add}

\subsection{Definitions}

A navigation is defined as
a search, or
a click on an instructor’s name, or
a click on a crosslisted or prerequisite course link

The navigations-per-{add, bookmark} measure is
the number of navigations a user took (within one session) until a course was {added, bookmarked}

\subsection{Trends}

%figs

\subsection{Breaking them apart}

  behavioral patterns
  Direct search for specific course
  Discovery, browsing, exploring

  \subsubsection{Direct searches}

  \subsubsection{Browse}

  % STAT: Find major of users, find how many non-major courses they found

\subsection{Conclusion}

%figs

Effective++

\section{Users' search types over time}

\subsection{Definitions}

Points for search by (omits number and dept.):

description
credits
crosslisted
CRN
instructor
title
year
term
‘random’
upper-level writing

“CSC” → 0
“MTH 165” → 0
“taught by hema” → 1   ✓    (2 searches) 
“random mur 1-2 credits” → 2   ✓    (1 search)

\subsection{Trends}

% fig

First increase (60.5% of users)
Median: 2 searches
Average: 4.23 searches
(Starting at 1 counts as an increase value of 0)

Second increase (7.9% of users)
Median: 8 searches
Average: 17.52 searches

\subsection{Conclusion}

DSL++

### 2016-04-09 conclusion.tex
Chapters and sections placeholders

\chapter{Conclusions}

\section{Proposal to the University}

\section{Resources}

\subsection*{Source code}

The source code for Skedge is available online under an open source license:\\
\url{https://github.com/RocHack/skedge}.

\subsection*{Live site}

\noindent The site can be found at: \url{http://skedgeur.com}.

### 2016-04-09 introduction.tex
Chapters and sections placeholders

--- 2016-04-09 01-introduction.tex
@@
 
-This paper will begin by discussing Markov chains and how we set out to use them for our goal of creating an assistive program editor.
-We will then describe in detail how we implemented a Markov chain to house program syntax trees, how we were able to generate code using it, and finally we'll discuss our results and address the limitations of our product.
+This paper will begin by
 
-\section{Introduction}
+\section{Space of course explorers and schedulers}
 
-\subsection{Markov chain}
+\section{Overview of CDCS}
 
-A Markov chain is a mathematical system that undergoes a stochastic process whereby
+\section{Overview of Skedge}
 

### 2016-04-09 paper.tex
Placeholders
Chapters and sections placeholders
Updates

@@
 
-%\usepackage{qtree}
-%\usepackage{tikz}
-%\usepackage{tikz-qtree}
-
-%\usepackage[]{algorithm2e}
-
 \lstset{
@@
 \title{
-\vspace{80pt}\\
-\LARGE \bfseries SKEDGE
-\\\vspace{10pt}\Large Smarter course scheduling for our\\University of Rochester
+\vspace{60pt}\\
+\huge \bfseries SKEDGE
+\\
+\vspace{10pt}
+\Large
+Smarter course scheduling for our\\
+University of Rochester
 }
@@
     \vspace{5pt}\\
-    Supervised by Professor Philip Guo\\
-    \vspace{5pt}\\
+    Supervised by\\
+    Professor Philip Guo\\
+    \vspace{2pt}\\
     Department of Computer Science\\
@@
     the requirements for the degree\\
-    honors bachelor of science\\
+    \emph{honors bachelor of science}\\
 }
@@
 
+%%%%%%%%%%%
+
+\onehalfspacing
+
+\pagenumbering{gobble}
+\setcounter{tocdepth}{1}
+\tableofcontents
+
+\listoftables
+
+\listoffigures
+
+\clearpage
+
+%%%%%%%%%%%%%%%%%%
+
 \doublespacing
 
-% Direct search for specific course
-% Discovery, browsing, exploring
-% behavioral patterns
-
-% STAT: Find major of users, find how many non-major courses they found
+\addcontentsline{toc}{chapter}{Abstract}
 
 \begin{abstract}
+
+\thispagestyle{plain}
+\pagenumbering{roman}
 
@@
 
-\tableofcontents
+%%%%%%%%%%%%%%%
 
-\include{01-introduction}
+\pagenumbering{arabic}
 
+\include{introduction}
 \clearpage
 
-\include{04-conclusion}
+\include{01-grievances}
+\clearpage
 
+\include{02-tech}
+\clearpage
+
+\include{03-analytics}
+\clearpage
+
+\include{04-future}
+\clearpage
+
+\include{conclusion}
 \clearpage
@@
 \begin{thebibliography}{1}
-
-	\bibitem{esprima} Ariya Hidayat {\em Esprima}
-		\\\url{http://esprima.org/}
 
@@
 
-	\bibitem{hosa} Frank Pfenning, Conal Elliott
-		{\em Higher-Order Abstract Syntax.}
-		ACM SIGPLAN Notices. Vol. 23. No. 7. ACM, 1988.
-		\\\url{http://www.cs.cmu.edu/~fp/papers/pldi88.pdf}
+\end{thebibliography}
 
-	\bibitem{escodegen} Yusuke Suzuki {\em Escodegen}
-		\\\url{https://github.com/Constellation/escodegen}
-
-	\bibitem{parser api} Mozilla Developer Network {\em SpiderMonkey Parser API}
-		\\\url{https://developer.mozilla.org/en-US/docs/SpiderMonkey/Parser_API}
-
-\end{thebibliography}
+\addcontentsline{toc}{chapter}{Bibliography}
 

### 2016-04-09 design.tex
Remove unneeded files

--- 2016-04-09 01-grievances.tex
@@
 
-Improvements were made by USING CDCS, (bottom-up, not top-down)!
+From its very inception, Skedge's functionality and visual design were driven by the shortcomings of CDCS.
+It was built \emph{bottom-up}, not \emph{top-down}, as every aspect of the application was either made as a reaction to a particular grievance in CDCS or as the natural evolution of an existing feature.
+Skedge is thus rooted in \emph{usability} derived from real need, not mere conjecture along the question ``what could students want?''.
+Its success with students, shown in Chapter 4, demonstrates that this usability extends beyond my own standard and can fulfill the various discovered use-cases of students in general.
+
+In this chapter, I will
 
@@
   - requires your friends to share their schedules on FB publicly and you to see their post
+
+%%%%%%%%%%%%%%
+
   - is schedule-first, not search-first

### 2016-04-09 tech.tex
Remove unneeded files

\chapter{Technical overview}

\section{Back-end}

Skedge's infrastructure is built

\emph{nginx}, \emph{unicorn}, \emph{Ruby on Rails}, \emph{PostgreSQL}, \emph{React.js}, \emph{Ahoy}, and Google Analytics.

\section{Front-end}

\section{Analytics}

### 2016-04-09 design.tex
API + more images

@@
-\chapter{Design as a reaction to CDCS}
+\chapter{Design in reaction to CDCS}
 
 From its very inception, Skedge's functionality and visual design were driven by the shortcomings of CDCS.
-It was built \emph{bottom-up}, not \emph{top-down}, as every aspect of the application was either made as a reaction to a particular grievance in CDCS or as the natural evolution of an existing feature.
+It was built \emph{bottom-up}, not \emph{top-down}---every aspect of the application was either made as a reaction to a particular grievance in CDCS or as the natural evolution of an existing feature.
 Skedge is thus rooted in \emph{usability} derived from real need, not mere conjecture along the question ``what could students want?''.
@@
 
-CDCS is an old system.
+CDCS is an old system, relatively speaking, and its recent development has been stagnant.
+It launched in 2009, seven years ago, and has hardly changed since.
+Figure \ref{fig:cdcs2010} shows CDCS in July 2010, which, besides the addition of a few search fields, is identical to its current version.
+Since its introduction in 2009, we have seen the rise of mobile devices into ubiquity, a boom in ``hacker culture'' and public APIs, and the capability for standalone web applications to be as sophisticated and dynamic as desktop-class applications without the aid of browser extensions.
+
+\begin{figure}[ht]
+  \centering
+    \fbox{
+      \includegraphics[width=10cm]{images/cdcs/2010}
+    }
+  \caption{CDCS in July 2, 2010, courtesy of \emph{Archive.org}}
+  \label{fig:cdcs2010}
+\end{figure}
+
+\subsection{Mobile}
+
+- Important nowadays
+
+- Mobile traffic stat or smth
+
+\begin{figure}[ht]
+  \centering
+  \begin{tabular}{c c}
+    \begin{subfigure}[h]{4cm}
+      \centering
+      \includegraphics[width=1.00\textwidth]{images/cdcs/mobile}
+      \caption{CDCS} \label{fig:cdcs-mobile}
+    \end{subfigure}
+    \begin{subfigure}[h]{3.9cm}
+      \centering
+      \includegraphics[width=1.00\textwidth]{images/skedge/mobile}
+      \caption{Skedge} \label{fig:sk-mobile}
+    \end{subfigure}
+  \end{tabular}
+  \caption{CDCS and Skedge running on a mobile device}
+\end{figure}
+
+\subsection{Public API}
+
+With the increasing number of attendees at University of Rochester's hackathons, it is clear that the University's ``hacker culture'' is growing (e.g.
+more students collaborating to build side-projects that integrate resources and servies often benefitting the student community), and projects that include the University's course catalog are becoming more common.
+Open-source and open-information services greatly help to foster such innovation, and having public APIs is essential toward this end.
+
+Skedge provides a public JSON API at the root URL \url{http://www.skedgeur.com/api/}, made at the request of a student that was interested in using its course data, and the API has already been used in projects by several other student groups.
+The endpoints included are \url{/api/courses?q=query} (Skedge's query language---described in detail in section 2.3---is supported here), \url{/api/departments?q=optional_query}, and \url{/api/instructors?q=optional_query}.
+
+\subsection{Built-in scheduler vs. browser extension}
+
+- Better UX
+
+- Data is centralized
 
@@
 - Can use back button
+
 - Can send a link to a course or search
-
-\subsection{Built-in scheduler vs. browser extension}
-
-- Better UX
-- Data is centralized
-
-\subsection{Mobile}
-
-- Important nowadays
-- Mobile traffic stat or smth
-
-\subsection{Public API}
-
-- Important nowadays, extends student possibility
-- JSON
-- Brief demo of API
 
@@
 
+- More concise, space efficient (collapsable)
+- Eliminates redundancy of same stuff (instructor, description, course title, etc)
+- Don't have to pay more attention that you should to whether it's a different course or the same one
+
+Note that in Figure \ref{fig:cdcs-sections}, the first two boxes are sections for the same course, and the next two are labs for that course.
+Four more lab sections and \emph{twenty} more workshop sessions for that same course follow below the truncated screenshot.
+Figure \ref{fig:sk-sections} demonstrates how this information can be conveyed more concisely.
+
+\begin{figure}[ht]
+  \centering
+  \begin{tabular}{c c}
+    \begin{subfigure}[h]{6cm}
+      \centering
+      \fbox{
+        \includegraphics[width=1.00\textwidth]{images/cdcs/sections}
+      }
+      \caption{Ungrouped sections in CDCS} \label{fig:cdcs-sections}
+    \end{subfigure}
+    \hspace{15pt}
+    \begin{subfigure}[h]{7cm}
+      \centering
+      \fbox{
+        \includegraphics[width=1.00\textwidth]{images/skedge/sections}
+      }
+      \caption{Grouped sections in Skedge} \label{fig:sk-sections}
+    \end{subfigure}
+  \end{tabular}
+  \caption{Section presentation on CDCS and Skedge}
+\end{figure}
+
 \subsection{Course reference}
+
+\begin{figure}[ht]
+  \centering
+  \fbox{
+    \includegraphics[width=8cm]{images/skedge/hover}
+  }
+  \caption{Hoverable and clickable course and instructor mentions in Skedge} \label{fig:sk-hover}
+\end{figure}
 

### 2016-04-09 introduction.tex
Remove unneeded files
API + more images

@@
 
+%%%
+
 \section{Overview of CDCS}
 
+\label{fig:cdcs-index}
+
+\subsection{``Better CDCS''}
+
+\begin{figure}[ht]
+    \centering
+    \fbox{
+        \includegraphics[width=1.00\textwidth]{images/cdcs/better}
+    }
+    \caption{Better CDCS, a separate browser extension that embeds buttons into the CDCS course results interface, allowing users to add courses to a locally-stored schedule}
+    \label{fig:cdcs-better}
+\end{figure}
+
+%%%
+
 \section{Overview of Skedge}
+
+Skedge is a website I developed in 2014 and have been maintaining and developing since.
+
+\ref{fig:sk-index}
 
@@
             \centering
-            \includegraphics[width=1.00\textwidth]{images/cdcs}
+            \fbox{
+                \includegraphics[width=1.00\textwidth]{images/cdcs/index}
+            }
+            \caption{CDCS}
+            \label{fig:cdcs-index}
+        \end{subfigure}\\
+        \vspace{10pt}\\
+        \begin{subfigure}[h]{14cm}
+            \centering
+            \fbox{
+                \includegraphics[width=1.00\textwidth]{images/skedge/index}
+            }
+            \caption{Skedge}
+            \label{fig:sk-index}
         \end{subfigure}\\
         \vspace{20pt}\\
-        \begin{subfigure}[h]{15cm}
-            \centering
-            \includegraphics[width=1.00\textwidth]{images/skedge}
-        \end{subfigure}
-        \vspace{20pt}\\
-    \caption{CDCS (top) and Skedge (bottom) for the search query {\tt csc}.}
+    \caption{CDCS and Skedge for the search query {\tt csc}}
 \end{figure}

### 2016-04-10 design.tex
Lots of work in the feature section

@@
 From its very inception, Skedge's functionality and visual design were driven by the shortcomings of CDCS.
-It was built \emph{bottom-up}, not \emph{top-down}---every aspect of the application was either made as a reaction to a particular grievance in CDCS or as the natural evolution of an existing feature.
+Skedge is built \emph{bottom-up}, not \emph{top-down}---every aspect of the application was either made as a reaction to a particular grievance in CDCS or as the natural evolution of an existing feature.
 Skedge is thus rooted in \emph{usability} derived from real need, not mere conjecture along the question ``what could students want?''.
@@
 
-In this chapter, I will
+In this chapter, I invite the reader along on a tour of these grievances and their remedies.
 
@@
 
-CDCS is an old system, relatively speaking, and its recent development has been stagnant.
+CDCS is an old system, relatively speaking, and its development on user-facing features has been almost entirely stagnant.
 It launched in 2009, seven years ago, and has hardly changed since.
 Figure \ref{fig:cdcs2010} shows CDCS in July 2010, which, besides the addition of a few search fields, is identical to its current version.
-Since its introduction in 2009, we have seen the rise of mobile devices into ubiquity, a boom in ``hacker culture'' and public APIs, and the capability for standalone web applications to be as sophisticated and dynamic as desktop-class applications without the aid of browser extensions.
+Yet, since its introduction in 2009, we have seen the rise of mobile devices into ubiquity, a boom in ``hacker culture'' and public APIs, and the capability for standalone web applications to be as sophisticated and dynamic as desktop-class applications without the aid of browser extensions.
+With this in mind, Skedge brings course scheduling to the modern era.
 
@@
     }
-  \caption{CDCS in July 2, 2010, courtesy of \emph{Archive.org}}
+  \caption{CDCS in July 2, 2010, virtually unchanged from today, courtesy of \emph{Archive.org}}
   \label{fig:cdcs2010}
@@
 
+\subsection{AJAX vs. GET requests}
+
+CDCS makes an AJAX request with every submitted search, meaning that the server receives the request and returns a response all without any page navigation (i.e.
+the URL stays the same and no new page is loaded as the search results are displayed).
+Skedge, however, makes a GET request for every search submission, meaning that the user's browser loads a new page that contains the results and whose URL reflects the search query.
+This simple technical design decision substantially increases usability for two reasons:
+
+\begin{enumerate}
+  \item Page navigations allow users to leverage their browser history as it was designed---after making several searches, CDCS users who use the back button on their browsers will be brought to the page loaded before the very first use of CDCS, possibly losing time spent in crafting sophisticated searches.
+Skedge users can go backwards and forwards through their search histories and scroll locations using native browser functionality.
+  \item Every search query has a unique URL (e.g.
+\url{http://skedgeur.com/?q=csc} for {\tt csc}), so users are able to send links to a specific course or search result to others.
+With CDCS, the URL remains \url{https://cdcs.ur.rochester.edu} throughout the duration of the session.
+\end{enumerate}
+
 \subsection{Mobile}
 
-- Important nowadays
+According to Mary Meeker's 2015 Mobile Technology Trends from Kleiner Perkins Caufield \& Byers\cite{kpcb}, in 2014, 51\% of adult time spent per day on the Internet was from a mobile device, versus 42\% spent on a desktop computer or laptop.
+Time spent on the Internet with mobile devices reached three hours per day in 2014, compared to less than one hour in 2010 (12\% mobile vs.
+75\% desktop/laptop time share).
 
-- Mobile traffic stat or smth
+Undeniably, supporting mobile devices and tablets in web applications is crucial for usability nowadays.
+Note how Skedge is responsive to the user's device in Figure \ref{fig:sk-mobile}, compared with CDCS's lack of mobile support in Figure \ref{fig:cdcs-mobile}.
+CDCS on mobile requires the user to pinch and drag around both to read results and to make new searchs, while Skedge adapts content to the device's screen and fixes the search bar to the top of the screen for easy access while browsing.
+
+Moreover, since no major mobile browser currently supports browser extensions (and if one did, the extensions themselves would most likely need to be re-architected), CDCS on a mobile device loses all scheduling functionality, unlike with Skedge on mobile.
 
@@
   \centering
+  \vspace{10pt}
   \begin{tabular}{c c}
-    \begin{subfigure}[h]{4cm}
+    \begin{subfigure}[h]{4.5cm}
       \centering
@@
     \end{subfigure}
-    \begin{subfigure}[h]{3.9cm}
+    \begin{subfigure}[h]{4.4cm}
       \centering
@@
 
-With the increasing number of attendees at University of Rochester's hackathons, it is clear that the University's ``hacker culture'' is growing (e.g.
-more students collaborating to build side-projects that integrate resources and servies often benefitting the student community), and projects that include the University's course catalog are becoming more common.
+With the increasing number of attendees at University of Rochester's hackathons, it is clear that the University's ``hacker culture'' is growing---more students are collaborating to build side-projects that integrate resources and servies often benefitting the student community.
 Open-source and open-information services greatly help to foster such innovation, and having public APIs is essential toward this end.
@@
 
-\subsection{Built-in scheduler vs. browser extension}
+\subsection{Built-in scheduler}
 
-- Better UX
+As explained in the introduction, Skedge offers users a course schedule right in the page, unlike CDCS which requires the Better CDCS browser extension for this functionality.
+Having a schedule native in the application has several advantages:
 
-- Data is centralized
+\begin{enumerate}
+\item Besides some CDCS users possibly not even knowing about Better CDCS, not requiring a browser extension provides for a faster and more seamless user onboarding, especially when building schedules on public computers when extensions can't always be installed.
 
-\subsection{GET requests vs. AJAX}
+\item Skedge accommodates a schedule into its design, whereas Better CDCS has to work around an interface that wasn't designed with one.
+As a result, Better CDCS has lower usability, requiring the user to toggle between search results and their schedule.
+Skedge, conversely, gives users immediate visual feedback on how a course would fit into their schedule.
 
-- Can use back button
+\item Schedule data is centralized on Skedge's servers as opposed to locally in a browser cache, meaning that it can synchronize across a user's devices or in sessions on public computers, persists browser resets, and can be easily publicly shared to other users.
 
-- Can send a link to a course or search
+\item Extensions like Better CDCS have limited browser support. Internet Explorer and mobile browsers are unsupported, for example.
+\end{enumerate}
 
@@
 
-\subsection{Data quality}
+\subsection{Visual presentation}
 
-- Courses don't shout
-- Typos in comments
-- 12-hour time
+Skedge offers several improvements over CDCS in the quality of its data presentation:
+
+\begin{enumerate}
+  \item Displaying information in a rigid, tabular way, CDCS does not leverage fonts and styling to adhere to typographical standards.
+Instead of using larger or bolder type, for instance, course titles are listed entirely in uppercase (e.g.
+``INTRO TO PROGRAMMING''), which has been shown to be less readable than lower-case text \cite{caps}.
+This problem is compounded when users browse through possibly hundreds of courses.
+Skedge displays properly capitalized titles styled with large type that helps users to quickly group and locate them.
+
+  \item While possibly not a fault of the CDCS system itself, there are very frequently typos or missing spaces in the ``comments'' section of courses, which Skedge corrects.
+
+  \item CDCS displays all course times in 24-hour time, which, despite being concise and unambiguous, is not what most US students are used to.
+Skedge displays 12-hour time with AM/PM, and prevents ambiguities through the course-in-schedule visualization on hover.
+\end{enumerate}
 
@@
 
-- Grouped course sections
-- Embedded labs (A/B too), workshops, \& recitations
+Often, courses are offered at multiple timeslots, sometimes taught by different instructors and in different rooms.
+These are called \emph{sections} of a course.
+CDCS displays each section in a discrete ``section box'' (all of which are nondistinct and have equal size), even if two sections pertain to the same course.
+(In this regard, CDCS should really be \emph{SDSS}, ``\emph{Section} Description / \emph{Section} Scheduler'', because it operates on the level of sections, not courses.) As a result, course descriptions (which can be lengthy), titles, prerequisites, comments, etc.
+are all repeated for every section of the course.
 
-- More concise, space efficient (collapsable)
-- Eliminates redundancy of same stuff (instructor, description, course title, etc)
-- Don't have to pay more attention that you should to whether it's a different course or the same one
+To make matters worse, many courses in the University course catalog include what I call \emph{subsections}---secondary sections associated with a course that must be registered for separately.
+Namely, these are labs, lab lectures, workshops, and recitations.
+Once again, CDCS displays \emph{all of these} as separate ``section boxes'' by default, and the course description is yet again repeated for each subsection (which, this time can be tens of times), wasting valuable page space.
 
-Note that in Figure \ref{fig:cdcs-sections}, the first two boxes are sections for the same course, and the next two are labs for that course.
+Collapsing subsections within courses can result in massive improvements in filtering the data most relevant to the user.
+For instance, the search {\tt csc} for Fall 2016 on CDCS results in 147 ``section boxes'', while Skedge only shows 45 ``\emph{course} boxes'', with subsections collapsed within their respective course.
+This triage reduces the data (noise, more correctly) displayed by ~70\%, and is even higher for departments with more abundant labs and workshops, such as Physics (Skedge: 35 vs.
+CDCS: 226, an ~85\% reduction), or Chemistry (Skedge: 25 vs.
+CDCS: 171, an ~86\% reduction).
+
+Skedge can reduce the amount of data to scroll through---and thus the time taken to do so---by six- or seven-fold (and possibly more, counting the attention users otherwise have to pay to distinguish course from subsections), so this design decision has a large usability payoff.
+
+Additionally, some Physics courses (for instance) follow the ``A / B'' subsection structure, where a student registered for an ``A Section'' (as opposed to the ``B Section'') must also register for an ``A Lab'' and ``A Workshop''.
+Skedge organizes subsections for these cases to help sort the two out, which get mixed up in CDCS's linear output.
+
+Note that in Figure \ref{fig:cdcs-sections} (CDCS), the first two boxes are sections for the same course, and the next two are labs for that course.
 Four more lab sections and \emph{twenty} more workshop sessions for that same course follow below the truncated screenshot.
-Figure \ref{fig:sk-sections} demonstrates how this information can be conveyed more concisely.
+Figure \ref{fig:sk-sections} (Skedge) demonstrates how this information can be conveyed more concisely.
 
@@
   \centering
+  \vspace{10pt}
   \begin{tabular}{c c}
@@
   \end{tabular}
-  \caption{Section presentation on CDCS and Skedge}
+  \caption{Section and subsection presentation in CDCS and Skedge}
 \end{figure}
@@
 
+\emph{Course mentions} will often appear in the prequisites, crosslists, comments, or description fields of a course (e.g.
+``\textbf{Prerequisites:} CSC 171 or equivalent; MTH 150 is REQUIRED'').
+Users frequently want to find out more information about mentioned courses (frequency shown in Chapter 4).
+In CDCS, because course mentions are displayed as ordinary plaintext, users have to scroll back up, make a search for \emph{that} course, and lose their current search context as a result.
+
+Skedge solves this by hyperlinking each course mention to a search query for its respective course, in the style of Wikipedia.
+Moreover, it protects users from a context-switch by displaying a lightweight popover with that course's information when the user hovers their cursor over the course mention (see Figure \ref{fig:sk-hover}).
+
 \begin{figure}[ht]
   \centering
+  \vspace{10pt}
   \fbox{
@@
   }
-  \caption{Hoverable and clickable course and instructor mentions in Skedge} \label{fig:sk-hover}
+  \caption{Hoverable and clickable course mention in the \emph{Prerequisites} field of a course} \label{fig:sk-hover}
 \end{figure}
-
-- Clickable/hoverable course links, professor searches
 
@@
 - Old CDCS+betterCDCS system can't keep track of this, have conflicts when adding stuff
+- Multiple schedules per semester (for different schedule possibilities)
 
-\subsection{Exporting to GCal, .ics, image}
+\subsection{Exporting}
+
+to Google Calendar, {\tt .ics}, image
 
 - Mobile sync support
-- Security: BetterCDCS export gcal is currently broken and sends netID in PLAINTEXT over http(!!!)
-
-\subsection{Search}
-
-Most important usability concern is finding courses.
-
-%%%
+- Export gcal is currently broken
+- Security: BetterCDCS ``sign in'' sends netID in PLAINTEXT over http(!!!)
 
@@
 
-Use cases, natural language.
+Of course, the most important usability concern for a course explorer/scheduler is effectively \emph{finding courses}.
 
-\subsection{Course selection criteria}
-
-Narrowed it down to three criteria. Keep in mind that \emph{none} of the things listed below are supported by CDCS, and they are all supported by Skedge.
-
-  \subsubsection{Requirements}
-
-  - Finding crosslists
-  - Clusters
-
-  \subsubsection{Browsing}
-
-  - ``New'' courses
-  - ``Autofit'' search
-  - Random
-  - Sorts
-
-  \subsubsection{Friends}
-
-  - ``What are my friends taking?'' (``what are you taking this semester'' = probably most common smalltalk phrase uttered on campus)
-  - ``What do my friends recommend?'' - ``have you taken this class, and if so, what did you think of it?''
+First, I will list some of the criteria students have for finding courses and how Skedge improves
 
@@
 
-See figure.
+See the appendix for a full list of these natural language things.
 
@@
 
+  - Class size sort
+
   \subsubsection{Disadvantages}
+  grammar ambiguities (can be solved with a `did you mean')
 
-  Having to know the DSL, grammar ambiguities (can be solved with a `did you mean')
+    fall of the roman empire thing
 
-\subsection{Multipurpose}
+  Having to know the DSL,
 
-Used by other links (instructors, course references) around the site
+  DSL can be self-taught by the search system being multi-purposed. Used by other links (instructors, course references) around the site. Chapter 4.
 
-\subsection{Added features}
+\subsection{Course selection criteria}
 
-- CRN (!)
-- Crosslist
-- Class size
+I have identified \emph{three} use-cases for course searching (the existence of and distinction between these cases will be demonstrated with collected usage data in Chapter 4).
+The three cases are \textbf{requirements}, \textbf{electives}, and \textbf{peer recommendations}.
+The Skedge DSL and other application features offer substantial improvements over CDCS for each of these cases.
+
+  \subsubsection{Requirements}
+
+  These are courses that are required for a student to complete their degree, and are typically searched for directly.
+The functionality required here is simple and is mostly satisfied by CDCS, but Skedge offers the following improvements to the process:
+
+  \begin{enumerate}
+    \item \emph{Crosslisted courses:} For students with more than one major and/or minor, searching for courses that are crosslisted between departments can be valuable in reducing their requirement load.
+This search filter is unsupported by CDCS, and is supported by Skedge using the operator ``{\tt x}'' (e.g.
+``{\tt csc x ece}'' for courses listed under both Computer Science and Electrical \& Computer Engineering departments).
+
+    \item \emph{Clusters:} Skedge already stores a users' previously taken courses, so it can intelligently suggest either already-completed clusters or courses that would complete clusters that are missing one or two courses.
+For students with many degree requirements already, this could greatly reduce time spent navigating the University's Cluster Search Engine (for which I also have a long list of grievances, but that lies outside the scope of this paper.)\footnote{This feature is under development and is not currently live.
+It was, incidentally, requested by a Skedge user.}
+
+    \item \emph{CRN:} Surprisingly, search by Course Reference Number is unsupported by CDCS. It is supported by Skedge by just searching the 5-digit number.
+  \end{enumerate}
+
+  \subsubsection{Electives}
+
+  Elective courses can be courses within a user's major. Here, Skedge offers search, filter, and sorting features that substantially aid users in browsing courses that they might want to take.
+
+  \begin{enumerate}
+
+    \item ``New'' courses
+    \item ``Autofit'' search\footnote{This feature is under development and is not currently live.}
+    \item search by instructor
+    \item Random
+    \item Sorts
+
+  \end{enumerate}
+
+  \subsubsection{Peer Recommendations}
+
+  CDCS currently has no support at all for peer course recommendation, a highly undervalued resource for course finding.
+Skedge implements peer recomendations through ``Skedge Social,'' a system detailed in the next section.
 
@@
 \section{Social}
+
+The question ``what are you taking this semester?'' is certainly the most common smalltalk phrase uttered on campus within the first few weeks of the semester.
+Besides the pure motivation  and it is not unreasonable to assume that students want to take classes that their friends are taking.
+
+\begin{itemize}
+  \item ``What are my friends taking?''
+  \item ``What do my friends recommend?''
+\end{itemize}
 

### 2016-04-10 paper.tex
Remove unneeded files
Lots of work in the feature section

@@
 
+\setlength{\skip\footins}{0.75cm}
+
 \include{introduction}
@@
 
-\include{01-grievances}
+\include{design}
 \clearpage
 
-\include{02-tech}
+\include{tech}
 \clearpage
 
-\include{03-analytics}
+\include{analytics}
 \clearpage
 
-\include{04-future}
+\include{future}
 \clearpage
@@
 
-	\bibitem{markov} Takis Konstantopoulos {\em Introductory lecture notes on
+	\bibitem{kpcb} Takis Konstantopoulos {\em Introductory lecture notes on
 		Markov Chains and Random Walks.} Uppsala University,
 		\\\url{http://www2.math.uu.se/~takis/L/McRw/mcrw.pdf}
+
+    \bibitem{caps} Wheildon, Colin (1995). \emph{Type and Layout: How Typography and Design Can Get your Message Across - Or Get in the Way.} Berkeley: Strathmoor Press. p. 62. ISBN 0-9624891-5-8.
 

### 2016-04-10 design.tex
More design

@@
 
-\section{Search}
+\section{Search friendliness}
 
-Of course, the most important usability concern for a course explorer/scheduler is effectively \emph{finding courses}.
+Of course, the most important usability concern for a course explorer/scheduler is being able to effectively \emph{find courses}.
 
-First, I will list some of the criteria students have for finding courses and how Skedge improves
+In this section, I will present the \emph{Skedge DSQL}, a domain-specific query language that is based on natural language.
+Next, I will demonstrate how Skedge leverages the DSQL to handle the three search criteria students have for finding courses better than CDCS does.
 
@@
 
-%Figure
+In the spirit of this chapter's theme, Skedge's search method was, again, primarily designed as a reaction to CDCS's.
+CDCS's search method is a 15-field form (see Figure \ref{fig:cdcs-search}), of which I only found myself regularly using two (one of them being the \emph{required} ``Year/Term'' field).
+This prompted me to closely examine the fields to a) determine redundancies between them and b) find how to make some of the valuable filter fields more usable.
+From this came Skedge's unified, natural language based search method, which I call the \emph{Skedge Domain-Specific Query Language (Skedge DSQL)}.
+Figure \ref{fig:sk-search} is a list of example possible searches shown to users as they type.
 
-See the appendix for a full list of these natural language things.
+\begin{figure}[ht]
+  \centering
+  \vspace{5pt}
+  \begin{tabular}{c c}
+    \begin{subfigure}[w]{4.5cm}
+      \centering
+      \fbox{
+        \includegraphics[width=3.3cm]{images/cdcs/search}
+      }
+      \caption{Form-based search in CDCS} \label{fig:cdcs-search}
+    \end{subfigure}
+    \hspace{5pt}
+    \begin{subfigure}[w]{8.5cm}
+      \centering
+      \fbox{
+        \includegraphics[width=1.00\textwidth]{images/skedge/search}
+      }
+      \caption{Examples of the Skedge DSQL} \label{fig:sk-search}
+    \end{subfigure}
+  \end{tabular}
+  \caption{Two philosophies of search method}
+\end{figure}
 
-  \subsubsection{Advantages}
+\begin{figure}[ht]
+  \centering
+  \vspace{10pt}
+  \includegraphics[width=12cm]{images/search-mapping}
+  \caption{Mapping from CDCS's form-based search to the Skedge DSQL} \label{fig:search-mapping}
+\end{figure}
+
+Figure \ref{fig:search-mapping} is a mapping of each CDCS field to an element of the Skedge DSQL.
+
+  \subsubsection{Advantages of the DSQL}
 
@@
 
-  \subsubsection{Disadvantages}
+  \subsubsection{Disadvantages of the DSQL}
   grammar ambiguities (can be solved with a `did you mean')
@@
 
-\section{Social}
+\section{``Skedge Social''}
 
@@
 
-\subsection{Skedge Social}
+\subsection{Skedge's solution}
 

### 2016-04-10 paper.tex
More design

@@
 As a result, Skedge boasts strong user-retention rates, long session durations, and high student adoption despite having virtually no advertisement.
-Through collected usage data, I demonstrate that a) Skedge's differences from and additions to CDCS are usable and have real need, b) the two major use-cases associated with course browsing---direct search and exploratory search---are effectively accommodated by Skedge, and c) Skedge's search mechanism is user-friendly and self-teaches to users over time.
+Through collected usage data, I demonstrate that \textbf{a)} Skedge's differences from and additions to CDCS are usable and have real need, \textbf{b)} the three major use-cases associated with course browsing---direct search, exploratory search, and peer recommendation---are effectively accommodated by Skedge, and \textbf{c)} Skedge's search mechanism is user-friendly and self-teaches to users over time.
 

### 2016-04-10 design.tex
DSL stuff

@@
   \vspace{10pt}
-  \includegraphics[width=12cm]{images/search-mapping}
+  \includegraphics[width=14cm]{images/search-mapping}
   \caption{Mapping from CDCS's form-based search to the Skedge DSQL} \label{fig:search-mapping}
@@
 
-Figure \ref{fig:search-mapping} is a mapping of each CDCS field to an element of the Skedge DSQL.
+Figure \ref{fig:search-mapping} is a mapping of each CDCS field to an element of the Skedge DSQL or unified search.
+Dashed lines that end abruptly are fields that were deemed unnecessary or redundant due to a Skedge-unique feature.
+To briefly sum up my rationale for each field:
+
+\onehalfspacing
+
+\subsubsection{Removed fields}
+
+\begin{enumerate}
+
+  \item \emph{School} (e.g.
+``College of Arts, Sciences, \& Engineering'', ``Eastman School of Music'', etc.) was removed in favor of always searching all schools, as is the default on CDCS.
+Searching by department codes (which are unique across schools) designates a school regardless.
+
+  \item \emph{Course Type} (e.g. ``Main Course'', ``Lab'', ``Workshop'' etc.) was removed because Skedge already embeds subsections into courses.
+
+  \item \emph{Status} (i.e.
+``Open'', ``Closed'', or ``Cancelled'') was removed because it is often useful to see closed courses if there is possibility of the instructor letting in more students, and searching for cancelled courses did not seem useful.
+
+  \item \emph{Division} (i.e. ``Humanities'', ``Social Sciences'', etc.) was removed because it seems too broad to be useful, but it could be added to the Skedge DSQL if there is demand.
+
+\end{enumerate}
+
+\subsubsection{Left side}
+
+The \emph{unified search} components can be entered directly into the single Skedge search field, along with the Skedge DSQL. It includes the CDCS fields:
+
+\begin{enumerate}
+  \item \emph{Subject} (department).
+  \item \emph{Course} (department code and/or course number, e.g. ``csc 171'').
+  \item \emph{Title} (course title).
+\end{enumerate}
+
+\subsubsection{Right side}
+
+The \emph{Skedge DSQL} is enterable using special keywords and, optionally, qualifiers for those keywords. It includes the CDCS fields:
+
+\begin{enumerate}
+  \item \emph{Year/term}, which is no longer required.
+It maps to the keywords {\tt \{spring,fall,summer,winter\}} and a four-digit year to specify term and year, and defaults to the current year and term.
+
+  \item \emph{Description}, which is specified by surrounding the query in quotes (e.g. ````buddhism'''').
+
+  \item \emph{Instructor name}, which is specified with the keyword {\tt taught by} and given the instructor name as a qualifier (e.g. ``taught by Guo'').
+
+  \item \emph{Day}, which can be specified with the keywords {\tt \{m,t,w,r,f,s,u\}} for filtering by specific days of the week (e.g. ``mwf'' for Monday-Wednesday-Friday classes only).
+
+  \item \emph{Start time between} is replaced by the keywords {\tt \{early,late\}} which \emph{sort} (ascending and descending, respectively) the results by start time.
+
+  \item \emph{Credit hours between} is replaced by the keyword {\tt credits}, qualified with a number or range (e.g. ``dan 1-2 credits'' for Dance classes with 1-2 credits).
+
+  \item \emph{Upper level writing} is replaced by the single-letter keyword {\tt w} (e.g. ``his w'' for History upper-level writing courses).
+\end{enumerate}
+
+\doublespacing
+
+As noted in Figure \ref{fig:search-mapping}, many use-cases for the \emph{Day} and \emph{Start time between} fields can be obsoleted with an \emph{autofit} feature that will be described later in the next subsection.
 
@@
 
-  - 15 fields reduced to 1
+    The main advantage of Skedge's search system is in its reduction of 15 fields to a single one.
+I hypothesize that since the vast majority of searches fit the department code / course number pattern (this hypothesis later supported in Chapter 4), the time required to force the user to select a term and then find the correct box to enter this query is wasted.
 
-  vs form entry:
-  - Faster
-  - More intuitive
-  - More easily extendable
-
-  - Class size sort
+    In addition, once learned and remembered, using a natural language based query language can be faster and more intuitive than clicking on form fields.
+And more importantly, the Skedge DSQL is more easily extendable as new search features are invented.
+For CDCS, the more form fields that are added to the sidebar, the less usable its user interface becomes.
+For Skedge, while it may take longer for users to learn a bigger DSQL, there is no issue with crowding user interface real estate.
 
   \subsubsection{Disadvantages of the DSQL}
-  grammar ambiguities (can be solved with a `did you mean')
 
-    fall of the roman empire thing
+    One disadvantage of Skedge's DSQL is the possible grammatical ambiguity of some queries.
+For instance, the query ``Fall of the Roman Empire'' could be searching for courses with that as the title, or searching for courses in the \emph{fall term} with the title ``of the Roman Empire''.
+Unless multiple queries are run and more complex natural language processing is done, the system would have a hard time disambiguating between the two.
+This could also be solved with a ``did you mean?'' feature, which disambiguates a query into every possible DSQL breakup.
 
-  Having to know the DSL,
-
-  DSL can be self-taught by the search system being multi-purposed. Used by other links (instructors, course references) around the site. Chapter 4.
+    Another notable disadvantage of the DSQL is having to know it, and it having a possibly steep learning curve for some users.
+However, the argument will be made in Chapter 4 that simply by using the site over time, users will self-learn the Skedge DSQL.
+One of the ways this is possible is thanks to Skedge's search system being \emph{multi-purposed}.
+Hyperlinks around the site---instructor names (which link to ``{\tt taught by <name>}''), course references, or schedule course blocks for example---all direct the user to an ordinary search results page with the search field populated, signalling to the user that such a search is possible.
 
@@
     \item Random
-    \item Sorts
+    \item Sorts (class size sort)
 

### 2016-04-10 design.tex
multischedule stuff

@@
 
-Skedge solves this by hyperlinking each course mention to a search query for its respective course, in the style of Wikipedia.
-Moreover, it protects users from a context-switch by displaying a lightweight popover with that course's information when the user hovers their cursor over the course mention (see Figure \ref{fig:sk-hover}).
+Skedge solves this by linking each course mention to its search query, in the style of Wikipedia.
+Moreover, it prevents context-switches by displaying a popover containing information on that course when the user hovers their cursor over the course mention (see Figure \ref{fig:sk-hover}).
 
@@
   \centering
-  \vspace{10pt}
+  \vspace{2pt}
   \fbox{
@@
 
-- Old CDCS+betterCDCS system can't keep track of this, have conflicts when adding stuff
-- Multiple schedules per semester (for different schedule possibilities)
+With an active schedule still in use, students often want to plan their next semester's schedule.
+This is impossible with the Better CDCS extension, which instead adds courses from both terms to the same schedule, sometimes causing non-existent conflicts.
+A CDCS user must clear their entire schedule out before scheduling a new one.
+On Skedge, however, when a user adds a section of a term for which they do not yet have a schedule, a new one will automatically be created for that term.
+Users are also able to access their old schedules.
 
-\subsection{Exporting}
-
-to Google Calendar, {\tt .ics}, image
-
-- Mobile sync support
-- Export gcal is currently broken
-- Security: BetterCDCS ``sign in'' sends netID in PLAINTEXT over http(!!!)
+Additionally, a common feature request for Skedge is the implementation of multiple schedules \emph{per} term, allowing users to experiment with different schedule possibilities.\footnote{Currently under development, but simple to implement given the existing schedule management structure.}
 
@@
 
-\subsubsection{Left side}
+\subsubsection{Unified Search (left side)}
 
@@
 
-\subsubsection{Right side}
+\subsubsection{Skedge DSQL (right side)}
 

### 2016-04-10 design.tex
electives

@@
 
-  Elective courses can be courses within a user's major. Here, Skedge offers search, filter, and sorting features that substantially aid users in browsing courses that they might want to take.
+  Elective courses can either be courses that are required for a major or minor, but are chosen by the student, or they can be courses of particular interest to students who have the flexibility to take them.
+For this use-case, Skedge offers search, filter, and sorting features that substantially aid users in browsing through courses and discovering ones they might want to take.
+Note that none of the following are supported by CDCS.
 
@@
 
-    \item ``New'' courses
-    \item ``Autofit'' search\footnote{This feature is under development and is not currently live.}
-    \item search by instructor
-    \item Random
-    \item Sorts (class size sort)
+    \item The Skedge DSQL includes the ``new'' keyword, which qualifies an accompanying search by only displaying the courses that were not offered the previous year for that term.
+For instance, searching ``new csc'' today (Fall 2016) would show Fall Computer Science courses that have been added to the catalog or changed since Fall 2015.
+    \item The Skedge DSQL could include an ``autofit'' keyword, which would take into account the user's current schedule, and only show sections that would not conflict with any already scheduled times.\footnote{This feature is under development and is not currently live.}
+    \item The Skedge DSQL includes the ``random'' keyword, which displays a single random course matching the accompanying query.
+For instance, ``random csc'' would show a random course in Computer Science for the current term.
+Used in conjunction with ``autofit'', it can a powerful way to find interesting new classes that are available for users to try.
+    \item The Skedge DSQL includes sorting by class size---useful for students trying to find smaller, discussion based classes---or by class starting time---for students with preferences of morning or afternoon/evening classes.
+    \item Because instructor names appearing in courses automatically link to searches for more courses taught by that instructor, it is easier and more natural to browse by instructor.
+[NEED STAT HERE]
 

### 2016-04-10 design.tex
work on social... so slow

@@
   CDCS currently has no support at all for peer course recommendation, a highly undervalued resource for course finding.
-Skedge implements peer recomendations through ``Skedge Social,'' a system detailed in the next section.
+Skedge implements peer recomendations through Skedge Social, a system detailed in the next section.
 
@@
 
-\section{``Skedge Social''}
+\section{Peer connectivity}
 
-The question ``what are you taking this semester?'' is certainly the most common smalltalk phrase uttered on campus within the first few weeks of the semester.
-Besides the pure motivation  and it is not unreasonable to assume that students want to take classes that their friends are taking.
+The question ``what are you taking this semester?'' may possibly be the most common phrase uttered in campus smalltalk within the first few weeks of a semester (as well as its analogue ``what are you taking \emph{next} semester?'' in the last half of a semester).
+I hypothesize that this question seeks to do the following:
 
 \begin{itemize}
-  \item ``What are my friends taking?''
-  \item ``What do my friends recommend?''
+  \item Satisfy curiosity
+  \item Find common courses
+  \item Find common free time to meet
+  \item Seek or solicit advice about courses
 \end{itemize}
 
-\subsection{The issue}
+However, students' social lives do not solely exist in-person---social media dominates college social lives etc, v impt to consider [CITE].
+The ``what are you taking'' question typically translates to social media in the form of the schedule image exported from Better CDCS posted to Facebook (e.g.
+Figure \ref{fig:cdcs-social}).
+There are several limitations to this model.
+
+\begin{figure}[ht]
+  \centering
+  \vspace{10pt}
+  \includegraphics[width=6cm]{images/cdcs/social}
+  \caption{Static image} \label{fig:cdcs-social}
+\end{figure}
+
+\subsection{Limitations of the Facebook model, remedied by Skedge}
 
@@
 
-  - Edits don’t update
-  - Referencing courses
+    As mentioned above, the schedules generated with Better CDCS cannot be shared publicly, and therefore students can only post a static image of their schedule at the time of posting.
+This excludes the possibility of edits to the schedule being reflected in the same post to their friends.
 
-  \subsubsection{Finding common courses}
+    Moreover, referencing courses (i.e.
+looking up more information on a course a friend is taking) is not very usable under this model, since links cannot be embedded in a static image, and requires a manual search.
 
-  - requires your friends to share their schedules on FB publicly and you to see their post
+    Skedge allows users to generate a public link to a view-only, larger version of their schedules (see Figure \ref{fig:sk-share}).
+If a user visiting others' share pages also has their schedule saved on Skedge, the option to ``overlay'' the user's schedule ontop of the share page's schedule is available, making it easier for friends to find common free time (see Figure \ref{fig:sk-overlay}).
 
-%%%%%%%%%%%%%%
+    When this link is posted on Facebook, Skedge will include a rendered image of the schedule in the post, maintaining the element of visual bait with posting a static image.
 
-  - is schedule-first, not search-first
-  - typically only occurs for the current semester
+    \begin{figure}[H]
+      \centering
+      \begin{tabular}{c c}
+        \begin{subfigure}[w]{10.5cm}
+          \centering
+          \fbox{
+            \includegraphics[width=1.00\textwidth]{images/skedge/share}
+          }
+          \caption{Skedge's schedule ``share'' page, publicly accessible at \url{http://skedgeur.com/1234}, with the last four digits being the randomly generated ID for that schedule.
+Hovering over a course shows information about it, and clicking searches for it.} \label{fig:sk-share}
+        \end{subfigure} \vspace{15pt}\\
+
+        \begin{subfigure}[w]{10.5cm}
+          \centering
+          \fbox{
+            \includegraphics[width=1.00\textwidth]{images/skedge/share-compare}
+          }
+          \caption{The ``overlay'' feature enabled, superimposing the user's schedule over the share page's schedule.} \label{fig:sk-overlay}
+        \end{subfigure}
+      \end{tabular}
+      \caption{Skedge's per-schedule share page}
+    \end{figure}
+
+  \subsubsection{Robustness of finding courses taken by peers}
+
+  This model imposes several obstacles that prevents users from effectively finding the courses their peers are taking. Doing so within this model:
+
+  \begin{enumerate}
+    \item requires that \textbf{a)} peers share their schedules on Facebook publicly and \textbf{b)} the user incidentally checks Facebook while that post is still deemed relevant via Facebook's algorithms.
+
+    \item is exclusively schedule-oriented, not search-oriented.
+Course discovery happens from seeing peers' schedules first, not from \emph{browsing} first and being seeing, for each class, that their peers may be taking it.
+
+    \item typically only occurs for the \emph{current} semester.
+Under the Facebook model, it can be very difficult, nigh impossible, for students looking for opinions, tips, or even academic help for a specific class to find others who had taken that class before.
+  \end{enumerate}
+
+  Solving these issues would require platform infrastructure external to Facebook, but would still greatly benefit from its ubiquity and centralized role in social media.
+Since Skedge already saves and archives users' schedules, and Facebook has an integration API, building this platform into Skedge proved to be that opportunity.
 
@@
 
-\subsection{Skedge's solution}
+\subsection{Skedge Social}
 
-% Walkthru
+The \emph{Skedge Social} platform aims to leverage existing social media paradigms (such as the \emph{news feed}, \emph{likes}, and \emph{requests}) to more seamlessly give users answers to the questions alluded to earlier---``What are my friends taking?'' and ``What do my friends recommend?''.
+Skedge Social consists of the following several components:
 
-  \subsubsection{Friends' course enrollments}
+  \subsubsection{One-click integration}
+
+    Login
+
+    \begin{figure}[H]
+      \centering
+      \fbox{
+        \includegraphics[width=14cm]{images/skedge/social-login}
+      }
+      \caption{Skedge Social landing page for a user whose account has not yet been integrated with Facebook.
+Skedge Social appears in the context of the ordinary Skedge interface, and is accessible with the globe icon to the right of the searchbar.} \label{fig:sk-social-login}
+    \end{figure}
+
+  \subsubsection{Peer schedules}
 
@@
 
-  \subsubsection{Friends' course likes}
+    \begin{figure}[H]
+      \centering
+      \fbox{
+        \includegraphics[width=10cm]{images/skedge/social}
+      }
+      \caption{Skedge Social's schedule feed (the default of four tabs) for a user who has integrated their Skedge account with Facebook.} \label{fig:sk-social}
+    \end{figure}
 
-  \subsubsection{Likes \& enrollments embedded in results}
+  \subsubsection{Peer-liked courses}
+
+  \subsubsection{Likes \& enrollments embedded in search results}
+
+  ugh screenshot
 
@@
 
-  %figs of the 2 types
+  \begin{figure}[H]
+    \centering
+    \begin{tabular}{c c}
+      \begin{subfigure}[w]{3cm}
+        \centering
+        \includegraphics[height=1.5cm]{images/skedge/social-note1}
+        \caption{Note1.} \label{fig:sk-social-note1}
+      \end{subfigure}
+
+      \begin{subfigure}[w]{3cm}
+        \centering
+        \includegraphics[height=1.5cm]{images/skedge/social-note2}
+        \caption{Note2.} \label{fig:sk-social-note2}
+      \end{subfigure}
+    \end{tabular}
+    \caption{Notifications}
+  \end{figure}

### 2016-04-10 introduction.tex
Lots of work in the feature section
work on social... so slow

@@
 
-\section{Space of course explorers and schedulers}
+\section{Overview of CDCS}
 
-%%%
-
-\section{Overview of CDCS}
+CDCS is the University's official tool for
 
@@
 \subsection{``Better CDCS''}
-
-\begin{figure}[ht]
-    \centering
-    \fbox{
-        \includegraphics[width=1.00\textwidth]{images/cdcs/better}
-    }
-    \caption{Better CDCS, a separate browser extension that embeds buttons into the CDCS course results interface, allowing users to add courses to a locally-stored schedule}
-    \label{fig:cdcs-better}
-\end{figure}
 
@@
 Skedge is a website I developed in 2014 and have been maintaining and developing since.
+
+Bookmarks
+
+Students, parents, department coordinators, and faculty can all benefit from such tool improvements.
 
@@
             }
-            \caption{CDCS}
+            \caption{CDCS, with the search query {\tt csc}}
             \label{fig:cdcs-index}
@@
             \fbox{
+                \includegraphics[width=1.00\textwidth]{images/cdcs/better}
+            }
+            \caption{Better CDCS, a separate browser extension that embeds buttons into the CDCS course results interface, allowing users to add courses to a locally-stored schedule}
+            \label{fig:cdcs-better}
+        \end{subfigure}
+    \caption{CDCS and Better CDCS in their current states}
+\end{figure}
+
+\begin{figure}[ht]
+    \centering
+    \fbox{
         \includegraphics[width=1.00\textwidth]{images/skedge/index}
     }
-            \caption{Skedge}
+    \caption{Skedge, with the search query {\tt csc} and the user's current schedule on the right}
     \label{fig:sk-index}
-        \end{subfigure}\\
-        \vspace{20pt}\\
-    \caption{CDCS and Skedge for the search query {\tt csc}}
 \end{figure}

### 2016-04-11 design.tex
Digging into social

@@
 For this use-case, Skedge offers search, filter, and sorting features that substantially aid users in browsing through courses and discovering ones they might want to take.
-Note that none of the following are supported by CDCS.
+Note that none of the following features are supported by CDCS.
 
@@
     \item The Skedge DSQL includes sorting by class size---useful for students trying to find smaller, discussion based classes---or by class starting time---for students with preferences of morning or afternoon/evening classes.
-    \item Because instructor names appearing in courses automatically link to searches for more courses taught by that instructor, it is easier and more natural to browse by instructor.
-[NEED STAT HERE]
+    \item Instructor names appearing in courses automatically link to searches for more courses taught by that instructor, which is easier than having to use the CDCS ``instructor'' field to make a separate search.
 
@@
 The question ``what are you taking this semester?'' may possibly be the most common phrase uttered in campus smalltalk within the first few weeks of a semester (as well as its analogue ``what are you taking \emph{next} semester?'' in the last half of a semester).
-I hypothesize that this question seeks to do the following:
-
-\begin{itemize}
-  \item Satisfy curiosity
-  \item Find common courses
-  \item Find common free time to meet
-  \item Seek or solicit advice about courses
-\end{itemize}
 
@@
 
-\begin{figure}[ht]
+\begin{figure}[H]
   \centering
-  \vspace{10pt}
-  \includegraphics[width=6cm]{images/cdcs/social}
-  \caption{Static image} \label{fig:cdcs-social}
+  \vspace{5pt}
+  \begin{tabular}{c c}
+
+    \begin{subfigure}[w]{7cm}
+      \centering
+      \includegraphics[width=6.5cm]{images/cdcs/social}
+      \caption{Sharing a static schedule image generated by Better CDCS on Facebook} \label{fig:cdcs-social}
+    \end{subfigure}
+
+    \hspace{10pt}
+
+    \begin{subfigure}[w]{7.1cm}
+      \centering
+      \includegraphics[width=6.6cm]{images/skedge/facebook}
+      \caption{Sharing a Skedge schedule on Facebook, linked to the live and dynamic schedule share page on Skedge} \label{fig:sk-fb}
+    \end{subfigure}
+
+  \end{tabular}
+  \caption{Comparison of sharing schedules on Facebook}
 \end{figure}
@@
 
-    When this link is posted on Facebook, Skedge will include a rendered image of the schedule in the post, maintaining the element of visual bait with posting a static image.
+    When this link is posted on Facebook, Skedge will include a rendered image of the schedule in the post, maintaining the element of visual bait with posting a static image (see Figure \ref{fig:sk-fb}).
 
@@
       \begin{tabular}{c c}
-        \begin{subfigure}[w]{10.5cm}
+        \begin{subfigure}[w]{11cm}
           \centering
@@
 Hovering over a course shows information about it, and clicking searches for it.} \label{fig:sk-share}
-        \end{subfigure} \vspace{15pt}\\
+        \end{subfigure} \vspace{25pt}\\
 
-        \begin{subfigure}[w]{10.5cm}
+        \begin{subfigure}[w]{11cm}
           \centering
@@
     \item typically only occurs for the \emph{current} semester.
-Under the Facebook model, it can be very difficult, nigh impossible, for students looking for opinions, tips, or even academic help for a specific class to find others who had taken that class before.
+Under the Facebook model, it can be very difficult, nigh impossible, for students seeking advice, opinions, or even academic help for a specific class to find others who had taken that class before.
   \end{enumerate}
@@
 The \emph{Skedge Social} platform aims to leverage existing social media paradigms (such as the \emph{news feed}, \emph{likes}, and \emph{requests}) to more seamlessly give users answers to the questions alluded to earlier---``What are my friends taking?'' and ``What do my friends recommend?''.
-Skedge Social consists of the following several components:
 
-  \subsubsection{One-click integration}
-
-    Login
+Logging into Skedge Social is very simple---if the user is already logged into Facebook, it just requires one button of authorization (see Figure \ref{fig:sk-social-login}).
 
@@
       \fbox{
-        \includegraphics[width=14cm]{images/skedge/social-login}
+        \includegraphics[width=9cm]{images/skedge/social-login}
       }
-      \caption{Skedge Social landing page for a user whose account has not yet been integrated with Facebook.
+      \caption{Skedge Social landing page (\url{http://skedgeur.com/social}) for a user whose account has not yet been integrated with Facebook.
 Skedge Social appears in the context of the ordinary Skedge interface, and is accessible with the globe icon to the right of the searchbar.} \label{fig:sk-social-login}
@@
 
-  \subsubsection{Peer schedules}
+  \subsubsection{Schedule feed}
 
-  Mini-feed
+  Immediately after logging in, the user will see the ``schedule feed,'' the list of the user's peers' current schedules.
+This takes from a familiar paradigm of the \emph{feed} in social networks, which is routinely scrolled for updates.
+For each schedule in the feed, the user can easily visualize their own schedule side-by-side (using with the same ``overlay'' mechanic explained earlier), access the Facebook profile of the schedule's owner, and see the list of that users' course like.
+
+    \begin{figure}
+      \centering
+      \fbox{
+        \includegraphics[width=10cm]{images/skedge/social}
+      }
+      \caption{Skedge Social's schedule feed (the default of four tabs) for a user who has integrated their Skedge account with Facebook.
+The user can toggle between a peer's schedule and the list of their \emph{liked} courses.} \label{fig:sk-social}
+    \end{figure}
+
+  \subsubsection{Search result integration}
+
+  Skedge Social also integrates into course search results, showing users on the platform their Facebook friends who are taking, have taken, or \emph{liked} each particular course (see Figure \ref{fig:sk-social-course}).
+This has several advantages:
+
+  \begin{itemize}
+
+  \item It enhances the user's social engagement and allows them to better keep up with peers.
+  \item It promotes and facilitates communication for users either seeking or soliciting advice regarding courses (sending a Facebook message is merely two clicks away).
+  \item It brings attention to courses which are particularly popular or liked among the user's community of peers, increasing the chances of serendipitous course discovery.
+
+  \end{itemize}
 
@@
       \fbox{
-        \includegraphics[width=10cm]{images/skedge/social}
+        \includegraphics[width=12cm]{images/skedge/social-course}
       }
-      \caption{Skedge Social's schedule feed (the default of four tabs) for a user who has integrated their Skedge account with Facebook.} \label{fig:sk-social}
+      \caption{Skedge Social integrating into a course search result} \label{fig:sk-social-course}
     \end{figure}
-
-  \subsubsection{Peer-liked courses}
-
-  \subsubsection{Likes \& enrollments embedded in search results}
-
-  ugh screenshot
 

### 2016-04-11 design-modernity.tex
Split design files

--- 2016-04-10 design.tex
@@
-\chapter{Design in reaction to CDCS}
-
-From its very inception, Skedge's functionality and visual design were driven by the shortcomings of CDCS.
-Skedge is built \emph{bottom-up}, not \emph{top-down}---every aspect of the application was either made as a reaction to a particular grievance in CDCS or as the natural evolution of an existing feature.
-Skedge is thus rooted in \emph{usability} derived from real need, not mere conjecture along the question ``what could students want?''.
-Its success with students, shown in Chapter 4, demonstrates that this usability extends beyond my own standard and can fulfill the various discovered use-cases of students in general.
-
-In this chapter, I invite the reader along on a tour of these grievances and their remedies.
-
-%%%
-
 \section{Modernity}
@@
 \end{enumerate}
-
-%%%
-
-\section{Usability}
-
-\subsection{Visual presentation}
-
-Skedge offers several improvements over CDCS in the quality of its data presentation:
-
-\begin{enumerate}
-  \item Displaying information in a rigid, tabular way, CDCS does not leverage fonts and styling to adhere to typographical standards.
-Instead of using larger or bolder type, for instance, course titles are listed entirely in uppercase (e.g.
-``INTRO TO PROGRAMMING''), which has been shown to be less readable than lower-case text \cite{caps}.
-This problem is compounded when users browse through possibly hundreds of courses.
-Skedge displays properly capitalized titles styled with large type that helps users to quickly group and locate them.
-
-  \item While possibly not a fault of the CDCS system itself, there are very frequently typos or missing spaces in the ``comments'' section of courses, which Skedge corrects.
-
-  \item CDCS displays all course times in 24-hour time, which, despite being concise and unambiguous, is not what most US students are used to.
-Skedge displays 12-hour time with AM/PM, and prevents ambiguities through the course-in-schedule visualization on hover.
-\end{enumerate}
-
-\subsection{Section display}
-
-Often, courses are offered at multiple timeslots, sometimes taught by different instructors and in different rooms.
-These are called \emph{sections} of a course.
-CDCS displays each section in a discrete ``section box'' (all of which are nondistinct and have equal size), even if two sections pertain to the same course.
-(In this regard, CDCS should really be \emph{SDSS}, ``\emph{Section} Description / \emph{Section} Scheduler'', because it operates on the level of sections, not courses.) As a result, course descriptions (which can be lengthy), titles, prerequisites, comments, etc.
-are all repeated for every section of the course.
-
-To make matters worse, many courses in the University course catalog include what I call \emph{subsections}---secondary sections associated with a course that must be registered for separately.
-Namely, these are labs, lab lectures, workshops, and recitations.
-Once again, CDCS displays \emph{all of these} as separate ``section boxes'' by default, and the course description is yet again repeated for each subsection (which, this time can be tens of times), wasting valuable page space.
-
-Collapsing subsections within courses can result in massive improvements in filtering the data most relevant to the user.
-For instance, the search {\tt csc} for Fall 2016 on CDCS results in 147 ``section boxes'', while Skedge only shows 45 ``\emph{course} boxes'', with subsections collapsed within their respective course.
-This triage reduces the data (noise, more correctly) displayed by ~70\%, and is even higher for departments with more abundant labs and workshops, such as Physics (Skedge: 35 vs.
-CDCS: 226, an ~85\% reduction), or Chemistry (Skedge: 25 vs.
-CDCS: 171, an ~86\% reduction).
-
-Skedge can reduce the amount of data to scroll through---and thus the time taken to do so---by six- or seven-fold (and possibly more, counting the attention users otherwise have to pay to distinguish course from subsections), so this design decision has a large usability payoff.
-
-Additionally, some Physics courses (for instance) follow the ``A / B'' subsection structure, where a student registered for an ``A Section'' (as opposed to the ``B Section'') must also register for an ``A Lab'' and ``A Workshop''.
-Skedge organizes subsections for these cases to help sort the two out, which get mixed up in CDCS's linear output.
-
-Note that in Figure \ref{fig:cdcs-sections} (CDCS), the first two boxes are sections for the same course, and the next two are labs for that course.
-Four more lab sections and \emph{twenty} more workshop sessions for that same course follow below the truncated screenshot.
-Figure \ref{fig:sk-sections} (Skedge) demonstrates how this information can be conveyed more concisely.
-
-\begin{figure}[ht]
-  \centering
-  \vspace{10pt}
-  \begin{tabular}{c c}
-    \begin{subfigure}[h]{6cm}
-      \centering
-      \fbox{
-        \includegraphics[width=1.00\textwidth]{images/cdcs/sections}
-      }
-      \caption{Ungrouped sections in CDCS} \label{fig:cdcs-sections}
-    \end{subfigure}
-    \hspace{15pt}
-    \begin{subfigure}[h]{7cm}
-      \centering
-      \fbox{
-        \includegraphics[width=1.00\textwidth]{images/skedge/sections}
-      }
-      \caption{Grouped sections in Skedge} \label{fig:sk-sections}
-    \end{subfigure}
-  \end{tabular}
-  \caption{Section and subsection presentation in CDCS and Skedge}
-\end{figure}
-
-\subsection{Course reference}
-
-\emph{Course mentions} will often appear in the prequisites, crosslists, comments, or description fields of a course (e.g.
-``\textbf{Prerequisites:} CSC 171 or equivalent; MTH 150 is REQUIRED'').
-Users frequently want to find out more information about mentioned courses (frequency shown in Chapter 4).
-In CDCS, because course mentions are displayed as ordinary plaintext, users have to scroll back up, make a search for \emph{that} course, and lose their current search context as a result.
-
-Skedge solves this by linking each course mention to its search query, in the style of Wikipedia.
-Moreover, it prevents context-switches by displaying a popover containing information on that course when the user hovers their cursor over the course mention (see Figure \ref{fig:sk-hover}).
-
-\begin{figure}[ht]
-  \centering
-  \vspace{2pt}
-  \fbox{
-    \includegraphics[width=8cm]{images/skedge/hover}
-  }
-  \caption{Hoverable and clickable course mention in the \emph{Prerequisites} field of a course} \label{fig:sk-hover}
-\end{figure}
-
-\subsection{Multiple schedule support}
-
-With an active schedule still in use, students often want to plan their next semester's schedule.
-This is impossible with the Better CDCS extension, which instead adds courses from both terms to the same schedule, sometimes causing non-existent conflicts.
-A CDCS user must clear their entire schedule out before scheduling a new one.
-On Skedge, however, when a user adds a section of a term for which they do not yet have a schedule, a new one will automatically be created for that term.
-Users are also able to access their old schedules.
-
-Additionally, a common feature request for Skedge is the implementation of multiple schedules \emph{per} term, allowing users to experiment with different schedule possibilities.\footnote{Currently under development, but simple to implement given the existing schedule management structure.}
-
-\section{Search friendliness}
-
-Of course, the most important usability concern for a course explorer/scheduler is being able to effectively \emph{find courses}.
-
-In this section, I will present the \emph{Skedge DSQL}, a domain-specific query language that is based on natural language.
-Next, I will demonstrate how Skedge leverages the DSQL to handle the three search criteria students have for finding courses better than CDCS does.
-
-\subsection{Natural language search}
-
-In the spirit of this chapter's theme, Skedge's search method was, again, primarily designed as a reaction to CDCS's.
-CDCS's search method is a 15-field form (see Figure \ref{fig:cdcs-search}), of which I only found myself regularly using two (one of them being the \emph{required} ``Year/Term'' field).
-This prompted me to closely examine the fields to a) determine redundancies between them and b) find how to make some of the valuable filter fields more usable.
-From this came Skedge's unified, natural language based search method, which I call the \emph{Skedge Domain-Specific Query Language (Skedge DSQL)}.
-Figure \ref{fig:sk-search} is a list of example possible searches shown to users as they type.
-
-\begin{figure}[ht]
-  \centering
-  \vspace{5pt}
-  \begin{tabular}{c c}
-    \begin{subfigure}[w]{4.5cm}
-      \centering
-      \fbox{
-        \includegraphics[width=3.3cm]{images/cdcs/search}
-      }
-      \caption{Form-based search in CDCS} \label{fig:cdcs-search}
-    \end{subfigure}
-    \hspace{5pt}
-    \begin{subfigure}[w]{8.5cm}
-      \centering
-      \fbox{
-        \includegraphics[width=1.00\textwidth]{images/skedge/search}
-      }
-      \caption{Examples of the Skedge DSQL} \label{fig:sk-search}
-    \end{subfigure}
-  \end{tabular}
-  \caption{Two philosophies of search method}
-\end{figure}
-
-\begin{figure}[ht]
-  \centering
-  \vspace{10pt}
-  \includegraphics[width=14cm]{images/search-mapping}
-  \caption{Mapping from CDCS's form-based search to the Skedge DSQL} \label{fig:search-mapping}
-\end{figure}
-
-Figure \ref{fig:search-mapping} is a mapping of each CDCS field to an element of the Skedge DSQL or unified search.
-Dashed lines that end abruptly are fields that were deemed unnecessary or redundant due to a Skedge-unique feature.
-To briefly sum up my rationale for each field:
-
-\onehalfspacing
-
-\subsubsection{Removed fields}
-
-\begin{enumerate}
-
-  \item \emph{School} (e.g.
-``College of Arts, Sciences, \& Engineering'', ``Eastman School of Music'', etc.) was removed in favor of always searching all schools, as is the default on CDCS.
-Searching by department codes (which are unique across schools) designates a school regardless.
-
-  \item \emph{Course Type} (e.g. ``Main Course'', ``Lab'', ``Workshop'' etc.) was removed because Skedge already embeds subsections into courses.
-
-  \item \emph{Status} (i.e.
-``Open'', ``Closed'', or ``Cancelled'') was removed because it is often useful to see closed courses if there is possibility of the instructor letting in more students, and searching for cancelled courses did not seem useful.
-
-  \item \emph{Division} (i.e. ``Humanities'', ``Social Sciences'', etc.) was removed because it seems too broad to be useful, but it could be added to the Skedge DSQL if there is demand.
-
-\end{enumerate}
-
-\subsubsection{Unified Search (left side)}
-
-The \emph{unified search} components can be entered directly into the single Skedge search field, along with the Skedge DSQL. It includes the CDCS fields:
-
-\begin{enumerate}
-  \item \emph{Subject} (department).
-  \item \emph{Course} (department code and/or course number, e.g. ``csc 171'').
-  \item \emph{Title} (course title).
-\end{enumerate}
-
-\subsubsection{Skedge DSQL (right side)}
-
-The \emph{Skedge DSQL} is enterable using special keywords and, optionally, qualifiers for those keywords. It includes the CDCS fields:
-
-\begin{enumerate}
-  \item \emph{Year/term}, which is no longer required.
-It maps to the keywords {\tt \{spring,fall,summer,winter\}} and a four-digit year to specify term and year, and defaults to the current year and term.
-
-  \item \emph{Description}, which is specified by surrounding the query in quotes (e.g. ````buddhism'''').
-
-  \item \emph{Instructor name}, which is specified with the keyword {\tt taught by} and given the instructor name as a qualifier (e.g. ``taught by Guo'').
-
-  \item \emph{Day}, which can be specified with the keywords {\tt \{m,t,w,r,f,s,u\}} for filtering by specific days of the week (e.g. ``mwf'' for Monday-Wednesday-Friday classes only).
-
-  \item \emph{Start time between} is replaced by the keywords {\tt \{early,late\}} which \emph{sort} (ascending and descending, respectively) the results by start time.
-
-  \item \emph{Credit hours between} is replaced by the keyword {\tt credits}, qualified with a number or range (e.g. ``dan 1-2 credits'' for Dance classes with 1-2 credits).
-
-  \item \emph{Upper level writing} is replaced by the single-letter keyword {\tt w} (e.g. ``his w'' for History upper-level writing courses).
-\end{enumerate}
-
-\doublespacing
-
-As noted in Figure \ref{fig:search-mapping}, many use-cases for the \emph{Day} and \emph{Start time between} fields can be obsoleted with an \emph{autofit} feature that will be described later in the next subsection.
-
-  \subsubsection{Advantages of the DSQL}
-
-    The main advantage of Skedge's search system is in its reduction of 15 fields to a single one.
-I hypothesize that since the vast majority of searches fit the department code / course number pattern (this hypothesis later supported in Chapter 4), the time required to force the user to select a term and then find the correct box to enter this query is wasted.
-
-    In addition, once learned and remembered, using a natural language based query language can be faster and more intuitive than clicking on form fields.
-And more importantly, the Skedge DSQL is more easily extendable as new search features are invented.
-For CDCS, the more form fields that are added to the sidebar, the less usable its user interface becomes.
-For Skedge, while it may take longer for users to learn a bigger DSQL, there is no issue with crowding user interface real estate.
-
-  \subsubsection{Disadvantages of the DSQL}
-
-    One disadvantage of Skedge's DSQL is the possible grammatical ambiguity of some queries.
-For instance, the query ``Fall of the Roman Empire'' could be searching for courses with that as the title, or searching for courses in the \emph{fall term} with the title ``of the Roman Empire''.
-Unless multiple queries are run and more complex natural language processing is done, the system would have a hard time disambiguating between the two.
-This could also be solved with a ``did you mean?'' feature, which disambiguates a query into every possible DSQL breakup.
-
-    Another notable disadvantage of the DSQL is having to know it, and it having a possibly steep learning curve for some users.
-However, the argument will be made in Chapter 4 that simply by using the site over time, users will self-learn the Skedge DSQL.
-One of the ways this is possible is thanks to Skedge's search system being \emph{multi-purposed}.
-Hyperlinks around the site---instructor names (which link to ``{\tt taught by <name>}''), course references, or schedule course blocks for example---all direct the user to an ordinary search results page with the search field populated, signalling to the user that such a search is possible.
-
-\subsection{Course selection criteria}
-
-I have identified \emph{three} use-cases for course searching (the existence of and distinction between these cases will be demonstrated with collected usage data in Chapter 4).
-The three cases are \textbf{requirements}, \textbf{electives}, and \textbf{peer recommendations}.
-The Skedge DSL and other application features offer substantial improvements over CDCS for each of these cases.
-
-  \subsubsection{Requirements}
-
-  These are courses that are required for a student to complete their degree, and are typically searched for directly.
-The functionality required here is simple and is mostly satisfied by CDCS, but Skedge offers the following improvements to the process:
-
-  \begin{enumerate}
-    \item \emph{Crosslisted courses:} For students with more than one major and/or minor, searching for courses that are crosslisted between departments can be valuable in reducing their requirement load.
-This search filter is unsupported by CDCS, and is supported by Skedge using the operator ``{\tt x}'' (e.g.
-``{\tt csc x ece}'' for courses listed under both Computer Science and Electrical \& Computer Engineering departments).
-
-    \item \emph{Clusters:} Skedge already stores a users' previously taken courses, so it can intelligently suggest either already-completed clusters or courses that would complete clusters that are missing one or two courses.
-For students with many degree requirements already, this could greatly reduce time spent navigating the University's Cluster Search Engine (for which I also have a long list of grievances, but that lies outside the scope of this paper.)\footnote{This feature is under development and is not currently live.
-It was, incidentally, requested by a Skedge user.}
-
-    \item \emph{CRN:} Surprisingly, search by Course Reference Number is unsupported by CDCS. It is supported by Skedge by just searching the 5-digit number.
-  \end{enumerate}
-
-  \subsubsection{Electives}
-
-  Elective courses can either be courses that are required for a major or minor, but are chosen by the student, or they can be courses of particular interest to students who have the flexibility to take them.
-For this use-case, Skedge offers search, filter, and sorting features that substantially aid users in browsing through courses and discovering ones they might want to take.
-Note that none of the following are supported by CDCS.
-
-  \begin{enumerate}
-
-    \item The Skedge DSQL includes the ``new'' keyword, which qualifies an accompanying search by only displaying the courses that were not offered the previous year for that term.
-For instance, searching ``new csc'' today (Fall 2016) would show Fall Computer Science courses that have been added to the catalog or changed since Fall 2015.
-    \item The Skedge DSQL could include an ``autofit'' keyword, which would take into account the user's current schedule, and only show sections that would not conflict with any already scheduled times.\footnote{This feature is under development and is not currently live.}
-    \item The Skedge DSQL includes the ``random'' keyword, which displays a single random course matching the accompanying query.
-For instance, ``random csc'' would show a random course in Computer Science for the current term.
-Used in conjunction with ``autofit'', it can a powerful way to find interesting new classes that are available for users to try.
-    \item The Skedge DSQL includes sorting by class size---useful for students trying to find smaller, discussion based classes---or by class starting time---for students with preferences of morning or afternoon/evening classes.
-    \item Because instructor names appearing in courses automatically link to searches for more courses taught by that instructor, it is easier and more natural to browse by instructor.
-[NEED STAT HERE]
-
-  \end{enumerate}
-
-  \subsubsection{Peer Recommendations}
-
-  CDCS currently has no support at all for peer course recommendation, a highly undervalued resource for course finding.
-Skedge implements peer recomendations through ``Skedge Social,'' a system detailed in the next section.
-
-%%%
-
-\section{``Skedge Social''}
-
-The question ``what are you taking this semester?'' is certainly the most common smalltalk phrase uttered on campus within the first few weeks of the semester.
-Besides the pure motivation  and it is not unreasonable to assume that students want to take classes that their friends are taking.
-
-\begin{itemize}
-  \item ``What are my friends taking?''
-  \item ``What do my friends recommend?''
-\end{itemize}
-
-\subsection{The issue}
-
-  \subsubsection{Static image vs. live site}
-
-  - Edits don’t update
-  - Referencing courses
-
-  \subsubsection{Finding common courses}
-
-  - requires your friends to share their schedules on FB publicly and you to see their post
-
-%%%%%%%%%%%%%%
-
-  - is schedule-first, not search-first
-  - typically only occurs for the current semester
-
-%figure
-
-\subsection{Skedge's solution}
-
-% Walkthru
-
-  \subsubsection{Friends' course enrollments}
-
-  Mini-feed
-
-  \subsubsection{Friends' course likes}
-
-  \subsubsection{Likes \& enrollments embedded in results}
-
-  \subsubsection{Personal schedule synchronization}
-
-  \subsubsection{Privacy}
-
-  \subsubsection{Notifications}
-
-  %figs of the 2 types

### 2016-04-11 design-social.tex
Split design files

--- 2016-04-11 design.tex
@@
-\chapter{Design in reaction to CDCS}
-
-From its very inception, Skedge's functionality and visual design were driven by the shortcomings of CDCS.
-Skedge is built \emph{bottom-up}, not \emph{top-down}---every aspect of the application was either made as a reaction to a particular grievance in CDCS or as the natural evolution of an existing feature.
-Skedge is thus rooted in \emph{usability} derived from real need, not mere conjecture along the question ``what could students want?''.
-Its success with students, shown in Chapter 4, demonstrates that this usability extends beyond my own standard and can fulfill the various discovered use-cases of students in general.
-
-In this chapter, I invite the reader along on a tour of these grievances and their remedies.
-
-%%%
-
-\section{Modernity}
-
-CDCS is an old system, relatively speaking, and its development on user-facing features has been almost entirely stagnant.
-It launched in 2009, seven years ago, and has hardly changed since.
-Figure \ref{fig:cdcs2010} shows CDCS in July 2010, which, besides the addition of a few search fields, is identical to its current version.
-Yet, since its introduction in 2009, we have seen the rise of mobile devices into ubiquity, a boom in ``hacker culture'' and public APIs, and the capability for standalone web applications to be as sophisticated and dynamic as desktop-class applications without the aid of browser extensions.
-With this in mind, Skedge brings course scheduling to the modern era.
-
-\begin{figure}[ht]
-  \centering
-    \fbox{
-      \includegraphics[width=10cm]{images/cdcs/2010}
-    }
-  \caption{CDCS in July 2, 2010, virtually unchanged from today, courtesy of \emph{Archive.org}}
-  \label{fig:cdcs2010}
-\end{figure}
-
-\subsection{AJAX vs. GET requests}
-
-CDCS makes an AJAX request with every submitted search, meaning that the server receives the request and returns a response all without any page navigation (i.e.
-the URL stays the same and no new page is loaded as the search results are displayed).
-Skedge, however, makes a GET request for every search submission, meaning that the user's browser loads a new page that contains the results and whose URL reflects the search query.
-This simple technical design decision substantially increases usability for two reasons:
-
-\begin{enumerate}
-  \item Page navigations allow users to leverage their browser history as it was designed---after making several searches, CDCS users who use the back button on their browsers will be brought to the page loaded before the very first use of CDCS, possibly losing time spent in crafting sophisticated searches.
-Skedge users can go backwards and forwards through their search histories and scroll locations using native browser functionality.
-  \item Every search query has a unique URL (e.g.
-\url{http://skedgeur.com/?q=csc} for {\tt csc}), so users are able to send links to a specific course or search result to others.
-With CDCS, the URL remains \url{https://cdcs.ur.rochester.edu} throughout the duration of the session.
-\end{enumerate}
-
-\subsection{Mobile}
-
-According to Mary Meeker's 2015 Mobile Technology Trends from Kleiner Perkins Caufield \& Byers\cite{kpcb}, in 2014, 51\% of adult time spent per day on the Internet was from a mobile device, versus 42\% spent on a desktop computer or laptop.
-Time spent on the Internet with mobile devices reached three hours per day in 2014, compared to less than one hour in 2010 (12\% mobile vs.
-75\% desktop/laptop time share).
-
-Undeniably, supporting mobile devices and tablets in web applications is crucial for usability nowadays.
-Note how Skedge is responsive to the user's device in Figure \ref{fig:sk-mobile}, compared with CDCS's lack of mobile support in Figure \ref{fig:cdcs-mobile}.
-CDCS on mobile requires the user to pinch and drag around both to read results and to make new searchs, while Skedge adapts content to the device's screen and fixes the search bar to the top of the screen for easy access while browsing.
-
-Moreover, since no major mobile browser currently supports browser extensions (and if one did, the extensions themselves would most likely need to be re-architected), CDCS on a mobile device loses all scheduling functionality, unlike with Skedge on mobile.
-
-\begin{figure}[ht]
-  \centering
-  \vspace{10pt}
-  \begin{tabular}{c c}
-    \begin{subfigure}[h]{4.5cm}
-      \centering
-      \includegraphics[width=1.00\textwidth]{images/cdcs/mobile}
-      \caption{CDCS} \label{fig:cdcs-mobile}
-    \end{subfigure}
-    \begin{subfigure}[h]{4.4cm}
-      \centering
-      \includegraphics[width=1.00\textwidth]{images/skedge/mobile}
-      \caption{Skedge} \label{fig:sk-mobile}
-    \end{subfigure}
-  \end{tabular}
-  \caption{CDCS and Skedge running on a mobile device}
-\end{figure}
-
-\subsection{Public API}
-
-With the increasing number of attendees at University of Rochester's hackathons, it is clear that the University's ``hacker culture'' is growing---more students are collaborating to build side-projects that integrate resources and servies often benefitting the student community.
-Open-source and open-information services greatly help to foster such innovation, and having public APIs is essential toward this end.
-
-Skedge provides a public JSON API at the root URL \url{http://www.skedgeur.com/api/}, made at the request of a student that was interested in using its course data, and the API has already been used in projects by several other student groups.
-The endpoints included are \url{/api/courses?q=query} (Skedge's query language---described in detail in section 2.3---is supported here), \url{/api/departments?q=optional_query}, and \url{/api/instructors?q=optional_query}.
-
-\subsection{Built-in scheduler}
-
-As explained in the introduction, Skedge offers users a course schedule right in the page, unlike CDCS which requires the Better CDCS browser extension for this functionality.
-Having a schedule native in the application has several advantages:
-
-\begin{enumerate}
-\item Besides some CDCS users possibly not even knowing about Better CDCS, not requiring a browser extension provides for a faster and more seamless user onboarding, especially when building schedules on public computers when extensions can't always be installed.
-
-\item Skedge accommodates a schedule into its design, whereas Better CDCS has to work around an interface that wasn't designed with one.
-As a result, Better CDCS has lower usability, requiring the user to toggle between search results and their schedule.
-Skedge, conversely, gives users immediate visual feedback on how a course would fit into their schedule.
-
-\item Schedule data is centralized on Skedge's servers as opposed to locally in a browser cache, meaning that it can synchronize across a user's devices or in sessions on public computers, persists browser resets, and can be easily publicly shared to other users.
-
-\item Extensions like Better CDCS have limited browser support. Internet Explorer and mobile browsers are unsupported, for example.
-\end{enumerate}
-
-%%%
-
-\section{Usability}
-
-\subsection{Visual presentation}
-
-Skedge offers several improvements over CDCS in the quality of its data presentation:
-
-\begin{enumerate}
-  \item Displaying information in a rigid, tabular way, CDCS does not leverage fonts and styling to adhere to typographical standards.
-Instead of using larger or bolder type, for instance, course titles are listed entirely in uppercase (e.g.
-``INTRO TO PROGRAMMING''), which has been shown to be less readable than lower-case text \cite{caps}.
-This problem is compounded when users browse through possibly hundreds of courses.
-Skedge displays properly capitalized titles styled with large type that helps users to quickly group and locate them.
-
-  \item While possibly not a fault of the CDCS system itself, there are very frequently typos or missing spaces in the ``comments'' section of courses, which Skedge corrects.
-
-  \item CDCS displays all course times in 24-hour time, which, despite being concise and unambiguous, is not what most US students are used to.
-Skedge displays 12-hour time with AM/PM, and prevents ambiguities through the course-in-schedule visualization on hover.
-\end{enumerate}
-
-\subsection{Section display}
-
-Often, courses are offered at multiple timeslots, sometimes taught by different instructors and in different rooms.
-These are called \emph{sections} of a course.
-CDCS displays each section in a discrete ``section box'' (all of which are nondistinct and have equal size), even if two sections pertain to the same course.
-(In this regard, CDCS should really be \emph{SDSS}, ``\emph{Section} Description / \emph{Section} Scheduler'', because it operates on the level of sections, not courses.) As a result, course descriptions (which can be lengthy), titles, prerequisites, comments, etc.
-are all repeated for every section of the course.
-
-To make matters worse, many courses in the University course catalog include what I call \emph{subsections}---secondary sections associated with a course that must be registered for separately.
-Namely, these are labs, lab lectures, workshops, and recitations.
-Once again, CDCS displays \emph{all of these} as separate ``section boxes'' by default, and the course description is yet again repeated for each subsection (which, this time can be tens of times), wasting valuable page space.
-
-Collapsing subsections within courses can result in massive improvements in filtering the data most relevant to the user.
-For instance, the search {\tt csc} for Fall 2016 on CDCS results in 147 ``section boxes'', while Skedge only shows 45 ``\emph{course} boxes'', with subsections collapsed within their respective course.
-This triage reduces the data (noise, more correctly) displayed by ~70\%, and is even higher for departments with more abundant labs and workshops, such as Physics (Skedge: 35 vs.
-CDCS: 226, an ~85\% reduction), or Chemistry (Skedge: 25 vs.
-CDCS: 171, an ~86\% reduction).
-
-Skedge can reduce the amount of data to scroll through---and thus the time taken to do so---by six- or seven-fold (and possibly more, counting the attention users otherwise have to pay to distinguish course from subsections), so this design decision has a large usability payoff.
-
-Additionally, some Physics courses (for instance) follow the ``A / B'' subsection structure, where a student registered for an ``A Section'' (as opposed to the ``B Section'') must also register for an ``A Lab'' and ``A Workshop''.
-Skedge organizes subsections for these cases to help sort the two out, which get mixed up in CDCS's linear output.
-
-Note that in Figure \ref{fig:cdcs-sections} (CDCS), the first two boxes are sections for the same course, and the next two are labs for that course.
-Four more lab sections and \emph{twenty} more workshop sessions for that same course follow below the truncated screenshot.
-Figure \ref{fig:sk-sections} (Skedge) demonstrates how this information can be conveyed more concisely.
-
-\begin{figure}[ht]
-  \centering
-  \vspace{10pt}
-  \begin{tabular}{c c}
-    \begin{subfigure}[h]{6cm}
-      \centering
-      \fbox{
-        \includegraphics[width=1.00\textwidth]{images/cdcs/sections}
-      }
-      \caption{Ungrouped sections in CDCS} \label{fig:cdcs-sections}
-    \end{subfigure}
-    \hspace{15pt}
-    \begin{subfigure}[h]{7cm}
-      \centering
-      \fbox{
-        \includegraphics[width=1.00\textwidth]{images/skedge/sections}
-      }
-      \caption{Grouped sections in Skedge} \label{fig:sk-sections}
-    \end{subfigure}
-  \end{tabular}
-  \caption{Section and subsection presentation in CDCS and Skedge}
-\end{figure}
-
-\subsection{Course reference}
-
-\emph{Course mentions} will often appear in the prequisites, crosslists, comments, or description fields of a course (e.g.
-``\textbf{Prerequisites:} CSC 171 or equivalent; MTH 150 is REQUIRED'').
-Users frequently want to find out more information about mentioned courses (frequency shown in Chapter 4).
-In CDCS, because course mentions are displayed as ordinary plaintext, users have to scroll back up, make a search for \emph{that} course, and lose their current search context as a result.
-
-Skedge solves this by linking each course mention to its search query, in the style of Wikipedia.
-Moreover, it prevents context-switches by displaying a popover containing information on that course when the user hovers their cursor over the course mention (see Figure \ref{fig:sk-hover}).
-
-\begin{figure}[ht]
-  \centering
-  \vspace{2pt}
-  \fbox{
-    \includegraphics[width=8cm]{images/skedge/hover}
-  }
-  \caption{Hoverable and clickable course mention in the \emph{Prerequisites} field of a course} \label{fig:sk-hover}
-\end{figure}
-
-\subsection{Multiple schedule support}
-
-With an active schedule still in use, students often want to plan their next semester's schedule.
-This is impossible with the Better CDCS extension, which instead adds courses from both terms to the same schedule, sometimes causing non-existent conflicts.
-A CDCS user must clear their entire schedule out before scheduling a new one.
-On Skedge, however, when a user adds a section of a term for which they do not yet have a schedule, a new one will automatically be created for that term.
-Users are also able to access their old schedules.
-
-Additionally, a common feature request for Skedge is the implementation of multiple schedules \emph{per} term, allowing users to experiment with different schedule possibilities.\footnote{Currently under development, but simple to implement given the existing schedule management structure.}
-
-\section{Search friendliness}
-
-Of course, the most important usability concern for a course explorer/scheduler is being able to effectively \emph{find courses}.
-
-In this section, I will present the \emph{Skedge DSQL}, a domain-specific query language that is based on natural language.
-Next, I will demonstrate how Skedge leverages the DSQL to handle the three search criteria students have for finding courses better than CDCS does.
-
-\subsection{Natural language search}
-
-In the spirit of this chapter's theme, Skedge's search method was, again, primarily designed as a reaction to CDCS's.
-CDCS's search method is a 15-field form (see Figure \ref{fig:cdcs-search}), of which I only found myself regularly using two (one of them being the \emph{required} ``Year/Term'' field).
-This prompted me to closely examine the fields to a) determine redundancies between them and b) find how to make some of the valuable filter fields more usable.
-From this came Skedge's unified, natural language based search method, which I call the \emph{Skedge Domain-Specific Query Language (Skedge DSQL)}.
-Figure \ref{fig:sk-search} is a list of example possible searches shown to users as they type.
-
-\begin{figure}[ht]
-  \centering
-  \vspace{5pt}
-  \begin{tabular}{c c}
-    \begin{subfigure}[w]{4.5cm}
-      \centering
-      \fbox{
-        \includegraphics[width=3.3cm]{images/cdcs/search}
-      }
-      \caption{Form-based search in CDCS} \label{fig:cdcs-search}
-    \end{subfigure}
-    \hspace{5pt}
-    \begin{subfigure}[w]{8.5cm}
-      \centering
-      \fbox{
-        \includegraphics[width=1.00\textwidth]{images/skedge/search}
-      }
-      \caption{Examples of the Skedge DSQL} \label{fig:sk-search}
-    \end{subfigure}
-  \end{tabular}
-  \caption{Two philosophies of search method}
-\end{figure}
-
-\begin{figure}[ht]
-  \centering
-  \vspace{10pt}
-  \includegraphics[width=14cm]{images/search-mapping}
-  \caption{Mapping from CDCS's form-based search to the Skedge DSQL} \label{fig:search-mapping}
-\end{figure}
-
-Figure \ref{fig:search-mapping} is a mapping of each CDCS field to an element of the Skedge DSQL or unified search.
-Dashed lines that end abruptly are fields that were deemed unnecessary or redundant due to a Skedge-unique feature.
-To briefly sum up my rationale for each field:
-
-\onehalfspacing
-
-\subsubsection{Removed fields}
-
-\begin{enumerate}
-
-  \item \emph{School} (e.g.
-``College of Arts, Sciences, \& Engineering'', ``Eastman School of Music'', etc.) was removed in favor of always searching all schools, as is the default on CDCS.
-Searching by department codes (which are unique across schools) designates a school regardless.
-
-  \item \emph{Course Type} (e.g. ``Main Course'', ``Lab'', ``Workshop'' etc.) was removed because Skedge already embeds subsections into courses.
-
-  \item \emph{Status} (i.e.
-``Open'', ``Closed'', or ``Cancelled'') was removed because it is often useful to see closed courses if there is possibility of the instructor letting in more students, and searching for cancelled courses did not seem useful.
-
-  \item \emph{Division} (i.e. ``Humanities'', ``Social Sciences'', etc.) was removed because it seems too broad to be useful, but it could be added to the Skedge DSQL if there is demand.
-
-\end{enumerate}
-
-\subsubsection{Unified Search (left side)}
-
-The \emph{unified search} components can be entered directly into the single Skedge search field, along with the Skedge DSQL. It includes the CDCS fields:
-
-\begin{enumerate}
-  \item \emph{Subject} (department).
-  \item \emph{Course} (department code and/or course number, e.g. ``csc 171'').
-  \item \emph{Title} (course title).
-\end{enumerate}
-
-\subsubsection{Skedge DSQL (right side)}
-
-The \emph{Skedge DSQL} is enterable using special keywords and, optionally, qualifiers for those keywords. It includes the CDCS fields:
-
-\begin{enumerate}
-  \item \emph{Year/term}, which is no longer required.
-It maps to the keywords {\tt \{spring,fall,summer,winter\}} and a four-digit year to specify term and year, and defaults to the current year and term.
-
-  \item \emph{Description}, which is specified by surrounding the query in quotes (e.g. ````buddhism'''').
-
-  \item \emph{Instructor name}, which is specified with the keyword {\tt taught by} and given the instructor name as a qualifier (e.g. ``taught by Guo'').
-
-  \item \emph{Day}, which can be specified with the keywords {\tt \{m,t,w,r,f,s,u\}} for filtering by specific days of the week (e.g. ``mwf'' for Monday-Wednesday-Friday classes only).
-
-  \item \emph{Start time between} is replaced by the keywords {\tt \{early,late\}} which \emph{sort} (ascending and descending, respectively) the results by start time.
-
-  \item \emph{Credit hours between} is replaced by the keyword {\tt credits}, qualified with a number or range (e.g. ``dan 1-2 credits'' for Dance classes with 1-2 credits).
-
-  \item \emph{Upper level writing} is replaced by the single-letter keyword {\tt w} (e.g. ``his w'' for History upper-level writing courses).
-\end{enumerate}
-
-\doublespacing
-
-As noted in Figure \ref{fig:search-mapping}, many use-cases for the \emph{Day} and \emph{Start time between} fields can be obsoleted with an \emph{autofit} feature that will be described later in the next subsection.
-
-  \subsubsection{Advantages of the DSQL}
-
-    The main advantage of Skedge's search system is in its reduction of 15 fields to a single one.
-I hypothesize that since the vast majority of searches fit the department code / course number pattern (this hypothesis later supported in Chapter 4), the time required to force the user to select a term and then find the correct box to enter this query is wasted.
-
-    In addition, once learned and remembered, using a natural language based query language can be faster and more intuitive than clicking on form fields.
-And more importantly, the Skedge DSQL is more easily extendable as new search features are invented.
-For CDCS, the more form fields that are added to the sidebar, the less usable its user interface becomes.
-For Skedge, while it may take longer for users to learn a bigger DSQL, there is no issue with crowding user interface real estate.
-
-  \subsubsection{Disadvantages of the DSQL}
-
-    One disadvantage of Skedge's DSQL is the possible grammatical ambiguity of some queries.
-For instance, the query ``Fall of the Roman Empire'' could be searching for courses with that as the title, or searching for courses in the \emph{fall term} with the title ``of the Roman Empire''.
-Unless multiple queries are run and more complex natural language processing is done, the system would have a hard time disambiguating between the two.
-This could also be solved with a ``did you mean?'' feature, which disambiguates a query into every possible DSQL breakup.
-
-    Another notable disadvantage of the DSQL is having to know it, and it having a possibly steep learning curve for some users.
-However, the argument will be made in Chapter 4 that simply by using the site over time, users will self-learn the Skedge DSQL.
-One of the ways this is possible is thanks to Skedge's search system being \emph{multi-purposed}.
-Hyperlinks around the site---instructor names (which link to ``{\tt taught by <name>}''), course references, or schedule course blocks for example---all direct the user to an ordinary search results page with the search field populated, signalling to the user that such a search is possible.
-
-\subsection{Course selection criteria}
-
-I have identified \emph{three} use-cases for course searching (the existence of and distinction between these cases will be demonstrated with collected usage data in Chapter 4).
-The three cases are \textbf{requirements}, \textbf{electives}, and \textbf{peer recommendations}.
-The Skedge DSL and other application features offer substantial improvements over CDCS for each of these cases.
-
-  \subsubsection{Requirements}
-
-  These are courses that are required for a student to complete their degree, and are typically searched for directly.
-The functionality required here is simple and is mostly satisfied by CDCS, but Skedge offers the following improvements to the process:
-
-  \begin{enumerate}
-    \item \emph{Crosslisted courses:} For students with more than one major and/or minor, searching for courses that are crosslisted between departments can be valuable in reducing their requirement load.
-This search filter is unsupported by CDCS, and is supported by Skedge using the operator ``{\tt x}'' (e.g.
-``{\tt csc x ece}'' for courses listed under both Computer Science and Electrical \& Computer Engineering departments).
-
-    \item \emph{Clusters:} Skedge already stores a users' previously taken courses, so it can intelligently suggest either already-completed clusters or courses that would complete clusters that are missing one or two courses.
-For students with many degree requirements already, this could greatly reduce time spent navigating the University's Cluster Search Engine (for which I also have a long list of grievances, but that lies outside the scope of this paper.)\footnote{This feature is under development and is not currently live.
-It was, incidentally, requested by a Skedge user.}
-
-    \item \emph{CRN:} Surprisingly, search by Course Reference Number is unsupported by CDCS. It is supported by Skedge by just searching the 5-digit number.
-  \end{enumerate}
-
-  \subsubsection{Electives}
-
-  Elective courses can either be courses that are required for a major or minor, but are chosen by the student, or they can be courses of particular interest to students who have the flexibility to take them.
-For this use-case, Skedge offers search, filter, and sorting features that substantially aid users in browsing through courses and discovering ones they might want to take.
-Note that none of the following features are supported by CDCS.
-
-  \begin{enumerate}
-
-    \item The Skedge DSQL includes the ``new'' keyword, which qualifies an accompanying search by only displaying the courses that were not offered the previous year for that term.
-For instance, searching ``new csc'' today (Fall 2016) would show Fall Computer Science courses that have been added to the catalog or changed since Fall 2015.
-    \item The Skedge DSQL could include an ``autofit'' keyword, which would take into account the user's current schedule, and only show sections that would not conflict with any already scheduled times.\footnote{This feature is under development and is not currently live.}
-    \item The Skedge DSQL includes the ``random'' keyword, which displays a single random course matching the accompanying query.
-For instance, ``random csc'' would show a random course in Computer Science for the current term.
-Used in conjunction with ``autofit'', it can a powerful way to find interesting new classes that are available for users to try.
-    \item The Skedge DSQL includes sorting by class size---useful for students trying to find smaller, discussion based classes---or by class starting time---for students with preferences of morning or afternoon/evening classes.
-    \item Instructor names appearing in courses automatically link to searches for more courses taught by that instructor, which is easier than having to use the CDCS ``instructor'' field to make a separate search.
-
-  \end{enumerate}
-
-  \subsubsection{Peer Recommendations}
-
-  CDCS currently has no support at all for peer course recommendation, a highly undervalued resource for course finding.
-Skedge implements peer recomendations through Skedge Social, a system detailed in the next section.
-
-%%%
-
 \section{Peer connectivity}
 
-The question ``what are you taking this semester?'' may possibly be the most common phrase uttered in campus smalltalk within the first few weeks of a semester (as well as its analogue ``what are you taking \emph{next} semester?'' in the last half of a semester).
-
-However, students' social lives do not solely exist in-person---social media dominates college social lives etc, v impt to consider [CITE].
+The question ``what are you taking this semester?'' may possibly be the most common phrase uttered in campus smalltalk within the first few weeks of a semester.
+But students' social lives do not solely exist in-person---social media dominates college social lives etc, v impt to consider [CITE].
 The ``what are you taking'' question typically translates to social media in the form of the schedule image exported from Better CDCS posted to Facebook (e.g.
@@
       \centering
-      \includegraphics[width=6.5cm]{images/cdcs/social}
+      \includegraphics[width=6cm]{images/cdcs/social}
       \caption{Sharing a static schedule image generated by Better CDCS on Facebook} \label{fig:cdcs-social}
@@
 
-    \begin{subfigure}[w]{7.1cm}
+    \begin{subfigure}[w]{7cm}
       \centering
-      \includegraphics[width=6.6cm]{images/skedge/facebook}
+      \includegraphics[width=6.1cm]{images/skedge/facebook}
       \caption{Sharing a Skedge schedule on Facebook, linked to the live and dynamic schedule share page on Skedge} \label{fig:sk-fb}
@@
 This excludes the possibility of edits to the schedule being reflected in the same post to their friends.
+Moreover, referencing courses (i.e.
+looking up more information on a course a friend is taking) is not very usable under this model as it requires a manual search.
 
-    Moreover, referencing courses (i.e.
-looking up more information on a course a friend is taking) is not very usable under this model, since links cannot be embedded in a static image, and requires a manual search.
-
-    Skedge allows users to generate a public link to a view-only, larger version of their schedules (see Figure \ref{fig:sk-share}).
+    Skedge allows users to generate a public link (of the format \url{http://skedgeur.com/1234}, with the last four digits being a random ID for that schedule) to a view-only, larger version of their schedules (see Figure \ref{fig:sk-share}).
 If a user visiting others' share pages also has their schedule saved on Skedge, the option to ``overlay'' the user's schedule ontop of the share page's schedule is available, making it easier for friends to find common free time (see Figure \ref{fig:sk-overlay}).
@@
       \begin{tabular}{c c}
-        \begin{subfigure}[w]{11cm}
+        \vspace{10pt}
+
+        \begin{subfigure}[w]{10cm}
           \centering
@@
           }
-          \caption{Skedge's schedule ``share'' page, publicly accessible at \url{http://skedgeur.com/1234}, with the last four digits being the randomly generated ID for that schedule.
-Hovering over a course shows information about it, and clicking searches for it.} \label{fig:sk-share}
-        \end{subfigure} \vspace{25pt}\\
+          \caption{Skedge's schedule ``share'' page. Hovering over a course shows information about it, and clicking searches for it.} \label{fig:sk-share}
+        \end{subfigure} \vspace{10pt}\\
 
-        \begin{subfigure}[w]{11cm}
+        \begin{subfigure}[w]{10cm}
           \centering
@@
 
-  This model imposes several obstacles that prevents users from effectively finding the courses their peers are taking. Doing so within this model:
+  The Facebook model stimies users from finding courses peers are taking in several ways. It
 
@@
 
-    \item is exclusively schedule-oriented, not search-oriented.
-Course discovery happens from seeing peers' schedules first, not from \emph{browsing} first and being seeing, for each class, that their peers may be taking it.
+    \item is unorganized, as schedules are spread across space and time (different posts and profiles).
 
-    \item typically only occurs for the \emph{current} semester.
-Under the Facebook model, it can be very difficult, nigh impossible, for students seeking advice, opinions, or even academic help for a specific class to find others who had taken that class before.
+    \item is only effective for the \emph{current semester}. It can be very difficult for students seeking advice (or even academic help) for a specific class to find others who had taken it before.
+
+    \item is \emph{schedule}-oriented, not \emph{search}-oriented. Course discovery can only happen from seeing peers' schedules, not from browsing and finding classes that their peers may be taking.
+
   \end{enumerate}
 
-  Solving these issues would require platform infrastructure external to Facebook, but would still greatly benefit from its ubiquity and centralized role in social media.
-Since Skedge already saves and archives users' schedules, and Facebook has an integration API, building this platform into Skedge proved to be that opportunity.
+  Solving these issues requires platform infrastructure external to Facebook, but Facebook's ubiquity is invaluable.
+Since Skedge already saves and archives users' schedules and Facebook has an integration API, building this platform into Skedge was the natural solution.
 
@@
 
-The \emph{Skedge Social} platform aims to leverage existing social media paradigms (such as the \emph{news feed}, \emph{likes}, and \emph{requests}) to more seamlessly give users answers to the questions alluded to earlier---``What are my friends taking?'' and ``What do my friends recommend?''.
-
-Logging into Skedge Social is very simple---if the user is already logged into Facebook, it just requires one button of authorization (see Figure \ref{fig:sk-social-login}).
+The \emph{Skedge Social} platform aims to leverage existing social media paradigms (e.g.
+\emph{news feed}, \emph{likes}, and \emph{requests}) to more seamlessly give users answers to the questions alluded to earlier---``what are my friends taking?'' and ``what do my friends recommend?''.
+Skedge Social currently only supports Facebook, which can be integrated with a single click (see Figure \ref{fig:sk-social-login}).
 
@@
 
-  Immediately after logging in, the user will see the ``schedule feed,'' the list of the user's peers' current schedules.
-This takes from a familiar paradigm of the \emph{feed} in social networks, which is routinely scrolled for updates.
-For each schedule in the feed, the user can easily visualize their own schedule side-by-side (using with the same ``overlay'' mechanic explained earlier), access the Facebook profile of the schedule's owner, and see the list of that users' course like.
+  Immediately after logging in, the user will see the ``schedule feed,'' the list of the user's peers' current schedules (see Figure \ref{fig:sk-feed}).
+This takes from a familiar paradigm of the \emph{feed} in social networks, which is routinely scrolled for updates.[1] For each schedule, the user can visualize their own schedule side-by-side (using with the same ``overlay'' mechanic explained earlier), access the Facebook profile of the schedule's owner, and see the list of courses that user likes.
+
+  This has the advantage of organizing and persisting peers' schedules, solving the first three issues of the ``Facebook post'' model listed above.
 
@@
       \fbox{
-        \includegraphics[width=10cm]{images/skedge/social}
+        \includegraphics[width=9cm]{images/skedge/social}
       }
       \caption{Skedge Social's schedule feed (the default of four tabs) for a user who has integrated their Skedge account with Facebook.
-The user can toggle between a peer's schedule and the list of their \emph{liked} courses.} \label{fig:sk-social}
+The user can toggle between a peer's schedule and the list of their \emph{liked} courses.} \label{fig:sk-feed}
     \end{figure}
@@
 
-  Skedge Social also integrates into course search results, showing users on the platform their Facebook friends who are taking, have taken, or \emph{liked} each particular course (see Figure \ref{fig:sk-social-course}).
-This has several advantages:
+  Skedge Social also integrates into course search results, showing the user any other peers on the platform who are taking, have taken, or \emph{liked} courses in the results list (see Figure \ref{fig:sk-social-course}).
+This has several advantages, which solve the last ``Facebook post'' model issue listed above:
 
@@
 
-    \begin{figure}[H]
+    \begin{figure}
       \centering
@@
 
-  \subsubsection{Privacy}
+  For the moment, Skedge leverages a user's
+
+  \subsubsection{Privacy and share requests}
+
+  Skedge offers users two privacy options:
+
+  \begin{enumerate}
+  \item ``Share my schedule and likes with all my friends'' (Public)\footnote{Note that even the ``public'' option is not truly public, as it only grants view permissions to people who are already friends with the user on Facebook.}
+  \item ``Share my schedule and likes only with friends I approve'' (Private)
+  \end{enumerate}
 

### 2016-04-11 design.tex
Split design files

--- 2016-04-10 design.tex
@@
 
-\section{Modernity}
-
-CDCS is an old system, relatively speaking, and its development on user-facing features has been almost entirely stagnant.
-It launched in 2009, seven years ago, and has hardly changed since.
-Figure \ref{fig:cdcs2010} shows CDCS in July 2010, which, besides the addition of a few search fields, is identical to its current version.
-Yet, since its introduction in 2009, we have seen the rise of mobile devices into ubiquity, a boom in ``hacker culture'' and public APIs, and the capability for standalone web applications to be as sophisticated and dynamic as desktop-class applications without the aid of browser extensions.
-With this in mind, Skedge brings course scheduling to the modern era.
-
-\begin{figure}[ht]
-  \centering
-    \fbox{
-      \includegraphics[width=10cm]{images/cdcs/2010}
-    }
-  \caption{CDCS in July 2, 2010, virtually unchanged from today, courtesy of \emph{Archive.org}}
-  \label{fig:cdcs2010}
-\end{figure}
-
-\subsection{AJAX vs. GET requests}
-
-CDCS makes an AJAX request with every submitted search, meaning that the server receives the request and returns a response all without any page navigation (i.e.
-the URL stays the same and no new page is loaded as the search results are displayed).
-Skedge, however, makes a GET request for every search submission, meaning that the user's browser loads a new page that contains the results and whose URL reflects the search query.
-This simple technical design decision substantially increases usability for two reasons:
-
-\begin{enumerate}
-  \item Page navigations allow users to leverage their browser history as it was designed---after making several searches, CDCS users who use the back button on their browsers will be brought to the page loaded before the very first use of CDCS, possibly losing time spent in crafting sophisticated searches.
-Skedge users can go backwards and forwards through their search histories and scroll locations using native browser functionality.
-  \item Every search query has a unique URL (e.g.
-\url{http://skedgeur.com/?q=csc} for {\tt csc}), so users are able to send links to a specific course or search result to others.
-With CDCS, the URL remains \url{https://cdcs.ur.rochester.edu} throughout the duration of the session.
-\end{enumerate}
-
-\subsection{Mobile}
-
-According to Mary Meeker's 2015 Mobile Technology Trends from Kleiner Perkins Caufield \& Byers\cite{kpcb}, in 2014, 51\% of adult time spent per day on the Internet was from a mobile device, versus 42\% spent on a desktop computer or laptop.
-Time spent on the Internet with mobile devices reached three hours per day in 2014, compared to less than one hour in 2010 (12\% mobile vs.
-75\% desktop/laptop time share).
-
-Undeniably, supporting mobile devices and tablets in web applications is crucial for usability nowadays.
-Note how Skedge is responsive to the user's device in Figure \ref{fig:sk-mobile}, compared with CDCS's lack of mobile support in Figure \ref{fig:cdcs-mobile}.
-CDCS on mobile requires the user to pinch and drag around both to read results and to make new searchs, while Skedge adapts content to the device's screen and fixes the search bar to the top of the screen for easy access while browsing.
-
-Moreover, since no major mobile browser currently supports browser extensions (and if one did, the extensions themselves would most likely need to be re-architected), CDCS on a mobile device loses all scheduling functionality, unlike with Skedge on mobile.
-
-\begin{figure}[ht]
-  \centering
-  \vspace{10pt}
-  \begin{tabular}{c c}
-    \begin{subfigure}[h]{4.5cm}
-      \centering
-      \includegraphics[width=1.00\textwidth]{images/cdcs/mobile}
-      \caption{CDCS} \label{fig:cdcs-mobile}
-    \end{subfigure}
-    \begin{subfigure}[h]{4.4cm}
-      \centering
-      \includegraphics[width=1.00\textwidth]{images/skedge/mobile}
-      \caption{Skedge} \label{fig:sk-mobile}
-    \end{subfigure}
-  \end{tabular}
-  \caption{CDCS and Skedge running on a mobile device}
-\end{figure}
-
-\subsection{Public API}
-
-With the increasing number of attendees at University of Rochester's hackathons, it is clear that the University's ``hacker culture'' is growing---more students are collaborating to build side-projects that integrate resources and servies often benefitting the student community.
-Open-source and open-information services greatly help to foster such innovation, and having public APIs is essential toward this end.
-
-Skedge provides a public JSON API at the root URL \url{http://www.skedgeur.com/api/}, made at the request of a student that was interested in using its course data, and the API has already been used in projects by several other student groups.
-The endpoints included are \url{/api/courses?q=query} (Skedge's query language---described in detail in section 2.3---is supported here), \url{/api/departments?q=optional_query}, and \url{/api/instructors?q=optional_query}.
-
-\subsection{Built-in scheduler}
-
-As explained in the introduction, Skedge offers users a course schedule right in the page, unlike CDCS which requires the Better CDCS browser extension for this functionality.
-Having a schedule native in the application has several advantages:
-
-\begin{enumerate}
-\item Besides some CDCS users possibly not even knowing about Better CDCS, not requiring a browser extension provides for a faster and more seamless user onboarding, especially when building schedules on public computers when extensions can't always be installed.
-
-\item Skedge accommodates a schedule into its design, whereas Better CDCS has to work around an interface that wasn't designed with one.
-As a result, Better CDCS has lower usability, requiring the user to toggle between search results and their schedule.
-Skedge, conversely, gives users immediate visual feedback on how a course would fit into their schedule.
-
-\item Schedule data is centralized on Skedge's servers as opposed to locally in a browser cache, meaning that it can synchronize across a user's devices or in sessions on public computers, persists browser resets, and can be easily publicly shared to other users.
-
-\item Extensions like Better CDCS have limited browser support. Internet Explorer and mobile browsers are unsupported, for example.
-\end{enumerate}
+\input{design-modernity}
+\clearpage
 
@@
 
-\section{Usability}
-
-\subsection{Visual presentation}
-
-Skedge offers several improvements over CDCS in the quality of its data presentation:
-
-\begin{enumerate}
-  \item Displaying information in a rigid, tabular way, CDCS does not leverage fonts and styling to adhere to typographical standards.
-Instead of using larger or bolder type, for instance, course titles are listed entirely in uppercase (e.g.
-``INTRO TO PROGRAMMING''), which has been shown to be less readable than lower-case text \cite{caps}.
-This problem is compounded when users browse through possibly hundreds of courses.
-Skedge displays properly capitalized titles styled with large type that helps users to quickly group and locate them.
-
-  \item While possibly not a fault of the CDCS system itself, there are very frequently typos or missing spaces in the ``comments'' section of courses, which Skedge corrects.
-
-  \item CDCS displays all course times in 24-hour time, which, despite being concise and unambiguous, is not what most US students are used to.
-Skedge displays 12-hour time with AM/PM, and prevents ambiguities through the course-in-schedule visualization on hover.
-\end{enumerate}
-
-\subsection{Section display}
-
-Often, courses are offered at multiple timeslots, sometimes taught by different instructors and in different rooms.
-These are called \emph{sections} of a course.
-CDCS displays each section in a discrete ``section box'' (all of which are nondistinct and have equal size), even if two sections pertain to the same course.
-(In this regard, CDCS should really be \emph{SDSS}, ``\emph{Section} Description / \emph{Section} Scheduler'', because it operates on the level of sections, not courses.) As a result, course descriptions (which can be lengthy), titles, prerequisites, comments, etc.
-are all repeated for every section of the course.
-
-To make matters worse, many courses in the University course catalog include what I call \emph{subsections}---secondary sections associated with a course that must be registered for separately.
-Namely, these are labs, lab lectures, workshops, and recitations.
-Once again, CDCS displays \emph{all of these} as separate ``section boxes'' by default, and the course description is yet again repeated for each subsection (which, this time can be tens of times), wasting valuable page space.
-
-Collapsing subsections within courses can result in massive improvements in filtering the data most relevant to the user.
-For instance, the search {\tt csc} for Fall 2016 on CDCS results in 147 ``section boxes'', while Skedge only shows 45 ``\emph{course} boxes'', with subsections collapsed within their respective course.
-This triage reduces the data (noise, more correctly) displayed by ~70\%, and is even higher for departments with more abundant labs and workshops, such as Physics (Skedge: 35 vs.
-CDCS: 226, an ~85\% reduction), or Chemistry (Skedge: 25 vs.
-CDCS: 171, an ~86\% reduction).
-
-Skedge can reduce the amount of data to scroll through---and thus the time taken to do so---by six- or seven-fold (and possibly more, counting the attention users otherwise have to pay to distinguish course from subsections), so this design decision has a large usability payoff.
-
-Additionally, some Physics courses (for instance) follow the ``A / B'' subsection structure, where a student registered for an ``A Section'' (as opposed to the ``B Section'') must also register for an ``A Lab'' and ``A Workshop''.
-Skedge organizes subsections for these cases to help sort the two out, which get mixed up in CDCS's linear output.
-
-Note that in Figure \ref{fig:cdcs-sections} (CDCS), the first two boxes are sections for the same course, and the next two are labs for that course.
-Four more lab sections and \emph{twenty} more workshop sessions for that same course follow below the truncated screenshot.
-Figure \ref{fig:sk-sections} (Skedge) demonstrates how this information can be conveyed more concisely.
-
-\begin{figure}[ht]
-  \centering
-  \vspace{10pt}
-  \begin{tabular}{c c}
-    \begin{subfigure}[h]{6cm}
-      \centering
-      \fbox{
-        \includegraphics[width=1.00\textwidth]{images/cdcs/sections}
-      }
-      \caption{Ungrouped sections in CDCS} \label{fig:cdcs-sections}
-    \end{subfigure}
-    \hspace{15pt}
-    \begin{subfigure}[h]{7cm}
-      \centering
-      \fbox{
-        \includegraphics[width=1.00\textwidth]{images/skedge/sections}
-      }
-      \caption{Grouped sections in Skedge} \label{fig:sk-sections}
-    \end{subfigure}
-  \end{tabular}
-  \caption{Section and subsection presentation in CDCS and Skedge}
-\end{figure}
-
-\subsection{Course reference}
-
-\emph{Course mentions} will often appear in the prequisites, crosslists, comments, or description fields of a course (e.g.
-``\textbf{Prerequisites:} CSC 171 or equivalent; MTH 150 is REQUIRED'').
-Users frequently want to find out more information about mentioned courses (frequency shown in Chapter 4).
-In CDCS, because course mentions are displayed as ordinary plaintext, users have to scroll back up, make a search for \emph{that} course, and lose their current search context as a result.
-
-Skedge solves this by linking each course mention to its search query, in the style of Wikipedia.
-Moreover, it prevents context-switches by displaying a popover containing information on that course when the user hovers their cursor over the course mention (see Figure \ref{fig:sk-hover}).
-
-\begin{figure}[ht]
-  \centering
-  \vspace{2pt}
-  \fbox{
-    \includegraphics[width=8cm]{images/skedge/hover}
-  }
-  \caption{Hoverable and clickable course mention in the \emph{Prerequisites} field of a course} \label{fig:sk-hover}
-\end{figure}
-
-\subsection{Multiple schedule support}
-
-With an active schedule still in use, students often want to plan their next semester's schedule.
-This is impossible with the Better CDCS extension, which instead adds courses from both terms to the same schedule, sometimes causing non-existent conflicts.
-A CDCS user must clear their entire schedule out before scheduling a new one.
-On Skedge, however, when a user adds a section of a term for which they do not yet have a schedule, a new one will automatically be created for that term.
-Users are also able to access their old schedules.
-
-Additionally, a common feature request for Skedge is the implementation of multiple schedules \emph{per} term, allowing users to experiment with different schedule possibilities.\footnote{Currently under development, but simple to implement given the existing schedule management structure.}
-
-\section{Search friendliness}
-
-Of course, the most important usability concern for a course explorer/scheduler is being able to effectively \emph{find courses}.
-
-In this section, I will present the \emph{Skedge DSQL}, a domain-specific query language that is based on natural language.
-Next, I will demonstrate how Skedge leverages the DSQL to handle the three search criteria students have for finding courses better than CDCS does.
-
-\subsection{Natural language search}
-
-In the spirit of this chapter's theme, Skedge's search method was, again, primarily designed as a reaction to CDCS's.
-CDCS's search method is a 15-field form (see Figure \ref{fig:cdcs-search}), of which I only found myself regularly using two (one of them being the \emph{required} ``Year/Term'' field).
-This prompted me to closely examine the fields to a) determine redundancies between them and b) find how to make some of the valuable filter fields more usable.
-From this came Skedge's unified, natural language based search method, which I call the \emph{Skedge Domain-Specific Query Language (Skedge DSQL)}.
-Figure \ref{fig:sk-search} is a list of example possible searches shown to users as they type.
-
-\begin{figure}[ht]
-  \centering
-  \vspace{5pt}
-  \begin{tabular}{c c}
-    \begin{subfigure}[w]{4.5cm}
-      \centering
-      \fbox{
-        \includegraphics[width=3.3cm]{images/cdcs/search}
-      }
-      \caption{Form-based search in CDCS} \label{fig:cdcs-search}
-    \end{subfigure}
-    \hspace{5pt}
-    \begin{subfigure}[w]{8.5cm}
-      \centering
-      \fbox{
-        \includegraphics[width=1.00\textwidth]{images/skedge/search}
-      }
-      \caption{Examples of the Skedge DSQL} \label{fig:sk-search}
-    \end{subfigure}
-  \end{tabular}
-  \caption{Two philosophies of search method}
-\end{figure}
-
-\begin{figure}[ht]
-  \centering
-  \vspace{10pt}
-  \includegraphics[width=14cm]{images/search-mapping}
-  \caption{Mapping from CDCS's form-based search to the Skedge DSQL} \label{fig:search-mapping}
-\end{figure}
-
-Figure \ref{fig:search-mapping} is a mapping of each CDCS field to an element of the Skedge DSQL or unified search.
-Dashed lines that end abruptly are fields that were deemed unnecessary or redundant due to a Skedge-unique feature.
-To briefly sum up my rationale for each field:
-
-\onehalfspacing
-
-\subsubsection{Removed fields}
-
-\begin{enumerate}
-
-  \item \emph{School} (e.g.
-``College of Arts, Sciences, \& Engineering'', ``Eastman School of Music'', etc.) was removed in favor of always searching all schools, as is the default on CDCS.
-Searching by department codes (which are unique across schools) designates a school regardless.
-
-  \item \emph{Course Type} (e.g. ``Main Course'', ``Lab'', ``Workshop'' etc.) was removed because Skedge already embeds subsections into courses.
-
-  \item \emph{Status} (i.e.
-``Open'', ``Closed'', or ``Cancelled'') was removed because it is often useful to see closed courses if there is possibility of the instructor letting in more students, and searching for cancelled courses did not seem useful.
-
-  \item \emph{Division} (i.e. ``Humanities'', ``Social Sciences'', etc.) was removed because it seems too broad to be useful, but it could be added to the Skedge DSQL if there is demand.
-
-\end{enumerate}
-
-\subsubsection{Unified Search (left side)}
-
-The \emph{unified search} components can be entered directly into the single Skedge search field, along with the Skedge DSQL. It includes the CDCS fields:
-
-\begin{enumerate}
-  \item \emph{Subject} (department).
-  \item \emph{Course} (department code and/or course number, e.g. ``csc 171'').
-  \item \emph{Title} (course title).
-\end{enumerate}
-
-\subsubsection{Skedge DSQL (right side)}
-
-The \emph{Skedge DSQL} is enterable using special keywords and, optionally, qualifiers for those keywords. It includes the CDCS fields:
-
-\begin{enumerate}
-  \item \emph{Year/term}, which is no longer required.
-It maps to the keywords {\tt \{spring,fall,summer,winter\}} and a four-digit year to specify term and year, and defaults to the current year and term.
-
-  \item \emph{Description}, which is specified by surrounding the query in quotes (e.g. ````buddhism'''').
-
-  \item \emph{Instructor name}, which is specified with the keyword {\tt taught by} and given the instructor name as a qualifier (e.g. ``taught by Guo'').
-
-  \item \emph{Day}, which can be specified with the keywords {\tt \{m,t,w,r,f,s,u\}} for filtering by specific days of the week (e.g. ``mwf'' for Monday-Wednesday-Friday classes only).
-
-  \item \emph{Start time between} is replaced by the keywords {\tt \{early,late\}} which \emph{sort} (ascending and descending, respectively) the results by start time.
-
-  \item \emph{Credit hours between} is replaced by the keyword {\tt credits}, qualified with a number or range (e.g. ``dan 1-2 credits'' for Dance classes with 1-2 credits).
-
-  \item \emph{Upper level writing} is replaced by the single-letter keyword {\tt w} (e.g. ``his w'' for History upper-level writing courses).
-\end{enumerate}
-
-\doublespacing
-
-As noted in Figure \ref{fig:search-mapping}, many use-cases for the \emph{Day} and \emph{Start time between} fields can be obsoleted with an \emph{autofit} feature that will be described later in the next subsection.
-
-  \subsubsection{Advantages of the DSQL}
-
-    The main advantage of Skedge's search system is in its reduction of 15 fields to a single one.
-I hypothesize that since the vast majority of searches fit the department code / course number pattern (this hypothesis later supported in Chapter 4), the time required to force the user to select a term and then find the correct box to enter this query is wasted.
-
-    In addition, once learned and remembered, using a natural language based query language can be faster and more intuitive than clicking on form fields.
-And more importantly, the Skedge DSQL is more easily extendable as new search features are invented.
-For CDCS, the more form fields that are added to the sidebar, the less usable its user interface becomes.
-For Skedge, while it may take longer for users to learn a bigger DSQL, there is no issue with crowding user interface real estate.
-
-  \subsubsection{Disadvantages of the DSQL}
-
-    One disadvantage of Skedge's DSQL is the possible grammatical ambiguity of some queries.
-For instance, the query ``Fall of the Roman Empire'' could be searching for courses with that as the title, or searching for courses in the \emph{fall term} with the title ``of the Roman Empire''.
-Unless multiple queries are run and more complex natural language processing is done, the system would have a hard time disambiguating between the two.
-This could also be solved with a ``did you mean?'' feature, which disambiguates a query into every possible DSQL breakup.
-
-    Another notable disadvantage of the DSQL is having to know it, and it having a possibly steep learning curve for some users.
-However, the argument will be made in Chapter 4 that simply by using the site over time, users will self-learn the Skedge DSQL.
-One of the ways this is possible is thanks to Skedge's search system being \emph{multi-purposed}.
-Hyperlinks around the site---instructor names (which link to ``{\tt taught by <name>}''), course references, or schedule course blocks for example---all direct the user to an ordinary search results page with the search field populated, signalling to the user that such a search is possible.
-
-\subsection{Course selection criteria}
-
-I have identified \emph{three} use-cases for course searching (the existence of and distinction between these cases will be demonstrated with collected usage data in Chapter 4).
-The three cases are \textbf{requirements}, \textbf{electives}, and \textbf{peer recommendations}.
-The Skedge DSL and other application features offer substantial improvements over CDCS for each of these cases.
-
-  \subsubsection{Requirements}
-
-  These are courses that are required for a student to complete their degree, and are typically searched for directly.
-The functionality required here is simple and is mostly satisfied by CDCS, but Skedge offers the following improvements to the process:
-
-  \begin{enumerate}
-    \item \emph{Crosslisted courses:} For students with more than one major and/or minor, searching for courses that are crosslisted between departments can be valuable in reducing their requirement load.
-This search filter is unsupported by CDCS, and is supported by Skedge using the operator ``{\tt x}'' (e.g.
-``{\tt csc x ece}'' for courses listed under both Computer Science and Electrical \& Computer Engineering departments).
-
-    \item \emph{Clusters:} Skedge already stores a users' previously taken courses, so it can intelligently suggest either already-completed clusters or courses that would complete clusters that are missing one or two courses.
-For students with many degree requirements already, this could greatly reduce time spent navigating the University's Cluster Search Engine (for which I also have a long list of grievances, but that lies outside the scope of this paper.)\footnote{This feature is under development and is not currently live.
-It was, incidentally, requested by a Skedge user.}
-
-    \item \emph{CRN:} Surprisingly, search by Course Reference Number is unsupported by CDCS. It is supported by Skedge by just searching the 5-digit number.
-  \end{enumerate}
-
-  \subsubsection{Electives}
-
-  Elective courses can either be courses that are required for a major or minor, but are chosen by the student, or they can be courses of particular interest to students who have the flexibility to take them.
-For this use-case, Skedge offers search, filter, and sorting features that substantially aid users in browsing through courses and discovering ones they might want to take.
-Note that none of the following are supported by CDCS.
-
-  \begin{enumerate}
-
-    \item The Skedge DSQL includes the ``new'' keyword, which qualifies an accompanying search by only displaying the courses that were not offered the previous year for that term.
-For instance, searching ``new csc'' today (Fall 2016) would show Fall Computer Science courses that have been added to the catalog or changed since Fall 2015.
-    \item The Skedge DSQL could include an ``autofit'' keyword, which would take into account the user's current schedule, and only show sections that would not conflict with any already scheduled times.\footnote{This feature is under development and is not currently live.}
-    \item The Skedge DSQL includes the ``random'' keyword, which displays a single random course matching the accompanying query.
-For instance, ``random csc'' would show a random course in Computer Science for the current term.
-Used in conjunction with ``autofit'', it can a powerful way to find interesting new classes that are available for users to try.
-    \item The Skedge DSQL includes sorting by class size---useful for students trying to find smaller, discussion based classes---or by class starting time---for students with preferences of morning or afternoon/evening classes.
-    \item Because instructor names appearing in courses automatically link to searches for more courses taught by that instructor, it is easier and more natural to browse by instructor.
-[NEED STAT HERE]
-
-  \end{enumerate}
-
-  \subsubsection{Peer Recommendations}
-
-  CDCS currently has no support at all for peer course recommendation, a highly undervalued resource for course finding.
-Skedge implements peer recomendations through ``Skedge Social,'' a system detailed in the next section.
+\input{design-usability}
+\clearpage
 
@@
 
-\section{``Skedge Social''}
+\input{design-search}
+\clearpage
 
-The question ``what are you taking this semester?'' is certainly the most common smalltalk phrase uttered on campus within the first few weeks of the semester.
-Besides the pure motivation  and it is not unreasonable to assume that students want to take classes that their friends are taking.
+%%%
 
-\begin{itemize}
-  \item ``What are my friends taking?''
-  \item ``What do my friends recommend?''
-\end{itemize}
-
-\subsection{The issue}
-
-  \subsubsection{Static image vs. live site}
-
-  - Edits don’t update
-  - Referencing courses
-
-  \subsubsection{Finding common courses}
-
-  - requires your friends to share their schedules on FB publicly and you to see their post
-
-%%%%%%%%%%%%%%
-
-  - is schedule-first, not search-first
-  - typically only occurs for the current semester
-
-%figure
-
-\subsection{Skedge's solution}
-
-% Walkthru
-
-  \subsubsection{Friends' course enrollments}
-
-  Mini-feed
-
-  \subsubsection{Friends' course likes}
-
-  \subsubsection{Likes \& enrollments embedded in results}
-
-  \subsubsection{Personal schedule synchronization}
-
-  \subsubsection{Privacy}
-
-  \subsubsection{Notifications}
-
-  %figs of the 2 types
+\input{design-social}

### 2016-04-11 design/search.tex
More splitup

\section{Search friendliness}

Of course, the most important usability concern for a course explorer/scheduler is being able to effectively \emph{find courses}.

In this section, I will present the \emph{Skedge DSQL}, a domain-specific query language that is based on natural language. Next, I will demonstrate how Skedge leverages the DSQL to handle the three search criteria students have for finding courses better than CDCS does.

\subsection{Natural language search}

In the spirit of this chapter's theme, Skedge's search method was, again, primarily designed as a reaction to CDCS's. CDCS's search method is a 15-field form (see Figure \ref{fig:cdcs-search}), of which I only found myself regularly using two (one of them being the \emph{required} ``Year/Term'' field). This prompted me to closely examine the fields to a) determine redundancies between them and b) find how to make some of the valuable filter fields more usable. From this came Skedge's unified, natural language based search method, which I call the \emph{Skedge Domain-Specific Query Language (Skedge DSQL)}. Figure \ref{fig:sk-search} is a list of example possible searches shown to users as they type.

\begin{figure}[ht]
  \centering
  \vspace{5pt}
  \begin{tabular}{c c}
    \begin{subfigure}[w]{4.5cm}
      \centering
      \fbox{
        \includegraphics[width=3.3cm]{images/cdcs/search}
      }
      \caption{Form-based search in CDCS} \label{fig:cdcs-search}
    \end{subfigure}
    \hspace{5pt}
    \begin{subfigure}[w]{8.5cm}
      \centering
      \fbox{
        \includegraphics[width=1.00\textwidth]{images/skedge/search}
      }
      \caption{Examples of the Skedge DSQL} \label{fig:sk-search}
    \end{subfigure}
  \end{tabular}
  \caption{Two philosophies of search method}
\end{figure}

\begin{figure}[ht]
  \centering
  \vspace{10pt}
  \includegraphics[width=14cm]{images/search-mapping}
  \caption{Mapping from CDCS's form-based search to the Skedge DSQL} \label{fig:search-mapping}
\end{figure}

Figure \ref{fig:search-mapping} is a mapping of each CDCS field to an element of the Skedge DSQL or unified search. Dashed lines that end abruptly are fields that were deemed unnecessary or redundant due to a Skedge-unique feature. To briefly sum up my rationale for each field:

\onehalfspacing

\subsubsection{Removed fields}

\begin{enumerate}

  \item \emph{School} (e.g. ``College of Arts, Sciences, \& Engineering'', ``Eastman School of Music'', etc.) was removed in favor of always searching all schools, as is the default on CDCS. Searching by department codes (which are unique across schools) designates a school regardless.

  \item \emph{Course Type} (e.g. ``Main Course'', ``Lab'', ``Workshop'' etc.) was removed because Skedge already embeds subsections into courses.

  \item \emph{Status} (i.e. ``Open'', ``Closed'', or ``Cancelled'') was removed because it is often useful to see closed courses if there is possibility of the instructor letting in more students, and searching for cancelled courses did not seem useful.

  \item \emph{Division} (i.e. ``Humanities'', ``Social Sciences'', etc.) was removed because it seems too broad to be useful, but it could be added to the Skedge DSQL if there is demand.

\end{enumerate}

\subsubsection{Unified Search (left side)}

The \emph{unified search} components can be entered directly into the single Skedge search field, along with the Skedge DSQL. It includes the CDCS fields:

\begin{enumerate}
  \item \emph{Subject} (department).
  \item \emph{Course} (department code and/or course number, e.g. ``csc 171'').
  \item \emph{Title} (course title).
\end{enumerate}

\subsubsection{Skedge DSQL (right side)}

The \emph{Skedge DSQL} is enterable using special keywords and, optionally, qualifiers for those keywords. It includes the CDCS fields:

\begin{enumerate}
  \item \emph{Year/term}, which is no longer required. It maps to the keywords {\tt \{spring,fall,summer,winter\}} and a four-digit year to specify term and year, and defaults to the current year and term.

  \item \emph{Description}, which is specified by surrounding the query in quotes (e.g. ````buddhism'''').

  \item \emph{Instructor name}, which is specified with the keyword {\tt taught by} and given the instructor name as a qualifier (e.g. ``taught by Guo'').

  \item \emph{Day}, which can be specified with the keywords {\tt \{m,t,w,r,f,s,u\}} for filtering by specific days of the week (e.g. ``mwf'' for Monday-Wednesday-Friday classes only).

  \item \emph{Start time between} is replaced by the keywords {\tt \{early,late\}} which \emph{sort} (ascending and descending, respectively) the results by start time.

  \item \emph{Credit hours between} is replaced by the keyword {\tt credits}, qualified with a number or range (e.g. ``dan 1-2 credits'' for Dance classes with 1-2 credits).

  \item \emph{Upper level writing} is replaced by the single-letter keyword {\tt w} (e.g. ``his w'' for History upper-level writing courses).
\end{enumerate}

\doublespacing

As noted in Figure \ref{fig:search-mapping}, many use-cases for the \emph{Day} and \emph{Start time between} fields can be obsoleted with an \emph{autofit} feature that will be described later in the next subsection.

  \subsubsection{Advantages of the DSQL}

    The main advantage of Skedge's search system is in its reduction of 15 fields to a single one. I hypothesize that since the vast majority of searches fit the department code / course number pattern (this hypothesis later supported in Chapter 4), the time required to force the user to select a term and then find the correct box to enter this query is wasted.

    In addition, once learned and remembered, using a natural language based query language can be faster and more intuitive than clicking on form fields. And more importantly, the Skedge DSQL is more easily extendable as new search features are invented. For CDCS, the more form fields that are added to the sidebar, the less usable its user interface becomes. For Skedge, while it may take longer for users to learn a bigger DSQL, there is no issue with crowding user interface real estate.

  \subsubsection{Disadvantages of the DSQL}

    One disadvantage of Skedge's DSQL is the possible grammatical ambiguity of some queries. For instance, the query ``Fall of the Roman Empire'' could be searching for courses with that as the title, or searching for courses in the \emph{fall term} with the title ``of the Roman Empire''. Unless multiple queries are run and more complex natural language processing is done, the system would have a hard time disambiguating between the two. This could also be solved with a ``did you mean?'' feature, which disambiguates a query into every possible DSQL breakup.

    Another notable disadvantage of the DSQL is having to know it, and it having a possibly steep learning curve for some users. However, the argument will be made in Chapter 4 that simply by using the site over time, users will self-learn the Skedge DSQL. One of the ways this is possible is thanks to Skedge's search system being \emph{multi-purposed}. Hyperlinks around the site---instructor names (which link to ``{\tt taught by <name>}''), course references, or schedule course blocks for example---all direct the user to an ordinary search results page with the search field populated, signalling to the user that such a search is possible.

\subsection{Course selection criteria}

I have identified \emph{three} use-cases for course searching (the existence of and distinction between these cases will be demonstrated with collected usage data in Chapter 4). The three cases are \textbf{requirements}, \textbf{electives}, and \textbf{peer recommendations}. The Skedge DSL and other application features offer substantial improvements over CDCS for each of these cases.

  \subsubsection{Requirements}

  These are courses that are required for a student to complete their degree, and are typically searched for directly. The functionality required here is simple and is mostly satisfied by CDCS, but Skedge offers the following improvements to the process:

  \begin{enumerate}
    \item \emph{Crosslisted courses:} For students with more than one major and/or minor, searching for courses that are crosslisted between departments can be valuable in reducing their requirement load. This search filter is unsupported by CDCS, and is supported by Skedge using the operator ``{\tt x}'' (e.g. ``{\tt csc x ece}'' for courses listed under both Computer Science and Electrical \& Computer Engineering departments).

    \item \emph{Clusters:} Skedge already stores a users' previously taken courses, so it can intelligently suggest either already-completed clusters or courses that would complete clusters that are missing one or two courses. For students with many degree requirements already, this could greatly reduce time spent navigating the University's Cluster Search Engine (for which I also have a long list of grievances, but that lies outside the scope of this paper.)\footnote{This feature is under development and is not currently live. It was, incidentally, requested by a Skedge user.}

    \item \emph{CRN:} Surprisingly, search by Course Reference Number is unsupported by CDCS. It is supported by Skedge by just searching the 5-digit number.
  \end{enumerate}

  \subsubsection{Electives}

  Elective courses can either be courses that are required for a major or minor, but are chosen by the student, or they can be courses of particular interest to students who have the flexibility to take them. For this use-case, Skedge offers search, filter, and sorting features that substantially aid users in browsing through courses and discovering ones they might want to take. Note that none of the following features are supported by CDCS.

  \begin{enumerate}

    \item The Skedge DSQL includes the ``new'' keyword, which qualifies an accompanying search by only displaying the courses that were not offered the previous year for that term. For instance, searching ``new csc'' today (Fall 2016) would show Fall Computer Science courses that have been added to the catalog or changed since Fall 2015.
    \item The Skedge DSQL could include an ``autofit'' keyword, which would take into account the user's current schedule, and only show sections that would not conflict with any already scheduled times.\footnote{This feature is under development and is not currently live.}
    \item The Skedge DSQL includes the ``random'' keyword, which displays a single random course matching the accompanying query. For instance, ``random csc'' would show a random course in Computer Science for the current term. Used in conjunction with ``autofit'', it can a powerful way to find interesting new classes that are available for users to try.
    \item The Skedge DSQL includes sorting by class size---useful for students trying to find smaller, discussion based classes---or by class starting time---for students with preferences of morning or afternoon/evening classes.
    \item Instructor names appearing in courses automatically link to searches for more courses taught by that instructor, which is easier than having to use the CDCS ``instructor'' field to make a separate search.

  \end{enumerate}

  \subsubsection{Peer Recommendations}

  CDCS currently has no support at all for peer course recommendation, a highly undervalued resource for course finding. Skedge implements peer recomendations through Skedge Social, a system detailed in the next section.

### 2016-04-11 design/usability.tex
More splitup

\section{Usability}

\subsection{Visual presentation}

Skedge offers several improvements over CDCS in the quality of its data presentation:

\begin{enumerate}
  \item Displaying information in a rigid, tabular way, CDCS does not leverage fonts and styling to adhere to typographical standards. Instead of using larger or bolder type, for instance, course titles are listed entirely in uppercase (e.g. ``INTRO TO PROGRAMMING''), which has been shown to be less readable than lower-case text \cite{caps}. This problem is compounded when users browse through possibly hundreds of courses. Skedge displays properly capitalized titles styled with large type that helps users to quickly group and locate them.

  \item While possibly not a fault of the CDCS system itself, there are very frequently typos or missing spaces in the ``comments'' section of courses, which Skedge corrects.

  \item CDCS displays all course times in 24-hour time, which, despite being concise and unambiguous, is not what most US students are used to. Skedge displays 12-hour time with AM/PM, and prevents ambiguities through the course-in-schedule visualization on hover.
\end{enumerate}

\subsection{Section display}

Often, courses are offered at multiple timeslots, sometimes taught by different instructors and in different rooms. These are called \emph{sections} of a course. CDCS displays each section in a discrete ``section box'' (all of which are nondistinct and have equal size), even if two sections pertain to the same course. (In this regard, CDCS should really be \emph{SDSS}, ``\emph{Section} Description / \emph{Section} Scheduler'', because it operates on the level of sections, not courses.) As a result, course descriptions (which can be lengthy), titles, prerequisites, comments, etc. are all repeated for every section of the course.

To make matters worse, many courses in the University course catalog include what I call \emph{subsections}---secondary sections associated with a course that must be registered for separately. Namely, these are labs, lab lectures, workshops, and recitations. Once again, CDCS displays \emph{all of these} as separate ``section boxes'' by default, and the course description is yet again repeated for each subsection (which, this time can be tens of times), wasting valuable page space.

Collapsing subsections within courses can result in massive improvements in filtering the data most relevant to the user. For instance, the search {\tt csc} for Fall 2016 on CDCS results in 147 ``section boxes'', while Skedge only shows 45 ``\emph{course} boxes'', with subsections collapsed within their respective course. This triage reduces the data (noise, more correctly) displayed by ~70\%, and is even higher for departments with more abundant labs and workshops, such as Physics (Skedge: 35 vs. CDCS: 226, an ~85\% reduction), or Chemistry (Skedge: 25 vs. CDCS: 171, an ~86\% reduction).

Skedge can reduce the amount of data to scroll through---and thus the time taken to do so---by six- or seven-fold (and possibly more, counting the attention users otherwise have to pay to distinguish course from subsections), so this design decision has a large usability payoff.

Additionally, some Physics courses (for instance) follow the ``A / B'' subsection structure, where a student registered for an ``A Section'' (as opposed to the ``B Section'') must also register for an ``A Lab'' and ``A Workshop''. Skedge organizes subsections for these cases to help sort the two out, which get mixed up in CDCS's linear output.

Note that in Figure \ref{fig:cdcs-sections} (CDCS), the first two boxes are sections for the same course, and the next two are labs for that course. Four more lab sections and \emph{twenty} more workshop sessions for that same course follow below the truncated screenshot. Figure \ref{fig:sk-sections} (Skedge) demonstrates how this information can be conveyed more concisely.

\begin{figure}[ht]
  \centering
  \vspace{10pt}
  \begin{tabular}{c c}
    \begin{subfigure}[h]{6cm}
      \centering
      \fbox{
        \includegraphics[width=1.00\textwidth]{images/cdcs/sections}
      }
      \caption{Ungrouped sections in CDCS} \label{fig:cdcs-sections}
    \end{subfigure}
    \hspace{15pt}
    \begin{subfigure}[h]{7cm}
      \centering
      \fbox{
        \includegraphics[width=1.00\textwidth]{images/skedge/sections}
      }
      \caption{Grouped sections in Skedge} \label{fig:sk-sections}
    \end{subfigure}
  \end{tabular}
  \caption{Section and subsection presentation in CDCS and Skedge}
\end{figure}

\subsection{Course reference}

\emph{Course mentions} will often appear in the prequisites, crosslists, comments, or description fields of a course (e.g. ``\textbf{Prerequisites:} CSC 171 or equivalent; MTH 150 is REQUIRED''). Users frequently want to find out more information about mentioned courses (frequency shown in Chapter 4). In CDCS, because course mentions are displayed as ordinary plaintext, users have to scroll back up, make a search for \emph{that} course, and lose their current search context as a result.

Skedge solves this by linking each course mention to its search query, in the style of Wikipedia. Moreover, it prevents context-switches by displaying a popover containing information on that course when the user hovers their cursor over the course mention (see Figure \ref{fig:sk-hover}).

\begin{figure}[ht]
  \centering
  \vspace{2pt}
  \fbox{
    \includegraphics[width=8cm]{images/skedge/hover}
  }
  \caption{Hoverable and clickable course mention in the \emph{Prerequisites} field of a course} \label{fig:sk-hover}
\end{figure}

\subsection{Multiple schedule support}

With an active schedule still in use, students often want to plan their next semester's schedule. This is impossible with the Better CDCS extension, which instead adds courses from both terms to the same schedule, sometimes causing non-existent conflicts. A CDCS user must clear their entire schedule out before scheduling a new one. On Skedge, however, when a user adds a section of a term for which they do not yet have a schedule, a new one will automatically be created for that term. Users are also able to access their old schedules.

Additionally, a common feature request for Skedge is the implementation of multiple schedules \emph{per} term, allowing users to experiment with different schedule possibilities.\footnote{Currently under development, but simple to implement given the existing schedule management structure.}

### 2016-04-11 design/search.tex
social is done

@@
 
-As noted in Figure \ref{fig:search-mapping}, many use-cases for the \emph{Day} and \emph{Start time between} fields can be obsoleted with an \emph{autofit} feature that will be described later in the next subsection.
+\noindent As noted in Figure \ref{fig:search-mapping}, many use-cases for the \emph{Day} and \emph{Start time between} fields can be obsoleted with an \emph{autofit} feature that will be described later in the next subsection.
 

### 2016-04-11 design/social.tex
social is done

--- 2016-04-11 design-social.tex
@@
 
-  Solving these issues requires platform infrastructure external to Facebook, but Facebook's ubiquity is invaluable.
+  \noindent Solving these issues requires platform infrastructure external to Facebook, but Facebook's ubiquity is invaluable.
 Since Skedge already saves and archives users' schedules and Facebook has an integration API, building this platform into Skedge was the natural solution.
@@
 
-  For the moment, Skedge leverages a user's
+  In Section 2.1.4, discussing the disadvantages of Better CDCS's decentralized schedule store, I mentioned the possibility of synchronizing a user's schedules across their devices.
+Skedge leverages a user's Facebook integration to provide this synchronization.
+Upon signing into Skedge with Facebook from another device, that device's browser will automatically update with all of the user's data.
+Modification from any device will propogate changes to the rest.
+
+  For this purpose, implementation of a simple login system (either in-house or using NetID, for instance) could easily replace the reliance on Facebook integration.
+This is currently unsupported, but is a space to explore in the near future.
 
@@
 
+  \noindent To implement privacy option \#2, Skedge Social includes a basic share request/approval system.
+The main interface has tabs \emph{Share requests} and \emph{Non-sharing friends} (visible in Figure \ref{fig:sk-feed}) for displaying pending share requests for approval, and for sending share requests to the user's Facebook friends on Skedge that have the private setting enabled, respectively.
+
   \subsubsection{Notifications}
 
-  \begin{figure}[H]
+  Skedge Social provides notifications for any pending share requests described above (Figure \ref{fig:sk-social-note1}), as well as for any Facebook friends that have joined the platform since the user's last visit to Skedge Social (Figure \ref{fig:sk-social-note2}).
+
+  CITE THAT NOTIFICATION BADGES ARE PAVLOVIAN
+
+  \begin{figure}
     \centering
     \begin{tabular}{c c}
-      \begin{subfigure}[w]{3cm}
+      \begin{subfigure}[w]{5.5cm}
         \centering
         \includegraphics[height=1.5cm]{images/skedge/social-note1}
-        \caption{Note1.} \label{fig:sk-social-note1}
+        \caption{\emph{Pending share requests} notification} \label{fig:sk-social-note1}
       \end{subfigure}
 
-      \begin{subfigure}[w]{3cm}
+      \hspace{20pt}
+
+      \begin{subfigure}[w]{6.5cm}
         \centering
         \includegraphics[height=1.5cm]{images/skedge/social-note2}
-        \caption{Note2.} \label{fig:sk-social-note2}
+        \caption{\emph{New friends} notification} \label{fig:sk-social-note2}
       \end{subfigure}
     \end{tabular}
-    \caption{Notifications}
+    \caption{Notifications: none, one, or both of them can adorn the Skedge Social logo}
   \end{figure}

### 2016-04-11 paper.tex
work on social... so slow
More splitup
Small fixes

--- 2016-04-10 paper.tex
@@
 \title{
-\vspace{60pt}\\
+\vspace{50pt}\\
 \huge \bfseries SKEDGE
@@
 \date{April 12, 2016\\
-    \vspace{150pt}\\
+    \vspace{100pt}\\
     submitted in partial fulfillment of\\
@@
     \emph{honors bachelor of science}\\
+    \vspace{30pt}\\
+    and as an open letter addressed to\\
+    the information technology staff\\
+    at the university of rochester
+    \vspace{-30pt}
 }
@@
 In this paper I present Skedge, a web application for students to comfortably and effectively engage with the University's course catalog.
-Skedge matches and surpasses the capabilities of the existing University tool for this purpose, ``Course Description / Course Schedule'' (CDCS) and presents its information in a more visually pleasing way.
+Skedge matches and surpasses the capabilities of the existing University tool for this purpose, ``Course Description / Course Schedule'' (CDCS) and presents its information in a more visually appealing way.
 As a result, Skedge boasts strong user-retention rates, long session durations, and high student adoption despite having virtually no advertisement.
-Through collected usage data, I demonstrate that a) Skedge's differences from and additions to CDCS are usable and have real need, b) the two major use-cases associated with course browsing---direct search and exploratory search---are effectively accommodated by Skedge, and c) Skedge's search mechanism is user-friendly and self-teaches to users over time.
+Through collected usage data, I demonstrate that \textbf{a)} Skedge's differences from and additions to CDCS are usable and have real-world need, \textbf{b)} the three major use-cases associated with course browsing---direct search, exploratory search, and peer recommendation---are effectively accommodated by Skedge, and \textbf{c)} Skedge's novel search mechanism is user-friendly and self-teaches to users over time.
 
@@
 
-\include{future}
-\clearpage
-
 \include{conclusion}
@@
 
-	\bibitem{kpcb} Takis Konstantopoulos {\em Introductory lecture notes on
-		Markov Chains and Random Walks.} Uppsala University,
-		\\\url{http://www2.math.uu.se/~takis/L/McRw/mcrw.pdf}
+	\bibitem{kpcb} Meeker, Mary (2015). http://www.slideshare.net/kleinerperkins/internet-trends-v1
 

### 2016-04-12 analytics.tex
Remove unneeded files
Analytics work

--- 2016-04-09 03-analytics.tex
@@
-\chapter{Data Analytics}
+\chapter{Data analytics}
 
-Hypotheses:
+In this chapter I propose three hypotheses. They are:
 
-1. Skedge's differences from and additions to CDCS are usable and have real need
+\begin{enumerate}
+  \item Skedge's differences from and additions to CDCS are \textbf{usable and have real need}.
 
-2. Skedge’s navigations-per-add and other metrics demonstrate effectiveness of the use cases
-a) direct searching, and
-b) course browsing
+  \item Skedge’s \emph{navigations-per-add} demonstrate the \textbf{effectiveness of its search features} for each of the following three user search paradigms:
 
-3. Skedge’s DSL is user-friendly; users learn more advanced search types over time by using it
+  \begin{enumerate}
+    \item direct search,
+    \item exploratory search, and
+    \item peer-guided search
+  \end{enumerate}
+
+  \item Skedge’s \textbf{DSQL is user-friendly}; users learn more advanced search types over time \textbf{simply by using it}.
+\end{enumerate}
+
+\noindent In the following sections I will present my analysis on collected usage data since November 3rd 2015, which will serve to validate each hypothesis.
 
@@
 
-Since November 3rd 2015 (137 days)
-3,768 unique users
-4,500 schedules
-Average 90 sessions/day
-Average 4.92 pages/session
-Average 5:31 minutes/session
-28\% of sessions are from new users
+From November 3, 2015 to April 11, 2016 (a period of 159 days), Skedge has seen\ldots
 
-MOBILE RESULT
+\begin{itemize}
+  \item 4,713 users that have added or bookmarked at least 1 course
+  \item 5,256 schedules
+  \item 17,012 sessions\footnote{A session is defined by a period of activity expiring after 30 minutes of inactivity}, 75\% of which are from returning visitors
+  \item 107 sessions per day, on average
+  \item 5.40 pages viewed per session, on average 
+  \item 6 minutes spent per session, on average
+\end{itemize}
+
+\begin{figure}
+  \centering
+
+  \includegraphics[width=1.0\textwidth]{images/graph/searches}
+
+  \caption{Etc}
+  \label{fig:searches}
+
+\end{figure}
+
+periodic visit cycle graph means people are actually using it and decided to stay with it, not declining, can be seen w exports & link-searches.
+
+also should be noted that NO seniors participated in the last, the majority of the userbase!
+
+success Considering skedge is OPTIONAL.
+
+\subsection{Exports}
+
+\begin{figure}
+  \centering
+  \includegraphics[width=1.0\textwidth]{images/graph/exports}
+
+  \caption{Etc}
+  \label{fig:searchtypes}
+\end{figure}
+
+Success! (broken on cdcs)
+
+\subsection{``Link-searches''}
+
+\begin{figure}
+  \centering
+  \includegraphics[width=1.0\textwidth]{images/graph/linksearches}
+
+  \caption{Etc}
+  \label{fig:searchtypes}
+\end{figure}
+
+Success, used!
+
+\subsection{Mobile}
+
+\begin{itemize}
+  \item 11.54\% of sessions
+  \item 2.74 page viewed per session, on average
+  \item 2 minutes spent per session, on average
+\end{itemize}
+
+success bc mobile doesn't exist on cdcs
+
+\subsection{Course Blocks}
+
+\begin{figure}
+  \centering
+
+  \fbox {
+    \includegraphics[width=8cm]{images/skedge/blocks}
+  }
+
+  \caption{Etc}
+  \label{fig:searchtypes}
+
+\end{figure}
+
+40\% of sessions have at least one block-click
+Average of 5.12 block-clicks per session
+
+success because you can't click schedule blocks in better cdcs!!
 
@@
 
-% figs
+  \subsubsection{Search types}
+
+  \begin{figure}
+    \centering
+    \includegraphics[width=1.0\textwidth]{images/graph/searchtypes}
+
+    \caption{Etc}
+    \label{fig:searchtypes}
+
+  \end{figure}
 
@@
 
-\subsection{Course blocks}
+\section{Effectiveness of search}
 
-40\% of sessions have at least one block-click
-Average of 4.94 block-clicks per session
+\subsection{Definitions}
 
-\subsection{Social}
+A navigation is defined as
+a search, or
+a click on an instructor’s name, or
+a click on a crosslisted or prerequisite course link
+
+The navigations-per-add measure is
+the number of navigations a user took (within one session) until a course was added, bookmarked
+
+\subsection{Trends}
+
+\begin{figure}
+  \centering
+  \includegraphics[width=1.0\textwidth]{images/graph/combined_navs}
+
+  \caption{Etc}
+  \label{fig:searchtypes}
+\end{figure}
+
+\subsection{Breaking them apart}
+
+  behavioral patterns
+  Direct search for specific course
+  Discovery, browsing, exploring
+
+  \subsubsection{Direct searches}
+
+  \begin{figure}
+    \centering
+    \includegraphics[width=1.0\textwidth]{images/graph/direct_navs}
+
+    \caption{Etc}
+    \label{fig:searchtypes}
+  \end{figure}
+
+  Why would 0-navs be so common with direct searches? 65\% subsections!
+
+  \subsubsection{Browse}
+
+  % STAT: Find major of users, find how many non-major courses they found
+
+  \begin{figure}
+    \centering
+    \includegraphics[width=1.0\textwidth]{images/graph/browse_navs}
+
+    \caption{Etc}
+    \label{fig:searchtypes}
+  \end{figure}
+
+  As expected, the 0-navs were mostly maincourses
+
+  Effective++
+
+  \subsubsection{Social}
 
@@
 
-\subsection{Conclusion}
+  success?????
 
-Success! Considering skedge is OPTIONAL.
-+ course blocks (obv usecase, can't click)
-+ exports (not supported by thing)
-+ mobile
-
-\section{Navigations-per-add}
-
-\subsection{Definitions}
-
-A navigation is defined as
-a search, or
-a click on an instructor’s name, or
-a click on a crosslisted or prerequisite course link
-
-The navigations-per-{add, bookmark} measure is
-the number of navigations a user took (within one session) until a course was {added, bookmarked}
-
-\subsection{Trends}
-
-%figs
-
-\subsection{Breaking them apart}
-
-  behavioral patterns
-  Direct search for specific course
-  Discovery, browsing, exploring
-
-  \subsubsection{Direct searches}
-
-  \subsubsection{Browse}
-
-  % STAT: Find major of users, find how many non-major courses they found
-
-\subsection{Conclusion}
-
-%figs
-
-Effective++
-
-\section{Users' search types over time}
+\section{Complexity of users' searches over time}
 
@@
 
-% fig
+\begin{figure}
+  \centering
 
-First increase (60.5% of users)
+  \includegraphics[width=1.00\textwidth]{images/graph/search_dt}
+
+  \\
+  \vspace{5pt}
+
+  \includegraphics[width=0.65\textwidth]{images/table/search_dt}
+
+  \caption{Etc}
+  \label{fig:searchtypes}
+\end{figure}
+
+First increase (60.5\% of users)
 Median: 2 searches
@@
 
-Second increase (7.9% of users)
+Second increase (7.9\% of users)
 Median: 8 searches
@@
 
-\subsection{Conclusion}
-
-DSL++
+DSQL++

### 2016-04-12 tech.tex
More design
work on social... so slow
Small fixes
Analytics work

@@
 
-\section{Back-end}
+\section{Architecture}
 
-Skedge's infrastructure is built
+\emph{nginx}, \emph{unicorn}, \emph{Ruby on Rails}, \emph{PostgreSQL}, \emph{React.js},
 
-\emph{nginx}, \emph{unicorn}, \emph{Ruby on Rails}, \emph{PostgreSQL}, \emph{React.js}, \emph{Ahoy}, and Google Analytics.
+\section{Data storage}
 
-\section{Front-end}
+cookies, likes not on fb
 
-\section{Analytics}
+\section{Usage data collection}
+
+As the saying goes, ``if it's not monitored, it's not in production'' \cite{monitored} (how else would we know that it was?) anonymous, \emph{Ahoy} and Google Analytics.

### 2016-04-12 design/modernity.tex
New bibliography and citations

--- 2016-04-11 design-modernity.tex
@@
 According to Mary Meeker's 2015 Mobile Technology Trends from Kleiner Perkins Caufield \& Byers\cite{kpcb}, in 2014, 51\% of adult time spent per day on the Internet was from a mobile device, versus 42\% spent on a desktop computer or laptop.
-Time spent on the Internet with mobile devices reached three hours per day in 2014, compared to less than one hour in 2010 (12\% mobile vs.
+Time spent on the Internet with mobile devices reached three hours per day in 2014, compared to less than one hour in 2010 (when there was 12\% mobile vs.
 75\% desktop/laptop time share).
@@
 Undeniably, supporting mobile devices and tablets in web applications is crucial for usability nowadays.
-Note how Skedge is responsive to the user's device in Figure \ref{fig:sk-mobile}, compared with CDCS's lack of mobile support in Figure \ref{fig:cdcs-mobile}.
+Note how Skedge responds to the device form-factor in Figure \ref{fig:sk-mobile}, compared with CDCS's lack of mobile support in Figure \ref{fig:cdcs-mobile}.
 CDCS on mobile requires the user to pinch and drag around both to read results and to make new searchs, while Skedge adapts content to the device's screen and fixes the search bar to the top of the screen for easy access while browsing.
 
-Moreover, since no major mobile browser currently supports browser extensions (and if one did, the extensions themselves would most likely need to be re-architected), CDCS on a mobile device loses all scheduling functionality, unlike with Skedge on mobile.
+Moreover, since no major mobile browser currently supports browser extensions (and if one did, the extensions themselves would most likely need to be re-architected), CDCS on a mobile device loses all scheduling functionality, unlike Skedge.
 
@@
 With the increasing number of attendees at University of Rochester's hackathons, it is clear that the University's ``hacker culture'' is growing---more students are collaborating to build side-projects that integrate resources and servies often benefitting the student community.
-Open-source and open-information services greatly help to foster such innovation, and having public APIs is essential toward this end.
+Open-source and open-information services greatly help to foster such innovation, and having public APIs is essential toward this end.\cite{Milberry}\cite{hackers}
 

### 2016-04-12 design/social.tex
New bibliography and citations

@@
 The question ``what are you taking this semester?'' may possibly be the most common phrase uttered in campus smalltalk within the first few weeks of a semester.
-But students' social lives do not solely exist in-person---social media dominates college social lives etc, v impt to consider [CITE].
+But college students' social lives do not solely exist in person---lately they are dominated largely by social media, most especially Facebook \cite{fb}.
 The ``what are you taking'' question typically translates to social media in the form of the schedule image exported from Better CDCS posted to Facebook (e.g.
@@
 
-  CITE THAT NOTIFICATION BADGES ARE PAVLOVIAN
+  Skedge users therefore do not have to seek out peer schedules---the schedules come to them.
+Social network notifications are an established, successful model that have been shown to be highly important for social networks' users and levels of engagement \cite{notifications}.
+The hope is that Skedge Social's notifications will evoke the almost Pavlovian ``click-and-check'' response in users, reminding them of the course network they are a part of and promoting a stronger sense of academic community.
 
-  \begin{figure}
+  \begin{figure}[H]
     \centering

### 2016-04-12 paper.tex
Analytics work
New bibliography and citations

@@
 \usepackage{setspace}
-
+\usepackage[comma,authoryear]{natbib}
 \usepackage[utf8]{inputenc}
@@
 
-\date{April 12, 2016\\
+\date{April 13, 2016\\
     \vspace{100pt}\\
@@
 
-\listoftables
-
 \listoffigures
@@
 As a result, Skedge boasts strong user-retention rates, long session durations, and high student adoption despite having virtually no advertisement.
-Through collected usage data, I demonstrate that \textbf{a)} Skedge's differences from and additions to CDCS are usable and have real-world need, \textbf{b)} the three major use-cases associated with course browsing---direct search, exploratory search, and peer recommendation---are effectively accommodated by Skedge, and \textbf{c)} Skedge's novel search mechanism is user-friendly and self-teaches to users over time.
+Through collected usage data, I demonstrate that \textbf{a)} Skedge's differences from and additions to CDCS are usable and have real-world need, \textbf{b)} the three major use-cases associated with course scheduling---direct search, exploratory search, and peer recommendation---are effectively accommodated by Skedge, and \textbf{c)} Skedge's novel search mechanism is user-friendly and self-teaches to users over time.
 
@@
 
-\begin{thebibliography}{1}
-
-	\bibitem{kpcb} Meeker, Mary (2015). http://www.slideshare.net/kleinerperkins/internet-trends-v1
-
-    \bibitem{caps} Wheildon, Colin (1995). \emph{Type and Layout: How Typography and Design Can Get your Message Across - Or Get in the Way.} Berkeley: Strathmoor Press. p. 62. ISBN 0-9624891-5-8.
-
-\end{thebibliography}
+% References
 
 \addcontentsline{toc}{chapter}{Bibliography}
+\singlespacing
+
+% Bibtex commands go here
+\bibliographystyle{ieeetr}
+\bibliography{bibliography}
 

### 2016-04-12 conclusion.tex
More design
More splitup
Small fixes
New bibliography and citations
open letter. can't believe there was a  wounded bat in the kitchen, that slowed me down

@@
-\chapter{Conclusions}
+\chapter{Conclusion}
 
+\vspace{-20pt}
 \section{Proposal to the University}
+
+On behalf of myself and the roughly \emph{one-thousand} students that regularly and voluntarily choose to use Skedge, I urge the University to consider adopting it as an official replacement to CDCS.
+
+Given the excellence of the University's inspiring faculty and academic programs, I know that I am not alone in my disappointment with what I find is a lack of regard for many student tools, as well as the lack of communication channels with students to improve them.
+Throughout the course of this paper, I have shown that the space of course scheduling holds vast potential for innovation, and that CDCS barely scratches that surface.
+An officially adopted Skedge---or an otherwise substantial overhaul of CDCS---would demonstrate that the University pays attention to its students and to how they use the tools it provides them.
+
+While I am aware that such matters are not extremely pressing, they nevertheless degrade student experience on campus and discolor perception of the administration's care for students, especially when modern, relevant, and engaging tools are within arm's reach, and \emph{most} especially at a time when University of Rochester's very high tuition rates continue to rise year-to-year.
+
+As a graduating senior, I cannot commit to continuing the upkeep of server costs and system maintenance for Skedge in the long term, and certainly not indefinitely.
+For the sake of the University's students, parents, faculty, and staff, I hope that my efforts with Skedge were not in vain and can lead to a substantial refresh of University tools that, most crucially, \textbf{involves a dialogue between the administration, students, and staff}.
+Thank you for reading.
+
+\\
+\vspace{10pt}
+
+\noindent (The views expressed in this section are entirely my own and were not reviewed by my thesis advisor or any others.)
 
@@
 
-\noindent The site can be found at: \url{http://skedgeur.com}.
+\noindent The site can be found at \url{http://skedgeur.com} or \url{http://skedge.org} (very recently acquired, simply redirects to the former).
+
+\section{Acknowledgments}
+
+I would like to give my foremost thanks to Dr.
+Philip Guo, who very thoughtfully reached out to \emph{me} about using Skedge for research, reigniting its development.
+I also want to give enormous thanks to Ani and Mark Gabrellian for their generous Mesrob Mashtots Innovation Grant donation which funded Skedge's development in the summer of 2015.
+Finally, a huge thanks to all those students and my close friends who reported bugs, gave feedback, and in general supported, promoted, got excited about, and complained about Skedge!

### 2016-04-12 conclusion.tex
conclusion

@@
 I would like to give my foremost thanks to Dr.
-Philip Guo, who very thoughtfully reached out to \emph{me} about using Skedge for research, reigniting its development.
+Philip Guo, who very thoughtfully reached out to \emph{me} about using Skedge for research, and whose deep care and support for the project has since reignited and bloomed its development.
 I also want to give enormous thanks to Ani and Mark Gabrellian for their generous Mesrob Mashtots Innovation Grant donation which funded Skedge's development in the summer of 2015.

### 2016-04-12 analytics.tex
Analytics

@@
 
-  \item Skedge’s \textbf{DSQL is user-friendly}; users learn more advanced search types over time \textbf{simply by using it}.
+  \item Skedge’s \textbf{DSQL is user-friendly}; users learn how to make more sophisticated search queries over time \textbf{simply by using it}.
 \end{enumerate}
@@
 \begin{itemize}
-  \item 4,713 users that have added or bookmarked at least 1 course
+  \item 4,713 users that have added or bookmarked at least one course
   \item 5,256 schedules
-  \item 17,012 sessions\footnote{A session is defined by a period of activity expiring after 30 minutes of inactivity}, 75\% of which are from returning visitors
-  \item 107 sessions per day, on average
-  \item 5.40 pages viewed per session, on average 
-  \item 6 minutes spent per session, on average
+  \item 17,124 sessions\footnote{A session is defined by a period of activity expiring after 30 minutes of inactivity}, 75\% of which are from returning visitors
+  \item an average of 107 sessions per day
+  \item an average of 5.4 pages viewed per session 
+  \item an average of 6 minutes spent per session
 \end{itemize}
 
-\begin{figure}
+\subsection{Search}
+
+\begin{figure}[H]
   \centering
@@
 
-  \caption{Etc}
+  \caption{Searches over time, with major relevant calendar events annotated}
   \label{fig:searches}
@@
 
-periodic visit cycle graph means people are actually using it and decided to stay with it, not declining, can be seen w exports & link-searches.
+Figure \ref{fig:searches} shows user search submissions over the time, and makes the periodicity of Skedge usage extremely apparent.
+Although it only shows a bit more than one ``phase,'' the cycle of events---\emph{semester begins} (students check course times, room locations, etc.
+while they get accustomed to the semester), \emph{next semester's course catalog released} (students explore the courses offered and build possible schedules), \emph{next semester registration opens} (cross-checking and officially registering each course with the Registrar)---clearly drives the majority of site traffic.
 
-also should be noted that NO seniors participated in the last, the majority of the userbase!
+What this means is that Skedge is not seen as an experiment or a proof of concept that is not production-ready.
+Rather, its traffic patterns show that it is consistenly \emph{relied upon} by students and used for its intended purpose in times of biggest course scheduling need.
+%, and the same can be seen with other usage measures such as schedule exporting and clicking on the Wikipedia-style embedded links (Figures \ref{fig:exports} and \ref{fig:linksearches}, respectively).
 
-success Considering skedge is OPTIONAL.
+It should also be noted that before March, the majority of the userbase as I knew it consisted of seniors, as the majority of people that I knew personally and shared Skedge with belonged to my class.
+The latest two major events reported in the graph above thus very impressively consist of \emph{very few seniors}, who have no courses left to schedule.
+March also marked the very first time I publicly advertised Skedge, although not very strongly.
+It was sent out to all students in the Computer Science mailing list as well as posted to two class-wide Facebook groups.
+
+The usability of and user satisfaction with a tool is hard to measure when it holds a monopoly.
+Considering that Skedge is an \emph{optional alternative} to CDCS, such consistent usage data (especially with 75\% of sessions being from returning users) undeniably prove its measures of usability and user satisfaction relative to CDCS.
+
+  \subsubsection{Search types}
+
+  \begin{figure}[H]
+    \centering
+    \includegraphics[width=1.0\textwidth]{images/graph/searchtypes}
+
+    \caption{Breakdown of search queries by type of search}
+    \label{fig:searchtypes}
+  \end{figure}
+
+Figure \ref{fig:searchtypes} validates the hypothesis made in Section 2.3 regarding the search interface design, demonstrating that the \emph{department code} and \emph{course number} search modes dominate user searches, accounting for over 85\% of all searches on Skedge.
+With this knowledge, bringing out the tucked-away, nondistinct field for this search in CDCS into prominence is a usability design choice that accomodates empirical usage.
+
+\subsubsection{Empty searches}
+
+  Of the total 45,466 searches on Skedge logged since November 2nd 2015, 89.4\% (40,625) of searches produce nonempty search results, whereas 10.6\% (4,841) of searches come up empty.
+This already is evidence of an effective search mechanism, but of those empty searches, the vast majority
+
+  Can learn from these searches.
+
+  \begin{figure}
+  {\renewcommand{\arraystretch}{1.5}
+  \centering
+  \begin{tabular}{ p{6cm} p{7.5cm} }
+  \hline
+
+    ``{\tt Prereq 280}'' \newline ``{\tt has-Prereq: 280}''
+    & Course dependencies \\ \hline
+
+    ``{\tt tuesday 3:25-4:40}'' \newline ``{\tt TR 3:25-4:40}'' \newline ``{\tt m/w}'' \newline ``{\tt classes tuesdays and thursdays between 11:05 and 12:20}'' \newline ``{\tt weekend summer 2016}''
+    & \parbox[c]{\hsize}{Better day-of-the-week handling} \\ \hline
+
+    ``{\tt 2 credits natural science}'' 
+    & Search by division (although there were very few) \\ \hline
+
+    ``{\tt religion and classics}'' \newline ``{\tt studio art}'' 
+    & Search by full department name \\ \hline
+
+    ``{\tt new csc courses}'' 
+    & The word ``{\tt courses}'' (somewhat common) \\ \hline
+
+    ``{\tt guo}'' 
+    & Instructor without ``{\tt taught by}'' (very common) \\ \hline
+
+    ``{\tt stt212 vs mth201}'' \newline ``{\tt stt212 OR mth201}'' 
+    & Searching multiple queries at the same time \\ \hline
+
+    ``{\tt crosslisted csc lin}'' 
+    & Probably a more reasonable syntax than the current Skedge DSQL one (``{\tt csc x lin}'') \\ \hline
+
+    ``{\tt place: hubbell}'' \newline ``{\tt hubbell friday}'' 
+    & Search by location, although motives unclear \\ \hline
+
+    ``{\tt MTH 2*}'' 
+    & Using ``{\tt *}'' instead of an omission (e.g. ``{\tt MTH 2}'' would work) \\ \hline
+
+    \hline
+  \end{tabular}
+  \caption{Example of search queries that produced no results but that indicate demand for added search features.
+The corresponding features that currently lack in the Skedge DSQL are described on the right.}
+  \end{figure}
+
+\begin{figure}
+  {\renewcommand{\arraystretch}{1.5}
+  \centering
+  \begin{tabular}{ | p{4cm} | p{4cm} | p{4cm} | }
+  \hline
+
+    ``{\tt Larry david}''
+    & ``{\tt taught by seligman}''
+    & ``{\tt cool stuff}''
+    \\ \hline
+
+    ``{\tt game of thrones}''
+    & ``{\tt fun classes}''
+    & ``{\tt fun classes in csc}''
+    \\ \hline
+
+    ``{\tt easy A}''
+    & ``{\tt flute birdman}''
+    & ``{\tt how do i make a new schedule}''
+    \\ \hline
+
+    \hline
+  \end{tabular}
+  \caption{Humerous search submissions}
+  \end{figure}
 
@@
 
-\begin{figure}
+\begin{figure}[H]
   \centering
@@
   \caption{Etc}
-  \label{fig:searchtypes}
+  \label{fig:exports}
 \end{figure}
@@
   \caption{Etc}
-  \label{fig:searchtypes}
+  \label{fig:linksearches}
 \end{figure}
@@
 success because you can't click schedule blocks in better cdcs!!
-
-\subsection{Search}
-
-  \subsubsection{Search types}
-
-  \begin{figure}
-    \centering
-    \includegraphics[width=1.0\textwidth]{images/graph/searchtypes}
-
-    \caption{Etc}
-    \label{fig:searchtypes}
-
-  \end{figure}
-
-  \subsubsection{Empty searches}
-
-  Can learn from these
-  Some funny ones
 
@@
 
-  Why would 0-navs be so common with direct searches? 65\% subsections!
+  Why would 0-navs be so common with direct searches? 64\% of them subsections!
 
   \subsubsection{Browse}
-
-  % STAT: Find major of users, find how many non-major courses they found
 
@@
 
-  As expected, the 0-navs were mostly maincourses
+  As expected, the 0-navs were mostly maincourses (62\%)
 
@@
 
-First increase (60.5\% of users)
-Median: 2 searches
-Average: 4.23 searches
-(Starting at 1 counts as an increase value of 0)
+First increase 69.4\% of users), with a median of 2 searches.
+Note that starting at search complexity one or greater counts as a ``searches before first increase value'' of 0.
 

### 2016-04-12 conclusion.tex
Analytics

@@
 
-On behalf of myself and the roughly \emph{one-thousand} students that regularly and voluntarily choose to use Skedge, I urge the University to consider adopting it as an official replacement to CDCS.
+On behalf of myself and the roughly \emph{one-thousand} students who regularly and voluntarily choose to use Skedge, I urge the University to consider adopting it as an official replacement to CDCS.
 
-Given the excellence of the University's inspiring faculty and academic programs, I know that I am not alone in my disappointment with what I find is a lack of regard for many student tools, as well as the lack of communication channels with students to improve them.
+Given the excellence of the University's inspiring faculty and academic programs, I know that I am not alone in my disappointment with what I consider a low regard for usability in many student tools, as well as the lack of communication channels with students to improve them.
 Throughout the course of this paper, I have shown that the space of course scheduling holds vast potential for innovation, and that CDCS barely scratches that surface.

### 2016-04-12 design/search.tex
Small fixes
Analytics

@@
 
-Of course, the most important usability concern for a course explorer/scheduler is being able to effectively \emph{find courses}.
-
-In this section, I will present the \emph{Skedge DSQL}, a domain-specific query language that is based on natural language.
+The most important usability concern of all for a course explorer/scheduler is being able to effectively \emph{find courses}.
+In this section, I will present \emph{SkedgeQL}, a domain-specific query language that is free-form and based on natural language.
 Next, I will demonstrate how Skedge leverages the DSQL to handle the three search criteria students have for finding courses better than CDCS does.
@@
 This prompted me to closely examine the fields to a) determine redundancies between them and b) find how to make some of the valuable filter fields more usable.
-From this came Skedge's unified, natural language based search method, which I call the \emph{Skedge Domain-Specific Query Language (Skedge DSQL)}.
+From this came Skedge's unified, natural language based search method, which I call the Skedge Query Language, or \emph{SkedgeQL}.
 Figure \ref{fig:sk-search} is a list of example possible searches shown to users as they type.
@@
 \begin{enumerate}
-  \item \emph{Subject} (department).
-  \item \emph{Course} (department code and/or course number, e.g. ``csc 171'').
-  \item \emph{Title} (course title).
+  \item \emph{Subject} (department)
+  \item \emph{Course} (department code and/or course number, e.g. ``csc 171'')
+  \item \emph{Title} (course title)
 \end{enumerate}
@@
 
-    In addition, once learned and remembered, using a natural language based query language can be faster and more intuitive than clicking on form fields.
-And more importantly, the Skedge DSQL is more easily extendable as new search features are invented.
+    It has been shown that once learned, natural language based query languages can be faster, more intuitive, and more usable than traditional form-based methods \cite{dsl}.
+WolframAlpha's hugely successful online free-form search is a notable pioneer in this field \cite{wolfram}.
+
+    Lastly, the Skedge DSQL is more easily extendable as new search features are invented.
 For CDCS, the more form fields that are added to the sidebar, the less usable its user interface becomes.

### 2016-04-12 analytics.tex
SkedgeQL

@@
 
-  \item Skedge’s \textbf{DSQL is user-friendly}; users learn how to make more sophisticated search queries over time \textbf{simply by using it}.
+  \item \textbf{SkedgeQL is user-friendly}; users learn how to make more sophisticated search queries over time \textbf{simply by using it}.
 \end{enumerate}
@@
 
-  Can learn from these searches.
+  Can learn from these searches. This is the same method WolframAlpha describes its team using to improve their freeform input system \cite{wolfram2}.
 
@@
     ``{\tt crosslisted csc lin}'' 
-    & Probably a more reasonable syntax than the current Skedge DSQL one (``{\tt csc x lin}'') \\ \hline
+    & Probably a more reasonable syntax than the current SkedgeQL one (``{\tt csc x lin}'') \\ \hline
 
@@
   \caption{Example of search queries that produced no results but that indicate demand for added search features.
-The corresponding features that currently lack in the Skedge DSQL are described on the right.}
+The corresponding features that currently lack in SkedgeQL are described on the right.}
   \end{figure}

### 2016-04-12 design/search.tex
SkedgeQL

@@
 In this section, I will present \emph{SkedgeQL}, a domain-specific query language that is free-form and based on natural language.
-Next, I will demonstrate how Skedge leverages the DSQL to handle the three search criteria students have for finding courses better than CDCS does.
+Next, I will demonstrate how Skedge leverages SkedgeQL to handle the three search criteria students have for finding courses better than CDCS does.
 
@@
       }
-      \caption{Examples of the Skedge DSQL} \label{fig:sk-search}
+      \caption{Examples of SkedgeQL} \label{fig:sk-search}
     \end{subfigure}
@@
   \includegraphics[width=14cm]{images/search-mapping}
-  \caption{Mapping from CDCS's form-based search to the Skedge DSQL} \label{fig:search-mapping}
+  \caption{Mapping from CDCS's form-based search to SkedgeQL} \label{fig:search-mapping}
 \end{figure}
 
-Figure \ref{fig:search-mapping} is a mapping of each CDCS field to an element of the Skedge DSQL or unified search.
+Figure \ref{fig:search-mapping} is a mapping of each CDCS field to an element of SkedgeQL or its unified search.
 Dashed lines that end abruptly are fields that were deemed unnecessary or redundant due to a Skedge-unique feature.
@@
 
-  \item \emph{Division} (i.e. ``Humanities'', ``Social Sciences'', etc.) was removed because it seems too broad to be useful, but it could be added to the Skedge DSQL if there is demand.
+  \item \emph{Division} (i.e. ``Humanities'', ``Social Sciences'', etc.) was removed because it seems too broad to be useful, but it could be added to SkedgeQL if there is demand.
 
@@
 
-The \emph{unified search} components can be entered directly into the single Skedge search field, along with the Skedge DSQL. It includes the CDCS fields:
+The \emph{unified search} components can be entered directly into the single Skedge search field, along with SkedgeQL. It includes the CDCS fields:
 
@@
 
-\subsubsection{Skedge DSQL (right side)}
+\subsubsection{SkedgeQL (right side)}
 
-The \emph{Skedge DSQL} is enterable using special keywords and, optionally, qualifiers for those keywords. It includes the CDCS fields:
+SkedgeQL is enterable using special keywords and, optionally, qualifiers for those keywords. It includes the CDCS fields:
 
@@
 
-  \subsubsection{Advantages of the DSQL}
+  \subsubsection{Advantages of SkedgeQL}
 
@@
 
-    Lastly, the Skedge DSQL is more easily extendable as new search features are invented.
+    Lastly, SkedgeQL is more easily extendable as new search features are invented.
 For CDCS, the more form fields that are added to the sidebar, the less usable its user interface becomes.
-For Skedge, while it may take longer for users to learn a bigger DSQL, there is no issue with crowding user interface real estate.
+For Skedge, while it may take longer for users to learn a bigger query language, there is no issue with crowding user interface real estate.
 
-  \subsubsection{Disadvantages of the DSQL}
+  \subsubsection{Disadvantages of SkedgeQL}
 
-    One disadvantage of Skedge's DSQL is the possible grammatical ambiguity of some queries.
+    One disadvantage of SkedgeQL is the possible grammatical ambiguity of some queries.
 For instance, the query ``Fall of the Roman Empire'' could be searching for courses with that as the title, or searching for courses in the \emph{fall term} with the title ``of the Roman Empire''.
 Unless multiple queries are run and more complex natural language processing is done, the system would have a hard time disambiguating between the two.
-This could also be solved with a ``did you mean?'' feature, which disambiguates a query into every possible DSQL breakup.
+This could also be solved with a ``did you mean?'' feature, which disambiguates a query into every possible SkedgeQL breakup.
 
-    Another notable disadvantage of the DSQL is having to know it, and it having a possibly steep learning curve for some users.
-However, the argument will be made in Chapter 4 that simply by using the site over time, users will self-learn the Skedge DSQL.
+    Another notable disadvantage of SkedgeQL is having to know it, and it having a possibly steep learning curve for some users.
+However, the argument will be made in Chapter 4 that simply by using the site over time, users will self-learn Skedge.
 One of the ways this is possible is thanks to Skedge's search system being \emph{multi-purposed}.
@@
 The three cases are \textbf{requirements}, \textbf{electives}, and \textbf{peer recommendations}.
-The Skedge DSL and other application features offer substantial improvements over CDCS for each of these cases.
+SkedgeQL and other application features offer substantial improvements over CDCS for each of these cases.
 
@@
 
-    \item The Skedge DSQL includes the ``new'' keyword, which qualifies an accompanying search by only displaying the courses that were not offered the previous year for that term.
+    \item SkedgeQL includes the ``new'' keyword, which qualifies an accompanying search by only displaying the courses that were not offered the previous year for that term.
 For instance, searching ``new csc'' today (Fall 2016) would show Fall Computer Science courses that have been added to the catalog or changed since Fall 2015.
-    \item The Skedge DSQL could include an ``autofit'' keyword, which would take into account the user's current schedule, and only show sections that would not conflict with any already scheduled times.\footnote{This feature is under development and is not currently live.}
-    \item The Skedge DSQL includes the ``random'' keyword, which displays a single random course matching the accompanying query.
+    \item SkedgeQL could include an ``autofit'' keyword, which would take into account the user's current schedule, and only show sections that would not conflict with any already scheduled times.\footnote{This feature is under development and is not currently live.}
+    \item SkedgeQL includes the ``random'' keyword, which displays a single random course matching the accompanying query.
 For instance, ``random csc'' would show a random course in Computer Science for the current term.
 Used in conjunction with ``autofit'', it can a powerful way to find interesting new classes that are available for users to try.
-    \item The Skedge DSQL includes sorting by class size---useful for students trying to find smaller, discussion based classes---or by class starting time---for students with preferences of morning or afternoon/evening classes.
+    \item SkedgeQL includes sorting by class size---useful for students trying to find smaller, discussion based classes---or by class starting time---for students with preferences of morning or afternoon/evening classes.
     \item Instructor names appearing in courses automatically link to searches for more courses taught by that instructor, which is easier than having to use the CDCS ``instructor'' field to make a separate search.

### 2016-04-12 analytics.tex
add export

@@
 Rather, its traffic patterns show that it is consistenly \emph{relied upon} by students and used for its intended purpose in times of biggest course scheduling need.
-%, and the same can be seen with other usage measures such as schedule exporting and clicking on the Wikipedia-style embedded links (Figures \ref{fig:exports} and \ref{fig:linksearches}, respectively).
 
@@
 Figure \ref{fig:searchtypes} validates the hypothesis made in Section 2.3 regarding the search interface design, demonstrating that the \emph{department code} and \emph{course number} search modes dominate user searches, accounting for over 85\% of all searches on Skedge.
-With this knowledge, bringing out the tucked-away, nondistinct field for this search in CDCS into prominence is a usability design choice that accomodates empirical usage.
+With this knowledge, bringing out the tucked-away, nondistinct field for this search in CDCS into prominence is a usability design choice that accomodates an empirically verified use case.
 
@@
   Of the total 45,466 searches on Skedge logged since November 2nd 2015, 89.4\% (40,625) of searches produce nonempty search results, whereas 10.6\% (4,841) of searches come up empty.
-This already is evidence of an effective search mechanism, but of those empty searches, the vast majority
 
-  Can learn from these searches. This is the same method WolframAlpha describes its team using to improve their freeform input system \cite{wolfram2}.
+  This already is evidence of an effective search mechanism, but of those empty searches, the vast majority (92\%) contain typos or are for courses that simply don't exist.
+The other significant portion of searches include instructor names without the requisite ``taught by'' for SkedgeQL to pick it up as an instructor search, and such a fix could easily be made.
+
+  This type of analysis can powerfully indicate holes in the query language (i.e.
+searches intuitive for users) that ought to be fixed, which is the same method WolframAlpha describes its team using to improve their freeform input system \cite{wolfram2}.
+Figure \ref{fig:empty-queries} lists a few notable examples of this nature.
+(Figure \ref{fig:funny-queries} lists a few funny queries I came across.)
 
@@
 
-    ``{\tt tuesday 3:25-4:40}'' \newline ``{\tt TR 3:25-4:40}'' \newline ``{\tt m/w}'' \newline ``{\tt classes tuesdays and thursdays between 11:05 and 12:20}'' \newline ``{\tt weekend summer 2016}''
+    ``{\tt tuesday 3:25-4:40}'' \newline ``{\tt m/w}'' \newline ``{\tt classes tuesdays and thursdays between 11:05 and 12:20}'' \newline ``{\tt weekend summer 2016}''
     & \parbox[c]{\hsize}{Better day-of-the-week handling} \\ \hline
@@
     ``{\tt new csc courses}'' 
-    & The word ``{\tt courses}'' (somewhat common) \\ \hline
+    & The word ``{\tt courses}'' (somewhat common with advanced query types) \\ \hline
+
+    ``{\tt curriculum change}'' 
+    & Alternative for ``{\tt new}'' keyword \\ \hline
 
@@
 
-    ``{\tt stt212 vs mth201}'' \newline ``{\tt stt212 OR mth201}'' 
+    ``{\tt stt212 vs mth201}'' \newline ``{\tt stt212 OR mth201}'' \newline ``{\tt csc 2\{4,8\}2}'' \newline ``{\tt csc 242|282}''
     & Searching multiple queries at the same time \\ \hline
@@
 
-    ``{\tt MTH 2*}'' 
+    ``{\tt MTH 2*}'' \newline ``{\tt csc 2xx}''
     & Using ``{\tt *}'' instead of an omission (e.g. ``{\tt MTH 2}'' would work) \\ \hline
@@
 The corresponding features that currently lack in SkedgeQL are described on the right.}
+  \label{fig:empty-queries}
   \end{figure}
@@
     ``{\tt game of thrones}''
-    & ``{\tt fun classes}''
+    & ``{\tt interesting courses}''
     & ``{\tt fun classes in csc}''
@@
   \end{tabular}
-  \caption{Humerous search submissions}
+  \caption{Humorous search submissions}
+  \label{fig:funny-queries}
   \end{figure}
@@
 
-\begin{figure}[H]
+\begin{figure}
   \centering
@@
 
-  \caption{Etc}
+  \caption{Different types of schedule exporting over time}
   \label{fig:exports}
@@
 
-Success! (broken on cdcs)
+In Section 2.2.5 (on the design of Skedge's schedule exporting), a major usability improvement claimed was the ability to export schedules to the {\tt .ics} format, a widely used format for many calendar applications, as well as offering functional Google Calendar export.
+Figure \ref{fig:exports} shows usage of the export functions over time, demonstrating its consistent and healthy use (traffic surges once again in line with major scheduling events.) This data confirms that there is strong demand for exporting to {\tt .ics}---it is even more common than Google Calendar export.
 

### 2016-04-12 design/usability.tex
add export

@@
 Additionally, a common feature request for Skedge is the implementation of multiple schedules \emph{per} term, allowing users to experiment with different schedule possibilities.\footnote{Currently under development, but simple to implement given the existing schedule management structure.}
+
+\subsection{Schedule export}
+
+Better CDCS provides two ways for users to export their schedule---as image or to Google calendar--except that the Google Calendar export feature is completely non-functional, telling users that ``{\tt This export tool will only work properly for the Spring of 2015}'' only \emph{after} they authorized the tool to access their Google account.
+
+Notably missing as an export option is the \emph{iCalendar} file format (extension {\tt .ics}), which can be imported into many calendar software products.
+It is the only format the Mac OS and iOS Calendar applications can accept, to offer one example of very popular products that are therefore unsupported by Better CDCS.
+
+Skedge provides the option for users to export their own or another user's schedule directly to \emph{Google Calendar} (even allowing users to choose a calendar---or create a new one for Skedge---to insert their courses into), as an \emph{image} (JPEG, PNG, or SVG), or in {\tt .ics} format (see Figure \ref{fig:sk-export}).
+The {\tt .ics} and Google Calendar exports include course codes, titles, locations, and properly repeat courses until the end of the current semester.
+
+\vspace{20pt}
+
+\begin{figure}[ht]
+  \centering
+  \vspace{2pt}
+  \includegraphics[width=14cm]{images/skedge/export}
+  \caption{Exporting to Google Calendar (note the other export options on the top right)} \label{fig:sk-export}
+\end{figure}
+
+\clearpage

### 2016-04-13 analytics.tex
Link searches

@@
     ``{\tt new csc courses}'' 
-    & The word ``{\tt courses}'' (somewhat common with advanced query types) \\ \hline
+    & The word ``{\tt courses}'' stimies the query here (somewhat common with advanced query types) \\ \hline
 
@@
     ``{\tt MTH 2*}'' \newline ``{\tt csc 2xx}''
-    & Using ``{\tt *}'' instead of an omission (e.g. ``{\tt MTH 2}'' would work) \\ \hline
+    & Using ``{\tt *}'' or ``{\tt x}'' instead of an omission (e.g. searching ``{\tt MTH 2}'' would work for this purpose) \\ \hline
 
@@
   \centering
-  \includegraphics[width=1.0\textwidth]{images/graph/exports}
+  \includegraphics[width=12cm]{images/graph/exports}
 
@@
 
-  \caption{Etc}
+  \caption{Different types of ``link-search'' clicks over time, with manual instructor search included for comparison with clicking on an instructor's name}
   \label{fig:linksearches}
@@
 
-Success, used!
+Figure \ref{fig:linksearches} shows the amount of clicks on the Wikipedia-style embedded links throughout Skedge (instructor names to see other courses they are teaching, and prerequisite and crosslist course mentions to search those courses).
+
+The number of manual instructor searches are also included in the graph for comparison with clicking on an instructor's name.
+As can be seen, clicking on an instructor's linked name is vastly more usable than searching it manually, as well as possibly offering different functionality (e.g.
+in the course browsing paradigm).
+Either way, usage of ``link-searches'' shows that they fulfill a need lacking in CDCS.
 

### 2016-04-13 analytics.tex
separate out analytics, almost done with 4.1

--- 2016-04-12 analytics.tex
@@
 
-\section{Usage}
+\input{analytics/usage}
+\clearpage
 
-\subsection{General}
+\input{analytics/npa}
+\clearpage
 
-From November 3, 2015 to April 11, 2016 (a period of 159 days), Skedge has seen\ldots
-
-\begin{itemize}
-  \item 4,713 users that have added or bookmarked at least one course
-  \item 5,256 schedules
-  \item 17,124 sessions\footnote{A session is defined by a period of activity expiring after 30 minutes of inactivity}, 75\% of which are from returning visitors
-  \item an average of 107 sessions per day
-  \item an average of 5.4 pages viewed per session 
-  \item an average of 6 minutes spent per session
-\end{itemize}
-
-\subsection{Search}
-
-\begin{figure}[H]
-  \centering
-
-  \includegraphics[width=1.0\textwidth]{images/graph/searches}
-
-  \caption{Searches over time, with major relevant calendar events annotated}
-  \label{fig:searches}
-
-\end{figure}
-
-Figure \ref{fig:searches} shows user search submissions over the time, and makes the periodicity of Skedge usage extremely apparent.
-Although it only shows a bit more than one ``phase,'' the cycle of events---\emph{semester begins} (students check course times, room locations, etc.
-while they get accustomed to the semester), \emph{next semester's course catalog released} (students explore the courses offered and build possible schedules), \emph{next semester registration opens} (cross-checking and officially registering each course with the Registrar)---clearly drives the majority of site traffic.
-
-What this means is that Skedge is not seen as an experiment or a proof of concept that is not production-ready.
-Rather, its traffic patterns show that it is consistenly \emph{relied upon} by students and used for its intended purpose in times of biggest course scheduling need.
-%, and the same can be seen with other usage measures such as schedule exporting and clicking on the Wikipedia-style embedded links (Figures \ref{fig:exports} and \ref{fig:linksearches}, respectively).
-
-It should also be noted that before March, the majority of the userbase as I knew it consisted of seniors, as the majority of people that I knew personally and shared Skedge with belonged to my class.
-The latest two major events reported in the graph above thus very impressively consist of \emph{very few seniors}, who have no courses left to schedule.
-March also marked the very first time I publicly advertised Skedge, although not very strongly.
-It was sent out to all students in the Computer Science mailing list as well as posted to two class-wide Facebook groups.
-
-The usability of and user satisfaction with a tool is hard to measure when it holds a monopoly.
-Considering that Skedge is an \emph{optional alternative} to CDCS, such consistent usage data (especially with 75\% of sessions being from returning users) undeniably prove its measures of usability and user satisfaction relative to CDCS.
-
-  \subsubsection{Search types}
-
-  \begin{figure}[H]
-    \centering
-    \includegraphics[width=1.0\textwidth]{images/graph/searchtypes}
-
-    \caption{Breakdown of search queries by type of search}
-    \label{fig:searchtypes}
-  \end{figure}
-
-Figure \ref{fig:searchtypes} validates the hypothesis made in Section 2.3 regarding the search interface design, demonstrating that the \emph{department code} and \emph{course number} search modes dominate user searches, accounting for over 85\% of all searches on Skedge.
-With this knowledge, bringing out the tucked-away, nondistinct field for this search in CDCS into prominence is a usability design choice that accomodates empirical usage.
-
-\subsubsection{Empty searches}
-
-  Of the total 45,466 searches on Skedge logged since November 2nd 2015, 89.4\% (40,625) of searches produce nonempty search results, whereas 10.6\% (4,841) of searches come up empty.
-This already is evidence of an effective search mechanism, but of those empty searches, the vast majority
-
-  Can learn from these searches. This is the same method WolframAlpha describes its team using to improve their freeform input system \cite{wolfram2}.
-
-  \begin{figure}
-  {\renewcommand{\arraystretch}{1.5}
-  \centering
-  \begin{tabular}{ p{6cm} p{7.5cm} }
-  \hline
-
-    ``{\tt Prereq 280}'' \newline ``{\tt has-Prereq: 280}''
-    & Course dependencies \\ \hline
-
-    ``{\tt tuesday 3:25-4:40}'' \newline ``{\tt TR 3:25-4:40}'' \newline ``{\tt m/w}'' \newline ``{\tt classes tuesdays and thursdays between 11:05 and 12:20}'' \newline ``{\tt weekend summer 2016}''
-    & \parbox[c]{\hsize}{Better day-of-the-week handling} \\ \hline
-
-    ``{\tt 2 credits natural science}'' 
-    & Search by division (although there were very few) \\ \hline
-
-    ``{\tt religion and classics}'' \newline ``{\tt studio art}'' 
-    & Search by full department name \\ \hline
-
-    ``{\tt new csc courses}'' 
-    & The word ``{\tt courses}'' (somewhat common) \\ \hline
-
-    ``{\tt guo}'' 
-    & Instructor without ``{\tt taught by}'' (very common) \\ \hline
-
-    ``{\tt stt212 vs mth201}'' \newline ``{\tt stt212 OR mth201}'' 
-    & Searching multiple queries at the same time \\ \hline
-
-    ``{\tt crosslisted csc lin}'' 
-    & Probably a more reasonable syntax than the current SkedgeQL one (``{\tt csc x lin}'') \\ \hline
-
-    ``{\tt place: hubbell}'' \newline ``{\tt hubbell friday}'' 
-    & Search by location, although motives unclear \\ \hline
-
-    ``{\tt MTH 2*}'' 
-    & Using ``{\tt *}'' instead of an omission (e.g. ``{\tt MTH 2}'' would work) \\ \hline
-
-    \hline
-  \end{tabular}
-  \caption{Example of search queries that produced no results but that indicate demand for added search features.
-The corresponding features that currently lack in SkedgeQL are described on the right.}
-  \end{figure}
-
-\begin{figure}
-  {\renewcommand{\arraystretch}{1.5}
-  \centering
-  \begin{tabular}{ | p{4cm} | p{4cm} | p{4cm} | }
-  \hline
-
-    ``{\tt Larry david}''
-    & ``{\tt taught by seligman}''
-    & ``{\tt cool stuff}''
-    \\ \hline
-
-    ``{\tt game of thrones}''
-    & ``{\tt fun classes}''
-    & ``{\tt fun classes in csc}''
-    \\ \hline
-
-    ``{\tt easy A}''
-    & ``{\tt flute birdman}''
-    & ``{\tt how do i make a new schedule}''
-    \\ \hline
-
-    \hline
-  \end{tabular}
-  \caption{Humerous search submissions}
-  \end{figure}
-
-\subsection{Exports}
-
-\begin{figure}[H]
-  \centering
-  \includegraphics[width=1.0\textwidth]{images/graph/exports}
-
-  \caption{Etc}
-  \label{fig:exports}
-\end{figure}
-
-Success! (broken on cdcs)
-
-\subsection{``Link-searches''}
-
-\begin{figure}
-  \centering
-  \includegraphics[width=1.0\textwidth]{images/graph/linksearches}
-
-  \caption{Etc}
-  \label{fig:linksearches}
-\end{figure}
-
-Success, used!
-
-\subsection{Mobile}
-
-\begin{itemize}
-  \item 11.54\% of sessions
-  \item 2.74 page viewed per session, on average
-  \item 2 minutes spent per session, on average
-\end{itemize}
-
-success bc mobile doesn't exist on cdcs
-
-\subsection{Course Blocks}
-
-\begin{figure}
-  \centering
-
-  \fbox {
-    \includegraphics[width=8cm]{images/skedge/blocks}
-  }
-
-  \caption{Etc}
-  \label{fig:searchtypes}
-
-\end{figure}
-
-40\% of sessions have at least one block-click
-Average of 5.12 block-clicks per session
-
-success because you can't click schedule blocks in better cdcs!!
-
-\section{Effectiveness of search}
-
-\subsection{Definitions}
-
-A navigation is defined as
-a search, or
-a click on an instructor’s name, or
-a click on a crosslisted or prerequisite course link
-
-The navigations-per-add measure is
-the number of navigations a user took (within one session) until a course was added, bookmarked
-
-\subsection{Trends}
-
-\begin{figure}
-  \centering
-  \includegraphics[width=1.0\textwidth]{images/graph/combined_navs}
-
-  \caption{Etc}
-  \label{fig:searchtypes}
-\end{figure}
-
-\subsection{Breaking them apart}
-
-  behavioral patterns
-  Direct search for specific course
-  Discovery, browsing, exploring
-
-  \subsubsection{Direct searches}
-
-  \begin{figure}
-    \centering
-    \includegraphics[width=1.0\textwidth]{images/graph/direct_navs}
-
-    \caption{Etc}
-    \label{fig:searchtypes}
-  \end{figure}
-
-  Why would 0-navs be so common with direct searches? 64\% of them subsections!
-
-  \subsubsection{Browse}
-
-  \begin{figure}
-    \centering
-    \includegraphics[width=1.0\textwidth]{images/graph/browse_navs}
-
-    \caption{Etc}
-    \label{fig:searchtypes}
-  \end{figure}
-
-  As expected, the 0-navs were mostly maincourses (62\%)
-
-  Effective++
-
-  \subsubsection{Social}
-
-  90 users have linked Skedge to Facebook
-  Since March 1st,
-  4,000+ visits (200 visits/day)
-  ~60\% of visits to /social were returning visitors
-  90 overlays onto friends’ schedules
-  10 clicks to Facebook profiles :(
-  - get stats from the fb dashboard
-
-  success?????
-
-\section{Complexity of users' searches over time}
-
-\subsection{Definitions}
-
-Points for search by (omits number and dept.):
-
-description
-credits
-crosslisted
-CRN
-instructor
-title
-year
-term
-‘random’
-upper-level writing
-
-“CSC” → 0
-“MTH 165” → 0
-“taught by hema” → 1   ✓    (2 searches) 
-“random mur 1-2 credits” → 2   ✓    (1 search)
-
-\subsection{Trends}
-
-\begin{figure}
-  \centering
-
-  \includegraphics[width=1.00\textwidth]{images/graph/search_dt}
-
-  \\
-  \vspace{5pt}
-
-  \includegraphics[width=0.65\textwidth]{images/table/search_dt}
-
-  \caption{Etc}
-  \label{fig:searchtypes}
-\end{figure}
-
-First increase 69.4\% of users), with a median of 2 searches.
-Note that starting at search complexity one or greater counts as a ``searches before first increase value'' of 0.
-
-Second increase (7.9\% of users)
-Median: 8 searches
-Average: 17.52 searches
-
-DSQL++
+\input{analytics/dsl}
+\clearpage

### 2016-04-13 analytics/dsl.tex
separate out analytics, almost done with 4.1

\section{Complexity of users' searches over time}

\subsection{Definitions}

Points for search by (omits number and dept.):

description
credits
crosslisted
CRN
instructor
title
year
term
‘random’
upper-level writing

“CSC” → 0
“MTH 165” → 0
“taught by hema” → 1   ✓    (2 searches) 
“random mur 1-2 credits” → 2   ✓    (1 search)

\subsection{Trends}

\begin{figure}
  \centering

  \includegraphics[width=1.00\textwidth]{images/graph/search_dt}

  \\
  \vspace{30pt}

  \includegraphics[width=0.65\textwidth]{images/table/search_dt}

  \caption{Etc}
  \label{fig:searchtypes}
\end{figure}

First increase 69.4\% of users), with a median of 2 searches.
Note that starting at search complexity one or greater counts as a ``searches before first increase value'' of 0.

Second increase (7.9\% of users)
Median: 8 searches
Average: 17.52 searches

DSQL++

### 2016-04-13 analytics/npa.tex
separate out analytics, almost done with 4.1

--- 2016-04-09 03-analytics.tex
@@
-\chapter{Data Analytics}
+\section{Effectiveness of search}
 
-Hypotheses:
+\subsection{Definitions}
 
-1. Skedge's differences from and additions to CDCS are usable and have real need
+A navigation is defined as
+a search, or
+a click on an instructor’s name, or
+a click on a crosslisted or prerequisite course link
 
-2. Skedge’s navigations-per-add and other metrics demonstrate effectiveness of the use cases
-a) direct searching, and
-b) course browsing
+The navigations-per-add measure is
+the number of navigations a user took (within one session) until a course was added, bookmarked
 
-3. Skedge’s DSL is user-friendly; users learn more advanced search types over time by using it
+\subsection{Trends}
 
-\section{Usage}
+\begin{figure}
+  \centering
+  \includegraphics[width=1.0\textwidth]{images/graph/combined_navs}
 
-\subsection{General}
+  \caption{Etc}
+  \label{fig:searchtypes}
+\end{figure}
 
-Since November 3rd 2015 (137 days)
-3,768 unique users
-4,500 schedules
-Average 90 sessions/day
-Average 4.92 pages/session
-Average 5:31 minutes/session
-28\% of sessions are from new users
+\subsection{Breaking them apart}
 
-MOBILE RESULT
+  behavioral patterns
+  Direct search for specific course
+  Discovery, browsing, exploring
 
-\subsection{Search}
+  \subsubsection{Direct searches}
 
-% figs
+  \begin{figure}
+    \centering
+    \includegraphics[width=1.0\textwidth]{images/graph/direct_navs}
 
-  \subsubsection{Empty searches}
+    \caption{Etc}
+    \label{fig:searchtypes}
+  \end{figure}
 
-  Can learn from these
-  Some funny ones
+  Why would 0-navs be so common with direct searches? 64\% of them subsections!
 
-\subsection{Course blocks}
+  \subsubsection{Browse}
 
-40\% of sessions have at least one block-click
-Average of 4.94 block-clicks per session
+  \begin{figure}
+    \centering
+    \includegraphics[width=1.0\textwidth]{images/graph/browse_navs}
 
-\subsection{Social}
+    \caption{Etc}
+    \label{fig:searchtypes}
+  \end{figure}
+
+  As expected, the 0-navs were mostly maincourses (62\%)
+
+  Effective++
+
+  \subsubsection{Social}
 
@@
 
-\subsection{Conclusion}
-
-Success! Considering skedge is OPTIONAL.
-+ course blocks (obv usecase, can't click)
-+ exports (not supported by thing)
-+ mobile
-
-\section{Navigations-per-add}
-
-\subsection{Definitions}
-
-A navigation is defined as
-a search, or
-a click on an instructor’s name, or
-a click on a crosslisted or prerequisite course link
-
-The navigations-per-{add, bookmark} measure is
-the number of navigations a user took (within one session) until a course was {added, bookmarked}
-
-\subsection{Trends}
-
-%figs
-
-\subsection{Breaking them apart}
-
-  behavioral patterns
-  Direct search for specific course
-  Discovery, browsing, exploring
-
-  \subsubsection{Direct searches}
-
-  \subsubsection{Browse}
-
-  % STAT: Find major of users, find how many non-major courses they found
-
-\subsection{Conclusion}
-
-%figs
-
-Effective++
-
-\section{Users' search types over time}
-
-\subsection{Definitions}
-
-Points for search by (omits number and dept.):
-
-description
-credits
-crosslisted
-CRN
-instructor
-title
-year
-term
-‘random’
-upper-level writing
-
-“CSC” → 0
-“MTH 165” → 0
-“taught by hema” → 1   ✓    (2 searches) 
-“random mur 1-2 credits” → 2   ✓    (1 search)
-
-\subsection{Trends}
-
-% fig
-
-First increase (60.5% of users)
-Median: 2 searches
-Average: 4.23 searches
-(Starting at 1 counts as an increase value of 0)
-
-Second increase (7.9% of users)
-Median: 8 searches
-Average: 17.52 searches
-
-\subsection{Conclusion}
-
-DSL++
+  success?????

### 2016-04-13 analytics/usage.tex
separate out analytics, almost done with 4.1

\section{Usage}

From November 3, 2015 to April 11, 2016 (a period of 159 days), Skedge has seen\ldots

\begin{center}
\begin{tabular}{l}
  \hline
  4,713 users that have added or bookmarked at least one course \\ 
  5,256 schedules \\ 
  17,124 sessions\footnote{A session is defined by a period of activity expiring after 30 minutes of inactivity}, 75\% of which are from returning visitors \\ 
  107 sessions averaged per day \\ 
  5.4 pages viewed averaged per session  \\ 
  6 minutes spent averaged per session \\ 
  \hline
\end{tabular}
\end{center}

\vspace{10pt}

\noindent In this section, I will address many of the features listed in Chapter 2 (Design) and demonstrate their usability.

\subsection{Search}

\begin{figure}[H]
  \centering

  \includegraphics[width=1.0\textwidth]{images/graph/searches}

  \caption{Searches over time, with major relevant calendar events annotated}
  \label{fig:searches}

\end{figure}

Figure \ref{fig:searches} shows user search submissions over the time, and makes the periodicity of Skedge usage extremely apparent. Although it only shows a bit more than one ``phase,'' the cycle of events---\emph{semester begins} (students check course times, room locations, etc. while they get accustomed to the semester), \emph{next semester's course catalog released} (students explore the courses offered and build possible schedules), \emph{next semester registration opens} (cross-checking and officially registering each course with the Registrar)---clearly drives the majority of site traffic.

What this means is that Skedge is not seen as an experiment or a proof of concept that is not production-ready. Rather, its traffic patterns show that it is consistenly \emph{relied upon} by students and used for its intended purpose in times of biggest course scheduling need.

It should also be noted that before March, the majority of the userbase as I knew it consisted of seniors, as the majority of people that I knew personally and shared Skedge with belonged to my class. The latest two major events reported in the graph above thus very impressively consist of \emph{very few seniors}, who have no courses left to schedule. March also marked the very first time I publicly advertised Skedge, although not very strongly. It was sent out to all students in the Computer Science mailing list as well as posted to two class-wide Facebook groups.

The usability of and user satisfaction with a tool is hard to measure when it holds a monopoly. Considering that Skedge is an \emph{optional alternative} to CDCS, such consistent usage data (especially with 75\% of sessions being from returning users) undeniably prove its measures of usability and user satisfaction relative to CDCS.

  \subsubsection{Search types}

  \begin{figure}[H]
    \centering
    \includegraphics[width=1.0\textwidth]{images/graph/searchtypes}

    \caption{Breakdown of search queries by type of search}
    \label{fig:searchtypes}
  \end{figure}

Figure \ref{fig:searchtypes} validates the hypothesis made in Section 2.3 regarding the search interface design, demonstrating that the \emph{department code} and \emph{course number} search modes dominate user searches, accounting for over 85\% of all searches on Skedge. Bringing the nondistinct field for this in CDCS into prominence is thus a design choice that accomodates empirical usage behavior.

\subsubsection{Empty searches}

  Of the total 45,466 searches on Skedge logged since November 2nd 2015, 89.4\% (40,625) of searches produce nonempty search results, whereas 10.6\% (4,841) of searches come up empty.

  This already is evidence of an effective search mechanism, but of those empty searches, the vast majority (92\%) contain typos or are for courses that simply don't exist. The other significant portion of searches include instructor names without the requisite ``taught by'' for SkedgeQL to pick it up as an instructor search, and such a fix could easily be made.

  This type of analysis can powerfully indicate holes in the query language (i.e. searches intuitive for users) that ought to be fixed, which is the same method WolframAlpha describes its team using to improve their freeform input system \cite{wolfram2}. Table \ref{fig:empty-queries} lists a few notable examples of this nature. (Table \ref{fig:funny-queries} lists a few funny queries I came across.)

\subsection{Exports}

In Section 2.2.5 (on the design of Skedge's schedule exporting), a major usability improvement claimed was the ability to export schedules to the {\tt .ics} format, a widely used format for many calendar applications, as well as offering functional Google Calendar export. Figure \ref{fig:exports} shows usage of the export functions over time, demonstrating its consistent and healthy use (traffic surges once again in line with major scheduling events.) This data confirms that there is strong demand for exporting to {\tt .ics}---it is even more common than Google Calendar export.

\subsection{``Link-searches''}

Figure \ref{fig:linksearches} shows the amount of clicks on the Wikipedia-style embedded links throughout Skedge (instructor names link to other courses they are teaching, prerequisite and crosslist course mentions link to those courses). The number of manual instructor searches is also included for comparison, which can be seen to be vastly less usable than the analogous link-click, which doesn't exist in CDCS. Additionally, link-clicks also possibly offer a different functionality (e.g. in the course browsing paradigm).

\singlespacing
{\renewcommand{\arraystretch}{1.5}
\begin{center}
\begin{table}[H]
  \begin{tabular}{ p{6cm} p{7.5cm} }
  \hline

    ``{\tt Prereq 280}'' \newline ``{\tt has-Prereq: 280}''
    & Course dependencies \\ \hline

    ``{\tt tuesday 3:25-4:40}'' \newline ``{\tt m/w}'' \newline ``{\tt classes tuesdays and thursdays between 11:05 and 12:20}'' \newline ``{\tt weekend summer 2016}''
    & \parbox[c]{\hsize}{Better day-of-the-week handling} \\ \hline

    ``{\tt 2 credits natural science}'' 
    & Search by division (although there were very few) \\ \hline

    ``{\tt religion and classics}'' \newline ``{\tt studio art}'' 
    & Search by full department name \\ \hline

    ``{\tt new csc courses}'' 
    & The word ``{\tt courses}'' stimies the query here (somewhat common with advanced query types) \\ \hline

    ``{\tt curriculum change}'' 
    & Alternative for ``{\tt new}'' keyword \\ \hline

    ``{\tt guo}'' 
    & Instructor without ``{\tt taught by}'' (very common) \\ \hline

    ``{\tt stt212 vs mth201}'' \newline ``{\tt stt212 OR mth201}'' \newline ``{\tt csc 2\{4,8\}2}'' \newline ``{\tt csc 242|282}''
    & Searching multiple queries at the same time \\ \hline

    ``{\tt crosslisted csc lin}'' 
    & Probably a more reasonable syntax than the current SkedgeQL one (``{\tt csc x lin}'') \\ \hline

    ``{\tt place: hubbell}'' \newline ``{\tt hubbell friday}'' 
    & Search by location, although motives unclear \\ \hline

    ``{\tt MTH 2*}'' \newline ``{\tt csc 2xx}''
    & Using ``{\tt *}'' or ``{\tt x}'' instead of an omission (e.g. searching ``{\tt MTH 2}'' would work for this purpose) \\ \hline

    \hline
  \end{tabular}
  \caption{Example of search queries that produced no results but that indicate demand for added search features. The corresponding features that currently lack in SkedgeQL are described on the right.}
  \label{fig:empty-queries}
\end{table}

\vspace{30pt}

\begin{table}[H]
  \begin{tabular}{ | p{4cm} | p{4cm} | p{4cm} | }
  \hline

    ``{\tt Larry david}''
    & ``{\tt taught by seligman}''
    & ``{\tt cool stuff}''
    \\ \hline

    ``{\tt game of thrones}''
    & ``{\tt interesting courses}''
    & ``{\tt fun classes in csc}''
    \\ \hline

    ``{\tt easy A}''
    & ``{\tt flute birdman}''
    & ``{\tt how do i make a new schedule}''
    \\ \hline

    \hline
  \end{tabular}
  \caption{Humorous search submissions}
  \label{fig:funny-queries}
\end{table}
\end{center}

\doublespacing

\begin{figure}[H]
  \centering
  \vspace{20pt}
    \begin{subfigure}[w]{14cm}
      \centering
      \includegraphics[width=1.00\textwidth]{images/graph/exports}
      \caption{Different types of schedule exporting over time} \label{fig:exports}
    \end{subfigure} \\
    \vspace{30pt}
    \begin{subfigure}[w]{14cm}
      \centering
      \includegraphics[width=1.00\textwidth]{images/graph/linksearches}
      \caption{Different types of ``link-search'' clicks over time, with manual instructor search included for comparison with clicking on an instructor's name} \label{fig:linksearches}
    \end{subfigure}

  \caption{Other metrics of Skedge usage over time}
\end{figure}

\clearpage

\subsection{Mobile}

Making the case for usability superior to CDCS's in the space of mobile support is easy, given that CDCS does not have a mobile-responsive website. But to illustrate the importance of ensuring good user experience on mobile devices, consder that 11.54\% of all Skedge sessions are from mobile devices, which is very substantial. On average, 2.74 pages viewed are viewed and 2 minutes are spent per session on mobile devices.

The nature of the content also makes Skedge commonly accessed on the go---quickly looking up room numbers while heading to class, for instance, or checking a friend's schedule on Skedge Social to meet them after a class, etc. There is still work to be done to determine the usability of Skedge's mobile site in particular. Surveys could be a good indicator, but I didn't have the chance to administer any.

\subsection{Course blocks}

\begin{figure}[H]
  \centering

  \fbox {
    \includegraphics[width=8cm]{images/skedge/blocks}
  }

  \caption{Etc}
  \label{fig:searchtypes}

\end{figure}

40\% of sessions have at least one block-click
Average of 5.12 block-clicks per session

success because you can't click schedule blocks in better cdcs!!

### 2016-04-13 analytics/usage.tex
Finish 4.1

@@
 
+\subsection{Social}
+
+Since March 1st (40 days), based on 963 sessions on {\tt /social}, Skedge Social has seen\ldots
+
+\vspace{5pt}
+
+\singlespacing
+\begin{center}
+\begin{tabular}{l}
+  \hline
+  141 users have linked Skedge to Facebook \\
+  10,242 page views (256 page views/day)  \\
+  10 pages/session (users cycle through tabs and go back to it) \\
+  a 5\% bounce rate (users interact with it) \\
+  67\% of visits being were returning visitors \\
+  284 overlays onto friends’ schedules (2 per user) \\
+  \hline
+\end{tabular}
+\end{center}
+\doublespacing
+
+\vspace{5pt}
+
+\noindent Although the feature is too young for in-depth data analysis, these initial reports show promise, especially in the rate of adoption (around 3.5 users per day on average), its level of interaction, and the rate of \emph{returning} users to Skedge Social.
+
 \subsection{Course blocks}
 
-\begin{figure}[H]
+\begin{figure}
   \centering
@@
   \fbox {
-    \includegraphics[width=8cm]{images/skedge/blocks}
+    \includegraphics[width=6cm]{images/skedge/blocks}
   }
 
-  \caption{Etc}
+  \caption{Hovering over a course block in Skedge's side-schedule shows at-a-glance information about it; clicking on it navigates the user to that course}
   \label{fig:searchtypes}
@@
 
-40\% of sessions have at least one block-click
-Average of 5.12 block-clicks per session
+As a user adds courses to their schedule, the courses appear as timeblocks on the schedule to the side of the search results.
+The blocks only display the course code (i.e.
+department code, course number, and if it is a subsection such as a lab or workshop), but hovering over them displays more useful information like the course title, location, and the exact start and end times.
 
-success because you can't click schedule blocks in better cdcs!!
+Interestingly, the only course information that is gained by clicking on courses is the description, instructor, and perhaps minor details like course comments, prerequisites, etc.
+Yet, of 12,944 sessions of users with schedules, \textbf{40\%} of those sessions had at least one block-click.
+Those sessions with block clicks averaged \textbf{5.12 block-clicks per session}, a strikingly high number considering its limited functional use!
+How do we explain this?
+
+I hypothesize that a combination of Skedge's presentation of information, its highly interactive user interface, and its concern for quality of user experience contribute to a sense of enjoyment regarding course scheduling.
+Clicking an average of five times \emph{per session} corresponds to clicking on the blocks for each of his or her courses, glancing over the details, and perhaps reading over the course description.
+This enamorment with one's course selection is a large component of Skedge's mission.
+It may seem trivial to state, but I believe the courses one takes in college are paramount to one's personal development and outlook on life.
+I believe ensuring that students are excited about their course selection and can also freely explore what more knowledge their university has to offer them is key to a well-balanced college experience.
+
+The above could only be confirmed with involved user studies, which I do not have any data for.
+At the very least, the average block-clicks per session is a strong case for Skedge's usability over CDCS for the simple reason that Better CDCS does not even allow users to click on their schedule blocks, and this usage data shows that there is clearly a desire to do so.

### 2016-04-13 analytics.tex
Cheeseball

@@
 
-In this chapter I propose three hypotheses. They are:
+In this chapter I propose three hypotheses, each to be supported by analysis in the coming sections. The hypotheses are:
 
@@
 
-  \item Skedge’s \emph{navigations-per-add} demonstrate the \textbf{effectiveness of its search features} for each of the following three user search paradigms:
-
-  \begin{enumerate}
-    \item direct search,
-    \item exploratory search, and
-    \item peer-guided search
-  \end{enumerate}
+  \item Skedge’s \emph{navigations-per-add} demonstrate its \textbf{effectiveness} in the user search paradigms of \textbf{a) direct search} and \textbf{b) exploratory search}.
 
@@
 \end{enumerate}
-
-\noindent In the following sections I will present my analysis on collected usage data since November 3rd 2015, which will serve to validate each hypothesis.
 

### 2016-04-13 analytics/usage.tex
Cheeseball

@@
 
-From November 3, 2015 to April 11, 2016 (a period of 159 days), Skedge has seen\ldots
+In this section, I will address many of the features listed in Chapter 2 (Design) and demonstrate their usability. As an overview, since November 3, 2015 (159 days ago), Skedge has seen\ldots
+
+\vspace{5pt}
 
@@
   4,713 users that have added or bookmarked at least one course \\ 
-  5,256 schedules \\ 
-  17,124 sessions\footnote{A session is defined by a period of activity expiring after 30 minutes of inactivity}, 75\% of which are from returning visitors \\ 
+  5,256 schedules, with an average of 5.13 and median of 5 sections per schedule \\ 
+  17,124 sessions, 75\% of which are from returning visitors \\ 
   107 sessions averaged per day \\ 
@@
 
-\vspace{10pt}
+The figure of 4,713 total users may be a little misleading,
 
-\noindent In this section, I will address many of the features listed in Chapter 2 (Design) and demonstrate their usability.
+\begin{itemize}
+  \item 
+\end{itemize}
 
@@
 
-I hypothesize that a combination of Skedge's presentation of information, its highly interactive user interface, and its concern for quality of user experience contribute to a sense of enjoyment regarding course scheduling.
-Clicking an average of five times \emph{per session} corresponds to clicking on the blocks for each of his or her courses, glancing over the details, and perhaps reading over the course description.
-This enamorment with one's course selection is a large component of Skedge's mission.
-It may seem trivial to state, but I believe the courses one takes in college are paramount to one's personal development and outlook on life.
-I believe ensuring that students are excited about their course selection and can also freely explore what more knowledge their university has to offer them is key to a well-balanced college experience.
+I hypothesize that a combination of Skedge's presentation of information, highly interactive user interface, and concern for quality of user experience contributes to a sense of enjoyment in course scheduling.
+Five clicks corresponds to five courses (both the median and mean for sections per schedule is 5), meaning that in every session, a user could be reviewing details and reading over descriptions for each of their courses.
+Enabling this sense of pride and admiration with one's course selection is a crucial aspect of Skedge's mission.
+I believe the courses one chooses to take are an integral aspect of one's personal development and outlook on life in college.
+I believe ensuring that students are excited not only about their courses, but also about \emph{exploring} what their university has to offer is key to an inspired, intellectually curious student body.
 
-The above could only be confirmed with involved user studies, which I do not have any data for.
+The above could only be confirmed with involved user studies, for which I have no data.
 At the very least, the average block-clicks per session is a strong case for Skedge's usability over CDCS for the simple reason that Better CDCS does not even allow users to click on their schedule blocks, and this usage data shows that there is clearly a desire to do so.

### 2016-04-13 conclusion.tex
Cheeseball

@@
 
-While I am aware that such matters are not extremely pressing, they nevertheless degrade student experience on campus and discolor perception of the administration's care for students, especially when modern, relevant, and engaging tools are within arm's reach, and \emph{most} especially at a time when University of Rochester's very high tuition rates continue to rise year-to-year.
+While functional, CDCS nevertheless degrades student experience on campus and discolors perception of the administration's care for students, especially when modern, relevant, and engaging tools are within arm's reach, and most especially at a time when University of Rochester's very high tuition rates continue to rise year-to-year.
 
@@
 For the sake of the University's students, parents, faculty, and staff, I hope that my efforts with Skedge were not in vain and can lead to a substantial refresh of University tools that, most crucially, \textbf{involves a dialogue between the administration, students, and staff}.
-Thank you for reading.
+Thank you.
 

### 2016-04-13 analytics/npa.tex
Finish 4.1
Cheeseball
navs per adds

@@
-\section{Effectiveness of search}
+\section{Effectiveness as a platform \& navigations-per-add}
 
@@
 
-A navigation is defined as
-a search, or
-a click on an instructor’s name, or
-a click on a crosslisted or prerequisite course link
+A \textbf{navigation} is defined as
 
-The navigations-per-add measure is
-the number of navigations a user took (within one session) until a course was added, bookmarked
+\onehalfspacing
+\begin{itemize}
+  \item a search, or
+  \item a click on an instructor’s name, or
+  \item a click on a crosslisted or prerequisite course link
+\end{itemize}
+\doublespacing
+
+\noindent The \textbf{navigations-per-add} metric is the number of navigations taken by a user (within one session) before a section was added.
+
+To illustrate this metric, imagine that a user searches for ``{\tt csc}'' (one navigation so far) and begins to scroll through the list of courses.
+The user sees that Philip Guo is teaching CSC 210 and clicks on his name (two navigations so far), and then adds the section for CSC 210.
+\emph{A navigations-per-add of 2 is tallied for that user}.
+The user then clicks on the prerequisite for that class, CSC 172 (one navigation so far), and adds it.
+\emph{A navigations-per-add of 1 is tallied for that user}.
+The user then opens the list of labs embedded in 172 and adds one of them.
+\emph{A navigations-per-add of 0 is tallied for that user.} (Note that a value of 0 can also occur when a user scrolls through a list of courses and adds more than one of them without leaving the page.)
 
@@
 
-  \caption{Etc}
-  \label{fig:searchtypes}
+  \caption{Frequency of different navigations-per-add values}
+  \label{fig:navs-combined}
 \end{figure}
+
+Figure \ref{fig:navs-combined} takes the navigations-per-add for every time a course was ever added by any Skedge user.
 
@@
     \centering
-    \includegraphics[width=1.0\textwidth]{images/graph/browse_navs}
+    \includegraphics[width=1.0\textwidth]{images/graph/browsed_navs}
 
@@
   As expected, the 0-navs were mostly maincourses (62\%)
-
-  Effective++
-
-  \subsubsection{Social}
-
-  90 users have linked Skedge to Facebook
-  Since March 1st,
-  4,000+ visits (200 visits/day)
-  ~60\% of visits to /social were returning visitors
-  90 overlays onto friends’ schedules
-  10 clicks to Facebook profiles :(
-  - get stats from the fb dashboard
-
-  success?????

### 2016-04-13 design/modernity.tex
random

@@
 With the increasing number of attendees at University of Rochester's hackathons, it is clear that the University's ``hacker culture'' is growing---more students are collaborating to build side-projects that integrate resources and servies often benefitting the student community.
-Open-source and open-information services greatly help to foster such innovation, and having public APIs is essential toward this end.\cite{Milberry}\cite{hackers}
+Open-source and open-information services greatly help to foster such innovation, and having public APIs is essential toward this end \cite{Milberry}\cite{hackers}.
 
 Skedge provides a public JSON API at the root URL \url{http://www.skedgeur.com/api/}, made at the request of a student that was interested in using its course data, and the API has already been used in projects by several other student groups.
-The endpoints included are \url{/api/courses?q=query} (Skedge's query language---described in detail in section 2.3---is supported here), \url{/api/departments?q=optional_query}, and \url{/api/instructors?q=optional_query}.
+The endpoints included as of now are \url{/api/courses?q=query} (Skedge's query language---described in detail in section 2.3---is supported here) and \url{/api/departments} for a list of department names and codes.
 

### 2016-04-13 design/usability.tex
random

@@
 For instance, the search {\tt csc} for Fall 2016 on CDCS results in 147 ``section boxes'', while Skedge only shows 45 ``\emph{course} boxes'', with subsections collapsed within their respective course.
-This triage reduces the data (noise, more correctly) displayed by ~70\%, and is even higher for departments with more abundant labs and workshops, such as Physics (Skedge: 35 vs.
-CDCS: 226, an ~85\% reduction), or Chemistry (Skedge: 25 vs.
-CDCS: 171, an ~86\% reduction).
+This triage reduces the data (noise, more correctly) displayed by \textbf{$\sim$70\%}, and is even higher for departments with more abundant labs and workshops, such as Physics (Skedge: 35 vs.
+CDCS: 226, \textbf{an $\sim$85\% reduction}), or Chemistry (Skedge: 25 vs.
+CDCS: 171, \textbf{an $\sim$86\% reduction}).
 

### 2016-04-13 analytics/usage.tex
random
Fix analytics intro

@@
 
-In this section, I will address many of the features listed in Chapter 2 (Design) and demonstrate their usability. As an overview, since November 3, 2015 (159 days ago), Skedge has seen\ldots
+In this section, I will address many of the features listed in Chapter 2 (Design) and demonstrate their usability. To date, Skedge has seen\ldots
 
@@
   5,256 schedules, with an average of 5.13 and median of 5 sections per schedule \\ 
-  17,124 sessions, 75\% of which are from returning visitors \\ 
+  75\% of sessions from returning visitors \\ 
   107 sessions averaged per day \\ 
@@
 
-The figure of 4,713 total users may be a little misleading,
+\noindent The figure of 4,713 total users may be misleading because, as explained in Chapter 3, if a user resets their browser cookies or changes browsers or devices, this registers as a distinct visitor.
+Unfortunately, even analytics tools that track unique visitors rely on cookies as well.
 
-\begin{itemize}
-  \item 
-\end{itemize}
+In an attempt to exclude ``stale users,'' we can time-constrain our analysis.
+In the past month, there were 1,416 visitors to the site and 672 visitors \emph{with schedules}.
+However, this metric excludes data from seniors who did not have need for Skedge in the past month.
+Going back to November 3, 2015, course registration day ($\sim$5 months ago), we find 4,647 visitors to the site and 1322 visitors with schedules.
+Thus, we can estimate that the amount of users who have added courses with Skedge is about \textbf{one-fifth} of all undergraduates (6,046) at the College.
 
@@
 
-  \caption{Searches over time, with major relevant calendar events annotated}
+  \caption{Searches over the period Oct 19 2015 to Apr 11 2016, annotated with major events}
   \label{fig:searches}
@@
 
-Figure \ref{fig:searches} shows user search submissions over the time, and makes the periodicity of Skedge usage extremely apparent.
-Although it only shows a bit more than one ``phase,'' the cycle of events---\emph{semester begins} (students check course times, room locations, etc.
+Figure \ref{fig:searches} shows searches over time, and makes the periodicity of Skedge usage extremely apparent.
+Although there is only enough data for a bit more than one ``phase,'' the cycle of events---\emph{semester begins} (students check course times, room locations, etc.
 while they get accustomed to the semester), \emph{next semester's course catalog released} (students explore the courses offered and build possible schedules), \emph{next semester registration opens} (cross-checking and officially registering each course with the Registrar)---clearly drives the majority of site traffic.
@@
 
-  This already is evidence of an effective search mechanism, but of those empty searches, the vast majority (92\%) contain typos or are for courses that simply don't exist.
-The other significant portion of searches include instructor names without the requisite ``taught by'' for SkedgeQL to pick it up as an instructor search, and such a fix could easily be made.
+  This already is evidence of an effective search mechanism, but of those empty searches, the vast majority contain typos or are for courses that simply don't exist.
+The other significant portion of searches include instructor names without the requisite ``taught by'' for SkedgeQL to pick it up as an instructor search, and this issue could easily be fixed.
 

### 2016-04-13 analytics/usage.tex
Update empty search tables

@@
 Figure \ref{fig:searchtypes} validates the hypothesis made in Section 2.3 regarding the search interface design, demonstrating that the \emph{department code} and \emph{course number} search modes dominate user searches, accounting for over 85\% of all searches on Skedge.
-Bringing the nondistinct field for this in CDCS into prominence is thus a design choice that accomodates empirical usage behavior.
+Bringing the nondistinct field for this in CDCS into prominence is thus a design choice that accomodates empirical behavior.
 
@@
 
-  This type of analysis can powerfully indicate holes in the query language (i.e.
-searches intuitive for users) that ought to be fixed, which is the same method WolframAlpha describes its team using to improve their freeform input system \cite{wolfram2}.
+  This type of analysis is powerful, as it can raise issues in the query language (i.e.
+searches I didn't implement but that users expect to work) with specific fixes.
+This is the same method WolframAlpha describes its team using to improve their freeform input system \cite{wolfram2}.
 Table \ref{fig:empty-queries} lists a few notable examples of this nature.
@@
 \singlespacing
-{\renewcommand{\arraystretch}{1.5}
+{\renewcommand{\arraystretch}{1.75}
 \begin{center}
 \begin{table}[H]
+\vspace{5pt}
   \begin{tabular}{ p{6cm} p{7.5cm} }
+
+    {\large Empty search examples}
+    & {\large Specific fix to SkedgeQL required} \\
+
+    \hline
   \hline
@@
 
-    \hline
   \end{tabular}
@@
 
-\vspace{30pt}
+\vspace{5pt}
 
 \begin{table}[H]
-  \begin{tabular}{ | p{4cm} | p{4cm} | p{4cm} | }
+  \centering
+  \begin{tabular}{ l l l }
+
     \hline
 
-    ``{\tt Larry david}''
+    ``{\tt why 123}''
     & ``{\tt taught by seligman}''
     & ``{\tt cool stuff}''
-    \\ \hline
+    \\
 
-    ``{\tt game of thrones}''
+    ``{\tt psy cool professor}''
     & ``{\tt interesting courses}''
     & ``{\tt fun classes in csc}''
-    \\ \hline
+    \\
 
@@
     & ``{\tt how do i make a new schedule}''
-    \\ \hline
+    \\
 
     \hline
+
   \end{tabular}
-  \caption{Humorous search submissions}
+  \vspace{10pt}
+  \caption{A sample of humorous searches}
   \label{fig:funny-queries}

### 2016-04-13 analytics/npa.tex
random
half of nps

@@
 
-\subsection{Trends}
+\subsection{General trends}
 
@@
 
-Figure \ref{fig:navs-combined} takes the navigations-per-add for every time a course was ever added by any Skedge user.
+Figure \ref{fig:navs-combined} shows the navigations-per-add breakdown for every time a section was added in Skedge.
+With 40\% of sections added in zero navigations and 39\% added within one navigation, it is immediately apparent that Skedge delivers users the content they look for in the vast majority of cases, assuming that the intention of the user was to add a course.
 
-\subsection{Breaking them apart}
+Considering the hypothesis claimed in Section 2.3.2 (that course selection criteria can be broken down to \emph{requirements}, \emph{electives}, and \emph{peer recommendations}), however, we ought examine if data for different use-cases are being conflated in this chart.
 
-  behavioral patterns
-  Direct search for specific course
-  Discovery, browsing, exploring
-
-  \subsubsection{Direct searches}
+\subsection{Direct search (requirement criteria)}
 
@@
 
-    \caption{Etc}
-    \label{fig:searchtypes}
+    \caption{Frequency of navigations-per-adds when the goal was to add a specific course}
+    \label{fig:navs-direct}
   \end{figure}
 
-  Why would 0-navs be so common with direct searches? 64\% of them subsections!
+  Direct searches are easy to distinguish, as we can take the subset of search queries that target a specific course, namely, those that \textbf{a)} contain both department and course code fields (e.g.
+``{\tt csc 171}''), or \textbf{b)} contain a title field (``{\tt intro to programming}'').
+Figure \ref{fig:navs-direct} shows the navigations-per-add when, under this criteria, the user's goal was to add a specific course.
 
-  \subsubsection{Browse}
+  The difference between these results and those in \ref{fig:navs-combined} is suble but logical.
+Users searching for a specific course should have an increased proportion of 1-navigation adds, as the graph shows, because the use-case is \emph{search for course}, \emph{add course}, perhaps occasionally taking an extra try to account for typos, not knowing the title or course code exactly, etc.
+
+  But given this reasoning, why would 0-navigation adds still be so common in direct searches?
+Further analysis on only the sections that were added from direct search and with 0 navigations shows that 64\% of those sections are in fact \emph{subsections}.
+As happened in the illustrative example at the beginning of this section, this shows that a very common use-case is to add labs, workshops, etc.
+of a course directly after adding the main section.
+
+  This hypothesis is further confirmed by doing the same analysis for ${navs}=1$ and ${navs}=2$, when the inverse behavior would be expected (it makes sense for subsections which were added seperately from their main sections to be more rate).
+Only 11\% of sections that were added from a direct search after 1 navigation are subsections, and that number drops to 8\% for 2 navigations.
+
+\subsection{Browse (elective criteria)}
+
+  Browsing behavior was done as a direct inverse of  harder to detect than direct search behavior, because there  query behavioral patterns
+  Discovery, browsing, exploring
 
@@
 
-    \caption{Etc}
+    \caption{Frequency of navigations-per-adds/bookmarks when the goal was to browse for courses}
     \label{fig:searchtypes}
@@
   As expected, the 0-navs were mostly maincourses (62\%)
+
+\subsection{Peer-guided (recommendation criteria)}
+
+  :(

### 2016-04-13 paper.tex
bib stuff
half of nps

@@
 As a result, Skedge boasts strong user-retention rates, long session durations, and high student adoption despite having virtually no advertisement.
-Through collected usage data, I demonstrate that \textbf{a)} Skedge's differences from and additions to CDCS are usable and have real-world need, \textbf{b)} the three major use-cases associated with course scheduling---direct search, exploratory search, and peer recommendation---are effectively accommodated by Skedge, and \textbf{c)} Skedge's novel search mechanism is user-friendly and self-teaches to users over time.
+Through collected usage data, I demonstrate that \textbf{a)} Skedge's differences from and additions to CDCS are usable and have real-world need, \textbf{b)} the two major use-cases associated with course scheduling---direct and exploratory search---are effectively accommodated by Skedge, and \textbf{c)} Skedge's novel search mechanism is user-friendly and self-teaches to users over time.
 
@@
 
-\addcontentsline{toc}{chapter}{Bibliography}
 \singlespacing
@@
 
+\addcontentsline{toc}{chapter}{Bibliography}
+
 \clearpage

### 2016-04-13 analytics/npa.tex
Cite undergrads

@@
 
-    \caption{Frequency of navigations-per-adds when the goal was to add a specific course}
+    \caption{Frequency of navigations-per-adds when the goal was to \emph{add a specific course}}
     \label{fig:navs-direct}
@@
 
-  Browsing behavior was done as a direct inverse of  harder to detect than direct search behavior, because there  query behavioral patterns
-  Discovery, browsing, exploring
-
   \begin{figure}
@@
 
-    \caption{Frequency of navigations-per-adds/bookmarks when the goal was to browse for courses}
-    \label{fig:searchtypes}
+    \caption{Frequency of navigations-per-add/bookmark when the goal was to \emph{browse for courses}}
+    \label{fig:navs-browse}
   \end{figure}
 
-  As expected, the 0-navs were mostly maincourses (62\%)
+  Browsing behavior was done as a direct inverse of direct searches, detecting searches that didn't specifically target any courses.
+This time, bookmarks were also counted in a user's flow of events (i.e.
+they are treated the same as an add), because bookmarking is also a goal of course browsing behavior.
+
+  The results are shown in Figure \ref{fig:navs-browse}.
+The difference between these results and the initial combined results of \ref{fig:navs-combined} are more substantial than last time, and may defy expectations at first glance.
+While it may seem intuitive that browsed-for courses should have higher amounts of navigations, note that the graph itself is showing \emph{effectiveness} of such browsing.
+In this case, it makes sense that a significantly higher proportion of adds/bookmarks occur within zero navigations, because that indicates a high-level search and subsequent adds/bookmarks from that list instead of moving on to another search.
+
+  As expected, and unlike the results of further direct-search analysis, the sections added/bookmarked when browsing within 0 navigations were mostly \emph{main} sections (only 37\% of them were subsections).
+This makes sense, as users may not be concerned with subsections if the goal of the browse is merely to mark interesting courses.
+This criterion saw more drastic drops in subsection amounts as well, as the number of navs increased: 5\% subsections for ${navs}=1$, and 4\% subsections for ${navs}=2$.
 

### 2016-04-13 analytics/usage.tex
linksearches graph
half of nps
Cite undergrads

@@
 Going back to November 3, 2015, course registration day ($\sim$5 months ago), we find 4,647 visitors to the site and 1322 visitors with schedules.
-Thus, we can estimate that the amount of users who have added courses with Skedge is about \textbf{one-fifth} of all undergraduates (6,046) at the College.
+Thus, we can estimate that roughly \textbf{one-fifth} of all undergraduates (6,046 students) at the College have added courses on Skedge \cite{undergrads}.
 
@@
   \vspace{10pt}
-  \caption{A sample of humorous searches}
+  \caption{A sample of some funny searches}
   \label{fig:funny-queries}
@@
   \centering
-  \vspace{20pt}
-    \begin{subfigure}[w]{14cm}
+    \begin{subfigure}[w]{14.5cm}
       \centering
@@
     \end{subfigure} \\
+
     \vspace{30pt}
-    \begin{subfigure}[w]{14cm}
+
+    \begin{subfigure}[w]{14.5cm}
       \centering
@@
 
-\vspace{5pt}
-
+{\renewcommand{\arraystretch}{1.5}
 \singlespacing

### 2016-04-13 design/usability.tex
Cite undergrads

@@
 
-Better CDCS provides two ways for users to export their schedule---as image or to Google calendar--except that the Google Calendar export feature is completely non-functional, telling users that ``{\tt This export tool will only work properly for the Spring of 2015}'' only \emph{after} they authorized the tool to access their Google account.
+Better CDCS provides two ways for users to export their schedule---as an image or to Google calendar---and the Google Calendar export feature is completely non-functional, telling users that ``{\tt This export tool will only work properly for the Spring of 2015}'' after the user authorizes the tool to access their Google account on a separate University page \cite{cdcs-google}.
 

### 2016-04-13 analytics.tex
npa done

@@
 
-  \item Skedge’s \emph{navigations-per-add} demonstrate its \textbf{effectiveness} in the user search paradigms of \textbf{a) direct search} and \textbf{b) exploratory search}.
+  \item Skedge’s \emph{navigations-per-add} demonstrate its \textbf{effectiveness} in the user search paradigms of \textbf{a) direct search}, \textbf{b) exploratory search}, and \textbf{c) peer-guided search}.
 

### 2016-04-13 analytics/npa.tex
npa done

@@
 
-  :(
+  Given more time with Skedge Social being live and more forethought in tracking its usage, similar data analysis could be done on the subject of the effectiveness of Skedge in peer-guided course finding.
+Unfortunately, usage data to answer questions such as ``how many Facebook friends were listed as liking or taking a course when it was added?'', and cannot be recreated due to Skedge's purposeful privacy limitations (Facebook friend data can only be retrieved client-side---no data is stored on the server that could authorize such a data request).
+Thus, part (c) of hypothesis \#2 cannot be verified.

### 2016-04-13 paper.tex
npa done

@@
 As a result, Skedge boasts strong user-retention rates, long session durations, and high student adoption despite having virtually no advertisement.
-Through collected usage data, I demonstrate that \textbf{a)} Skedge's differences from and additions to CDCS are usable and have real-world need, \textbf{b)} the two major use-cases associated with course scheduling---direct and exploratory search---are effectively accommodated by Skedge, and \textbf{c)} Skedge's novel search mechanism is user-friendly and self-teaches to users over time.
+Through collected usage data, I demonstrate that \textbf{a)} Skedge's differences from and additions to CDCS are usable and have real-world need, \textbf{b)} the three major use-cases associated with course scheduling---direct, exploratory, and peer-guided search---are effectively accommodated by Skedge, and \textbf{c)} Skedge's novel search mechanism is user-friendly and self-teaches to users over time.
 

### 2016-04-13 analytics/npa.tex
little updates

--- 2016-04-13 analytics/npa.tex
@@
 
-    \caption{Frequency of navigations-per-adds when the goal was to \emph{add a specific course}}
+    \caption{Frequency of navigations-per-adds for the goal of \emph{adding a specific course}}
     \label{fig:navs-direct}
@@
 
-    \caption{Frequency of navigations-per-add/bookmark when the goal was to \emph{browse for courses}}
+    \caption{Frequency of navigations-per-add/bookmark for the goal of \emph{browsing for courses}}
     \label{fig:navs-browse}
@@
 
-  :(
+  Given more time with Skedge Social being live and more forethought in tracking its usage, similar data analysis could be done on the subject of the effectiveness of Skedge in peer-guided course finding.
+Unfortunately, usage data to answer questions such as ``how many Facebook friends were listed as liking or taking a course when it was added?'', and cannot be retroactively computed due to Skedge's purposeful privacy limitations (Facebook friend data can only be retrieved client-side---no data is stored on the server that could authorize such a data request).
+Thus, part (c) of hypothesis \#2 cannot be verified.

### 2016-04-13 analytics/usage.tex
little updates

@@
 
-In this section, I will address many of the features listed in Chapter 2 (Design) and demonstrate their usability. To date, Skedge has seen\ldots
+In this section, I will adress the usability of the features listed in Chapter 2 (Design). But first, here are some general statistics of Skedge to date:
 
-\vspace{5pt}
+\vspace{7pt}
 
+\onehalfspacing
 \begin{center}
@@
 \end{center}
+\doublespacing
 
@@
 
-  \caption{Searches over the period Oct 19 2015 to Apr 11 2016, annotated with major events}
+  \caption[Searches over the period Oct 19 2015 to Apr 11 2016]{Searches over the period Oct 19 2015 to Apr 11 2016, annotated with major events}
   \label{fig:searches}
@@
   \hline
-  141 users have linked Skedge to Facebook \\
+  141 users have linked Skedge to Facebook ($\sim$3.5 per day) \\
   10,242 page views (256 page views/day)  \\
-  10 pages/session (users cycle through tabs and go back to it) \\
-  a 5\% bounce rate (users interact with it) \\
+  10 pages/session (users cycle through its tabs and repeatedly return to it) \\
+  a 5\% bounce rate (users interact with the page before leaving) \\
   67\% of visits being were returning visitors \\
@@
 
-\noindent Although the feature is too young for in-depth data analysis, these initial reports show promise, especially in the rate of adoption (around 3.5 users per day on average), its level of interaction, and the rate of \emph{returning} users to Skedge Social.
+\noindent Although the feature is too young for in-depth data analysis, these initial reports show promise, especially in the rate of adoption, its level of interaction, and the rate of returning users to Skedge Social.
 
@@
 
-  \caption{Hovering over a course block in Skedge's side-schedule shows at-a-glance information about it; clicking on it navigates the user to that course}
+  \caption[Course blocks in Skedge]{Hovering over a course block in Skedge's side-schedule shows at-a-glance information about it; clicking on it navigates the user to that course}
   \label{fig:searchtypes}

### 2016-04-13 design/social.tex
little updates

@@
       }
-      \caption{Skedge Social landing page (\url{http://skedgeur.com/social}) for a user whose account has not yet been integrated with Facebook.
+      \caption[Skedge Social landing page for a user who hasn't yet joined it]{Skedge Social landing page (\url{http://skedgeur.com/social}) for a user who hasn't yet joined it.
 Skedge Social appears in the context of the ordinary Skedge interface, and is accessible with the globe icon to the right of the searchbar.} \label{fig:sk-social-login}
@@
       }
-      \caption{Skedge Social's schedule feed (the default of four tabs) for a user who has integrated their Skedge account with Facebook.
+      \caption[Skedge Social's schedule feed]{Skedge Social's schedule feed (the default of four tabs) for a user who has integrated their Skedge account with Facebook.
 The user can toggle between a peer's schedule and the list of their \emph{liked} courses.} \label{fig:sk-feed}
@@
     \end{tabular}
-    \caption{Notifications: none, one, or both of them can adorn the Skedge Social logo}
+    \caption[Skedge Social's notifications]{Notifications: none, one, or both of them can adorn the Skedge Social logo}
   \end{figure}

### 2016-04-13 paper.tex
little updates

@@
 \title{
-\vspace{50pt}\\
-\huge \bfseries SKEDGE
-\\
+\vspace{60pt}\\
+\huge \bfseries Skedge \\
 \vspace{10pt}
 \Large
-Smarter course scheduling for our\\
-University of Rochester
+Smarter course scheduling for our University
 }
@@
 \date{April 13, 2016\\
-    \vspace{100pt}\\
+    \vspace{80pt}\\
     submitted in partial fulfillment of\\
     the requirements for the degree\\
-    \emph{honors bachelor of science}\\
-    \vspace{30pt}\\
+    \emph{honors bachelor of science},\\
+    \vspace{40pt}\\
     and as an open letter addressed to\\
@@
     at the university of rochester
-    \vspace{-30pt}
+    \vspace{-80pt}
 }

### 2016-04-13 analytics/dsl.tex
progress with dsl

@@
 
-Points for search by (omits number and dept.):
+The \textbf{complexity} of a search query is the number of fields that aren't course number or department that it spans (the accepted fields are shown in Table \ref{fig:sc-fields}).
+
+\singlespacing
+\begin{center}
+\begin{table}[H]
+  \centering
+  \begin{tabular}{ lllll }
+
+    \hline
 
     description
-credits
-crosslisted
-CRN
+    & title
+    & credits
+    & crosslisted
+    & CRN
+    \\
+
     instructor
-title
-year
-term
-‘random’
-upper-level writing
+    & year
+    & term
+    & upper-level writing
+    & random
+    \\
 
-“CSC” → 0
-“MTH 165” → 0
-“taught by hema” → 1   ✓    (2 searches) 
-“random mur 1-2 credits” → 2   ✓    (1 search)
+    \hline
+
+  \end{tabular}
+  \vspace{10pt}
+  \caption{``Complex'' search fields (omits only number and department)}
+  \label{fig:sc-fields}
+\end{table}
+\end{center}
+\doublespacing
+
+\noindent We can compute the number of searches it takes a user to reach a new highest search complexity record since the previous record. See Table \ref{fig:sc-example} for an illustrative example.
+
+A user's \textbf{first increase in search complexity} is defined as the number of searches a user made before searching with complexity greater than zero.
+If a user's first search has complexity 1, the user's \emph{first increase} is 0.
+
+A user's \textbf{second increase in search complexity} is defined as the number of searches a user made before searching with complexity greater than their \emph{first increase}'s search complexity.
+
+It is possible that user have neither, both, or only a \emph{first increase} metric.
+
+{\renewcommand{\arraystretch}{2.0}
+\singlespacing
+\begin{center}
+\begin{table}[H]
+  \centering
+  \begin{tabular}{ r | c c c | l }
+
+    Query no.
+    & Search query
+    &
+    & Complexity
+    & Searches till increase \\
+
+    \hline
+
+    \#1
+    & ``{\tt CSC}''
+    & \rightarrow
+    & 0
+    & 
+    \\
+
+    \#2
+    & ``{\tt MTH 165}''
+    & \rightarrow
+    & 0
+    & 
+    \\
+
+    \#3
+    & ``{\tt csc taught by guo}''
+    & \rightarrow
+    & 1
+    & $3-1=\textbf{2}$ \hspace{2pt} \emph{(first increase)}
+    \\
+
+    \#4
+    & ``{\tt random mur 1-2 credits}''
+    & \rightarrow
+    & 2
+    & $4-3=\textbf{1}$ \hspace{2pt} \emph{(second increase)}
+    \\
+
+  \end{tabular}
+  \vspace{10pt}
+  \caption{Illustration of computing the number of searches before a complexity increase}
+  \label{fig:sc-example}
+\end{table}
+\end{center}
+\doublespacing
+
+\subsection{Motivation}
+
+The purpose of this metric is to determine the usability and learning curve of SkedgeQL. The two ways Skedge currently offers to teach users about its query language are
+
+\begin{enumerate}
+  \item  by providing a table of example queries whenever the user clicks inside the search field from the Skedge homepage (see Figure \ref{fig:sk-search2}), and
+
+  \item  from embedded links on the site making searches with high complexity and implicitly demonstrating them to users by using the same input field they use to submit searches.\footnote{Clicking on an instructor's name, for instance, makes the advanced search ``for the user.'' Another example is clicking on a course box in the side schedule, which searches for that course code along with its term and year (e.g.
+``{\tt CSC 171 Fall 2016}'').
+
+  Other possibilities include linking the credits field of each course to a credits search, the CRN field of a section to a CRN search, the ``{\tt W}'' part of the course code to an upper-level writing search, having a crosslisted course link search the crosslisting of the two departments (instead of directly link to the crosslisted course), a ``feeling lucky?'' button for a search with ``{\tt random}''.}
+\end{enumerate}
+
+\begin{figure}[H]
+  \centering
+  \fbox{
+    \includegraphics[width=0.65\textwidth]{images/skedge/search}
+  }
+  \caption{Examples of SkedgeQL} \label{fig:sk-search2}
+\end{figure}
+
+\noindent The latter method is, of course, much preferred to the former as it teaches without requiring concerted effort on the part of the user to study and learn the query language, and mostly likely remains more memorable and intuitive as a result.
+
+A \emph{high} percentage of users with \emph{first increase} and \emph{second increase} metrics, with \emph{low} mean and median values for those increases would be desirable.
+This would strongly indicate the effect of teaching method \#2 because the factor in teaching users about the query language would simply be time (besides the case of users becoming curious about advanced search types and looking them up explicitly, which is difficult to account for because the query example box is shown to all users fairly regularly).
 
@@
   \centering
+  \vspace{-10pt}
 
+  \begin{subfigure}[w]{0.85\textwidth}
     \includegraphics[width=1.00\textwidth]{images/graph/search_dt}
+
+    \caption{Etc}
+    \label{fig:sc-graph}
+  \end{subfigure}
 
@@
 
-  \includegraphics[width=0.65\textwidth]{images/table/search_dt}
+  \begin{subfigure}[w]{0.75\textwidth}
+    \includegraphics[width=1.00\textwidth]{images/graph/search_dt_pie}
 
     \caption{Etc}
-  \label{fig:searchtypes}
+    \label{fig:sc-pie}
+  \end{subfigure}
+
+  \caption{Etc}
 \end{figure}
+
+The blue line in Figure \ref{fig:sc-graph} plots the number of users
 

### 2016-04-13 analytics/dsl.tex
chapter 4 done

@@
 
-\begin{figure}
+\subsubsection{Analysis of users with search complexity increases}
+
+\begin{figure}[H]
   \centering
-  \vspace{-10pt}
+  \includegraphics[width=0.9\textwidth]{images/graph/search_dt}
 
-  \begin{subfigure}[w]{0.85\textwidth}
-    \includegraphics[width=1.00\textwidth]{images/graph/search_dt}
-
-    \caption{Etc}
+  \caption{User-tracked search complexity over time}
   \label{fig:sc-graph}
-  \end{subfigure}
-
-  \\
-  \vspace{30pt}
-
-  \begin{subfigure}[w]{0.75\textwidth}
-    \includegraphics[width=1.00\textwidth]{images/graph/search_dt_pie}
-
-    \caption{Etc}
-    \label{fig:sc-pie}
-  \end{subfigure}
-
-  \caption{Etc}
 \end{figure}
 
-The blue line in Figure \ref{fig:sc-graph} plots the number of users
+The blue line in Figure \ref{fig:sc-graph} plots the number of users that had their \emph{first increase} in search complexity per search index (searches \#16--\#25 represent 47 users, and 26+ searches represents only 29 users).
+The green line plots the number of users for a \emph{second increase} at the given search index (searches \#16--\#25 represent 35 users, and 26+ searches represents 32).
 
-First increase 69.4\% of users), with a median of 2 searches.
-Note that starting at search complexity one or greater counts as a ``searches before first increase value'' of 0.
+The amount of users whose first searches are of nonzero complexity may be surprising.
+The vast majority of these are \textbf{title} searches however, which \emph{are} included in the ``complex'' field table.
+This is purposeful---title searches only represent 8\% of all search field uses, as we saw in Figure \ref{fig:searchtypes} of the section before last.
+Moreover, a user's understanding and/or intuition that the Skedge search field is free-form and can be used for different search fields simultaneously is already a significant conceptualization of SkedgeQL, albeit a relatively basic field.
 
-Second increase (7.9\% of users)
-Median: 8 searches
-Average: 17.52 searches
+\subsubsection{Users with increases vs. users without increases}
 
-DSQL++
+\begin{figure}[H]
+  \centering
+  \includegraphics[width=0.9\textwidth]{images/graph/search_dt_pie}
+
+  \caption{Users with a schedule and more than one search split by their number of increases in search complexity}
+  \label{fig:sc-pie}
+\end{figure}
+
+\vspace{5pt}
+
+Figure \ref{fig:sc-pie} shows that 69.4\% of Skedge users (\emph{users} here defined as having at least one course bookmarked or added to a schedule, with more than two searches since November 3rd---there are 1889 such users in total) have increased their record search complexity at least once, with a median of 2 searches to do so.
+Of those users, 18\% of them (12\% of total users) have had a \emph{second} increase in search complexity, with a median of 10 searches to do so.
+30\% of users have only searched with the department and course number fields.
+
+I believe this is a strong model for evaluating the friendliness and low learning curve of SkedgeQL, and the results are certainly promising, although further analysis on beginning with nonzero search complexity is needed.

### 2016-04-13 analytics/usage.tex
chapter 4 done

@@
 In an attempt to exclude ``stale users,'' we can time-constrain our analysis.
-In the past month, there were 1,416 visitors to the site and 672 visitors \emph{with schedules}.
-However, this metric excludes data from seniors who did not have need for Skedge in the past month.
-Going back to November 3, 2015, course registration day ($\sim$5 months ago), we find 4,647 visitors to the site and 1322 visitors with schedules.
-Thus, we can estimate that roughly \textbf{one-fifth} of all undergraduates (6,046 students) at the College have added courses on Skedge \cite{undergrads}.
+In the past month, there were 1,416 visitors (unique) to the site and 1,136 visitors \emph{with schedules}.
+However, this metric excludes data from seniors who did not have a need for Skedge in the past month.
+Going back since November 3, 2015, course registration day ($\sim$5 months ago), we find 4,647 visitors and 2,209 visitors with schedules.
+Thus, we can estimate that roughly \textbf{22\%} of undergraduate and graduate students (9,470 total) at the College have added courses on Skedge \cite{undergrads}.
 

### 2016-04-13 conclusion.tex
chapter 4 done

@@
 
-On behalf of myself and the roughly \emph{one-thousand} students who regularly and voluntarily choose to use Skedge, I urge the University to consider adopting it as an official replacement to CDCS.
+On behalf of myself and the roughly two thousand students who regularly and voluntarily choose to use Skedge, I urge the University to consider adopting it as an official replacement to CDCS.
 
@@
 For the sake of the University's students, parents, faculty, and staff, I hope that my efforts with Skedge were not in vain and can lead to a substantial refresh of University tools that, most crucially, \textbf{involves a dialogue between the administration, students, and staff}.
-Thank you.
+Thank you for reading.
 

### 2016-04-13 introduction.tex
Split design files
little updates
Tech

@@
 
-Skedge is a website I developed in 2014 and have been maintaining and developing since.
+Skedge is a website on which I started rapid development on December 9th 2013, and have been developing it on and off since.
 
@@
 Students, parents, department coordinators, and faculty can all benefit from such tool improvements.
+
+user accounts are lightweight, no login, cookie (browser) based
 
@@
     }
-    \caption{Skedge, with the search query {\tt csc} and the user's current schedule on the right}
+    \caption[Skedge with the search query {\tt csc}]{Skedge with the search query {\tt csc} and the user's current schedule on the right}
     \label{fig:sk-index}

### 2016-04-13 tech.tex
New bibliography and citations
Cheeseball
Fix analytics intro
Tech

--- 2016-04-09 tech.tex
@@
 
-\section{Back-end}
+\section{System architecture}
 
-Skedge's infrastructure is built
+Skedge is a \emph{Ruby on Rails} web application running on a webserver called \emph{unicorn}, and routes ports 80 to it using the \emph{nginx} webserver as a reverse-proxy.
+Skedge uses \emph{PostgreSQL} as database, and \emph{React.js} on the front-end, a framework developed by Facebook for making responsive user interfaces.
+This stack is hosted on a paid (by me) virtual machine instance on DigitalOcean.
 
-\emph{nginx}, \emph{unicorn}, \emph{Ruby on Rails}, \emph{PostgreSQL}, \emph{React.js}, \emph{Ahoy}, and Google Analytics.
+\section{Data storage}
 
-\section{Front-end}
+\subsection{Courses}
 
-\section{Analytics}
+At the end of each day, Skedge runs an update script that updates its course data with any new course data that exists or has changed on CDCS, endearingly called ``scraping.'' There is minimal text processing to get the data to its final form as it is on Skedge, such as title formatting into proper capitalization.
+Highly sensitive data such as times, location, or virtually any other field remains completely unmodified.
+A substianal amount of logic goes into connecting courses to their subsections (workshops, labs, etc.), as these are not explicitly connected in CDCS and sometimes requires some basic natural language processing (finding CRN numbers, title similarity, etc.) Any ambiguities in this process are reported to me and they can easily be resolved (most arise from data entry errors in CDCS).
+
+\subsection{Users \& schedules}
+
+Users are purposefully light-weight on Skedge, because it aims for as little friction and overhead as possible before users can begin to search for courses and build a schedule.
+Thus there are no logins (besides now with Skedge Social, which is optional), and yet Skedge keeps users' schedules persistent.
+This is done using a single cookie on the client side that records the user ID that client is associated with, as well as a long hexadecimal string that serves as the ``user secret,'' or ``password'', which must match the one on the server for any changes to be made.
+This prevents users from manually changing cookie data to modify others' schedules.
+
+The disadvantage of the cookie-based system is that each new browser or device is disjoint by default.
+Previously, I made it possible to retrieve and set the user secret from the user interface, allowing users to manually synchronize all of their devices' browser cookies to the same user ID and thus manage the same schedules from different devices, but this was replaced by the same feature offered through Skedge Social, as part of promoting the platform.
+A different way to solve the problem of better persistence while not sacrificing the trade-off of lightweight accounts, and one that doesn't require external services like Facebook will be looked into.
+A likely candidate at the moment is using University of Rochester NetID for schedule synchronization and improved persistence.
+
+\subsection{Skedge Social}
+
+Out of respect for privacy and a preference to \emph{not} store sensitive data, Skedge does not keep or store any Facebook data, not even users' names or lists of friends, the two pieces of data Skedge requests access to during the integration process.
+Skedge only stores an anonymous ID (uniquely generated for Skedge by Facebook) per account.
+Only a user’s name and list of friends also using Skedge are accessible via this ID, and Skedge doesn’t even store this in its database---both are used exclusively by the client, however---separate requests are made for if any given users (friends of the user) are taking certain classes, etc., and for when displaying the schedule feed.
+
+Course \emph{likes} are also entirely locally stored and have nothing to do with Facbeook---Skedge merely uses the same concept and a similar icon for its own feature.
+Comments on courses are not currently supported, although Facebook-integrated comments would be possible.
+
+\section{Usage data collection}
+
+As the saying goes, ``if it's not monitored, it's not in production'' (how else would we know that it was?) \cite{monitored}.
+
+Skedge collects usage data that is anonymous with regards to who a specific user is (because that information is not even known), but that is not anonymous with respect to a trackable user ID.
+Keeping track of certain properties across users is very important for the data analysis in Sections 4.2 and 4.3.
+This tracking is done by the open-source library \emph{Ahoy}, which also has a JavaScript library to log click events unobtrusively and efficiently (e.g.
+send them after the page has been loaded) using cookies.
+Note that this has nothing to do with advertisement tracking that is done by marketing agencies across the web---this tracking is entirely local to Skedge and Skedge does not advertise in any way.
+
+\emph{Google Analytics} is also enabled on Skedge, which also involves cookie-based tracking, and is used for higher-level analytics such as session duration, number of sessions, etc.
+(a session is defined by a period of activity expiring after 30 minutes of inactivity.)

### 2016-04-13 design/search.tex
start intro

@@
 \begin{enumerate}
-  \item \emph{Year/term}, which is no longer required.
+  \item \emph{Year/term}, which is no longer a required field.
 It maps to the keywords {\tt \{spring,fall,summer,winter\}} and a four-digit year to specify term and year, and defaults to the current year and term.
+Skedge also gets around the fundamental limitation of CDCS of only being able to search one year-term at a time, allowing users to browse \emph{all} of a department's courses at once, for instance.
 

### 2016-04-13 introduction.tex
start intro

@@
 
-This paper will begin by
+In this paper, I am proud to finally introduce the world to \emph{Skedge}, a website I have been developing in various degrees since December 9th, 2013.
+Skedge began as a weekend side-project for me after spending too much time frustrated with existing tools while trying to find myself a second major to pursue.
+It wasn't until I briefly presented it at an on-campus RocHack meeting when I realized it had potential to help other students as well, but since, it has always existed in the shadows and in continual ``beta''---unofficial, unsupported and dodging recognition.
+Until two summers ago, I had never expected it to end up being the subject of my honors senior thesis, so this opportunity represents a rather special privilege for me and, selfishly, an outlet for recognition.
 
@@
 
-CDCS is the University's official tool for
+``Course Description / Course Schedule,'' (or \emph{CDCS} for short) is the University's official tool for browsing through the University's current course catalog.
 
@@
 
-Skedge is a website on which I started rapid development on December 9th 2013, and have been developing it on and off since.
+Skedge is a website on which I started rapid development on , and have been developing it on and off since.
 

### 2016-04-13 design/modernity.tex
little updates
done

@@
     }
-  \caption{CDCS in July 2, 2010, virtually unchanged from today, courtesy of \emph{Archive.org}}
+  \caption[CDCS in July 2, 2010]{CDCS in July 2, 2010, virtually unchanged from today, courtesy of \emph{Archive.org}}
   \label{fig:cdcs2010}
@@
 
-\item Skedge accommodates a schedule into its design, whereas Better CDCS has to work around an interface that wasn't designed with one.
-As a result, Better CDCS has lower usability, requiring the user to toggle between search results and their schedule.
-Skedge, conversely, gives users immediate visual feedback on how a course would fit into their schedule.
+\item Skedge was designed to have a schedule, whereas Better CDCS works within an interface that wasn't.
+CDCS suffers lower usability as a result, requiring toggle between results and schedules.
+Skedge, conversely, immediately projects a course into the user's schedule.
 
-\item Schedule data is centralized on Skedge's servers as opposed to locally in a browser cache, meaning that it can synchronize across a user's devices or in sessions on public computers, persists browser resets, and can be easily publicly shared to other users.
+\item Schedule data is centralized on Skedge's servers as opposed to locally in a browser, meaning that it can synchronize across a user's devices and can be easily publicly shared to others.
 
 \item Extensions like Better CDCS have limited browser support. Internet Explorer and mobile browsers are unsupported, for example.
+
+\item As separate systems trying to deliver one cohesive experience, Better CDCS and CDCS are dependent on each other. If the format of one changes, the other is likely to break.
 \end{enumerate}

### 2016-04-13 introduction.tex
done

--- 2016-04-13 introduction.tex
@@
 
-This paper will begin by
+\vspace{-20pt}
+
+I am proud to finally, officially introduce the world to \emph{Skedge}, a website I have been developing since December 2013.
+It began as a weekend project after I spent too much time trying to find interesting elective classes, and it wasn't until I briefly presented it at a RocHack ``Hacker Night'' when I realized the potential it had to help others as well.
+I never expected the subject of my senior thesis to be something so close to my heart, so I am grateful for such a fun and motivating opportunity to continue work on Skedge, and, selfishly, for the outlet of recognition.
 
@@
 
-CDCS is the University's official tool for
-
-\label{fig:cdcs-index}
-
-\subsection{``Better CDCS''}
-
-%%%
-
-\section{Overview of Skedge}
-
-Skedge is a website on which I started rapid development on December 9th 2013, and have been developing it on and off since.
-
-Bookmarks
-
-Students, parents, department coordinators, and faculty can all benefit from such tool improvements.
-
-user accounts are lightweight, no login, cookie (browser) based
-
-\ref{fig:sk-index}
-
-\begin{figure}[ht]
+\begin{figure}
     \centering
@@
 
-\begin{figure}[ht]
+``Course Description / Course Schedule,'' (or \emph{CDCS} for short) is the University's official tool for browsing the University's current course catalog \cite{cdcs}.
+
+The CDCS interface consists of a series of form fields for inputting search queries on the left, and search results on the right side (see Figure \ref{fig:cdcs-index}).
+Despite ``Course Schedule'' being in its name, CDCS does not offer scheduling functionality.
+Only course times can viewed, and it is up to the user to keep track of the courses they are considering for their next semester.
+
+\subsection{``Better CDCS''}
+
+Better CDCS \cite{better-cdcs} is a browser extension, available for Firefox, Safari, and Chrome, that offers a solution to the lack of scheduling mentioned above.
+It works by injecting ``Add Section'' and ``Bookmark Section'' buttons into the existing CDCS interface, and providing a tab on top of the screen that will toggle between the user's schedule and the search results (see Figure \ref{fig:cdcs-better}).
+
+%%%
+
+\section{Overview of Skedge}
+
+\begin{figure}[H]
     \centering
@@
 \end{figure}
+
+Skedge \cite{skedge} is, in short, CDCS combined with Better CDCS on a single platform (see Figure \ref{fig:sk-index}).
+Like CDCS, it is lightweight, service-oriented, and doesn't require a login, but unlike CDCS, it doesn't require a separate browser extension.
+
+It matches feature-for-feature with the two tools (search options, information displayed, scheduling and course bookmarks, etc.), and aims to improve upon the work of both of them on several fronts, which will be investigated in this paper in great detail.
+It is my contention that students, parents, faculty, staff, and administration can all benefit from such improvements.
+
+\vspace{20pt}
+
+\noindent
+This paper is organized in two parts, with a brief intermission in between.
+First, I will explain many of Skedge's design decisions in relation to what I call my ``grievances'' with CDCS in its current state.
+The second part will look at collected data, extrapolating how students use Skedge and using several novel metrics to measure its usability and effectiveness as a platform.

### 2016-04-13 introduction.tex
Slight mods

@@
 It began as a weekend project after I spent too much time trying to find interesting elective classes, and it wasn't until I briefly presented it at a RocHack ``Hacker Night'' when I realized the potential it had to help others as well.
-I never expected the subject of my senior thesis to be something so close to my heart, so I am grateful for such a fun and motivating opportunity to continue work on Skedge, and, selfishly, for the outlet of recognition.
+I never expected the subject of my senior thesis to be something so close to my heart, so I am grateful for such a fun and motivating opportunity to continue work on Skedge, and, selfishly, for an outlet of recognition.
 
@@
 
-``Course Description / Course Schedule,'' (or \emph{CDCS} for short) is the University's official tool for browsing the University's current course catalog \cite{cdcs}.
-
-The CDCS interface consists of a series of form fields for inputting search queries on the left, and search results on the right side (see Figure \ref{fig:cdcs-index}).
+``Course Description / Course Schedule,'' (or \emph{CDCS} for short) is the University's official tool for browsing the course catalog \cite{cdcs}.
+The CDCS interface consists of fields on the left side for inputting a search, and search results on the right side (see Figure \ref{fig:cdcs-index}).
 Despite ``Course Schedule'' being in its name, CDCS does not offer scheduling functionality.
-Only course times can viewed, and it is up to the user to keep track of the courses they are considering for their next semester.
+Course times can only be viewed, and it is up to the user to keep track of courses considered for the next semester.
 
@@
 
-Better CDCS \cite{better-cdcs} is a browser extension, available for Firefox, Safari, and Chrome, that offers a solution to the lack of scheduling mentioned above.
+\emph{Better CDCS} \cite{better-cdcs} is a browser extension, available for Firefox, Safari, and Chrome, that offers a solution to the lack of scheduling mentioned above.
 It works by injecting ``Add Section'' and ``Bookmark Section'' buttons into the existing CDCS interface, and providing a tab on top of the screen that will toggle between the user's schedule and the search results (see Figure \ref{fig:cdcs-better}).
@@
 
-Skedge \cite{skedge} is, in short, CDCS combined with Better CDCS on a single platform (see Figure \ref{fig:sk-index}).
-Like CDCS, it is lightweight, service-oriented, and doesn't require a login, but unlike CDCS, it doesn't require a separate browser extension.
+\emph{Skedge} \cite{skedge} is, in short, CDCS combined with Better CDCS on a single platform (see Figure \ref{fig:sk-index}).
+Like CDCS, it is lightweight, service-oriented, and doesn't require login, but unlike Better CDCS, it also doesn't require a separate browser extension.
 
-It matches feature-for-feature with the two tools (search options, information displayed, scheduling and course bookmarks, etc.), and aims to improve upon the work of both of them on several fronts, which will be investigated in this paper in great detail.
+It matches feature-for-feature with the two tools (search options, information displayed, scheduling and bookmarks, etc.), and aims to improve upon the work of both on several fronts, which will be investigated in this paper in great detail.
 It is my contention that students, parents, faculty, staff, and administration can all benefit from such improvements.
 
-\vspace{20pt}
+\vspace{30pt}
 

### 2016-04-14 analytics/npa.tex
submitted

@@
   The results are shown in Figure \ref{fig:navs-browse}.
-The difference between these results and the initial combined results of \ref{fig:navs-combined} are more substantial than last time, and may defy expectations at first glance.
+The difference between these results and the initial combined results of Figure \ref{fig:navs-combined} are more substantial than last time, and may defy expectations at first glance.
 While it may seem intuitive that browsed-for courses should have higher amounts of navigations, note that the graph itself is showing \emph{effectiveness} of such browsing.
@@
 This makes sense, as users may not be concerned with subsections if the goal of the browse is merely to mark interesting courses.
-This criterion saw more drastic drops in subsection amounts as well, as the number of navs increased: 5\% subsections for ${navs}=1$, and 4\% subsections for ${navs}=2$.
+This criterion saw more drastic drops in subsection amounts as well, as the number of navigationss increased: 5\% subsections for ${navs}=1$, and 4\% subsections for ${navs}=2$.
 
@@
   Given more time with Skedge Social being live and more forethought in tracking its usage, similar data analysis could be done on the subject of the effectiveness of Skedge in peer-guided course finding.
-Unfortunately, usage data to answer questions such as ``how many Facebook friends were listed as liking or taking a course when it was added?'', and cannot be retroactively computed due to Skedge's purposeful privacy limitations (Facebook friend data can only be retrieved client-side---no data is stored on the server that could authorize such a data request).
+Unfortunately, usage data to answer questions such as ``how many Facebook friends were listed as liking or taking a course when it was added?'' was not tracked, and cannot be retroactively computed due to Skedge's purposeful privacy limitations (Facebook friend data can only be retrieved client-side---no data is stored on the server that could authorize such a data request).
 Thus, part (c) of hypothesis \#2 cannot be verified.

### 2016-04-14 analytics/usage.tex
submitted

@@
 
-It should also be noted that before March, the majority of the userbase as I knew it consisted of seniors, as the majority of people that I knew personally and shared Skedge with belonged to my class.
+It should also be noted that before March, the majority of the userbase as I knew it consisted of seniors, as the majority of people that I knew personally and shared Skedge with belonged in my class year.
 The latest two major events reported in the graph above thus very impressively consist of \emph{very few seniors}, who have no courses left to schedule.
@@
 The usability of and user satisfaction with a tool is hard to measure when it holds a monopoly.
-Considering that Skedge is an \emph{optional alternative} to CDCS, such consistent usage data (especially with 75\% of sessions being from returning users) undeniably prove its measures of usability and user satisfaction relative to CDCS.
+Considering that Skedge is an \emph{optional alternative} to CDCS, such consistent usage data (especially with 75\% of sessions being from returning users) undeniably proves its measures of usability and user satisfaction relative to CDCS.
 
@@
 
-Figure \ref{fig:searchtypes} validates the hypothesis made in Section 2.3 regarding the search interface design, demonstrating that the \emph{department code} and \emph{course number} search modes dominate user searches, accounting for over 85\% of all searches on Skedge.
+Figure \ref{fig:searchtypes} validates the hypothesis made in Section 2.3 regarding the search interface design, demonstrating that the \emph{department code} and \emph{course number} search modes dominate user searches, accounting for over 85\% of all search types on Skedge.
 Bringing the nondistinct field for this in CDCS into prominence is thus a design choice that accomodates empirical behavior.
@@
   This already is evidence of an effective search mechanism, but of those empty searches, the vast majority contain typos or are for courses that simply don't exist.
-The other significant portion of searches include instructor names without the requisite ``taught by'' for SkedgeQL to pick it up as an instructor search, and this issue could easily be fixed.
+The other significant portion of searches include instructor names without the requisite ``{\tt taught by}'' for SkedgeQL to pick it up as an instructor search, and this issue could easily be fixed.
 
@@
 Making the case for usability superior to CDCS's in the space of mobile support is easy, given that CDCS does not have a mobile-responsive website.
-But to illustrate the importance of ensuring good user experience on mobile devices, consder that 11.54\% of all Skedge sessions are from mobile devices, which is very substantial.
+But to illustrate the importance of ensuring good user experience on mobile devices, note that 11.54\% of all Skedge sessions are from mobile devices, which is very substantial.
 On average, 2.74 pages viewed are viewed and 2 minutes are spent per session on mobile devices.
@@
   \caption[Course blocks in Skedge]{Hovering over a course block in Skedge's side-schedule shows at-a-glance information about it; clicking on it navigates the user to that course}
-  \label{fig:searchtypes}
+  \label{fig:courseblocks}
 

### 2016-04-14 design.tex
More splitup
submitted

@@
 From its very inception, Skedge's functionality and visual design were driven by the shortcomings of CDCS.
-Skedge is built \emph{bottom-up}, not \emph{top-down}---every aspect of the application was either made as a reaction to a particular grievance in CDCS or as the natural evolution of an existing feature.
+Skedge is built \emph{bottom-up}, not \emph{top-down}---every aspect of the application was either made as a reaction to a particular grievance in CDCS or as the natural evolution of an existing Skedge feature.
 Skedge is thus rooted in \emph{usability} derived from real need, not mere conjecture along the question ``what could students want?''.
-Its success with students, shown in Chapter 4, demonstrates that this usability extends beyond my own standard and can fulfill the various discovered use-cases of students in general.
+Its success with students, shown in Chapter 4, demonstrates that this usability extends beyond my own standards and preferences and can fulfill the various discovered use-cases of students in general.
 
@@
 
-\input{design-modernity}
+\input{design/modernity}
 \clearpage
@@
 
-\input{design-usability}
+\input{design/usability}
 \clearpage
@@
 
-\input{design-search}
+\input{design/search}
 \clearpage
@@
 
-\input{design-social}
+\input{design/social}

### 2016-04-14 design/modernity.tex
Slight mods
submitted

@@
 Yet, since its introduction in 2009, we have seen the rise of mobile devices into ubiquity, a boom in ``hacker culture'' and public APIs, and the capability for standalone web applications to be as sophisticated and dynamic as desktop-class applications without the aid of browser extensions.
-With this in mind, Skedge brings course scheduling to the modern era.
+To this end, Skedge brings course scheduling to the modern era.
 
@@
     }
-  \caption[CDCS in July 2, 2010]{CDCS in July 2, 2010, virtually unchanged from today, courtesy of \emph{Archive.org}}
+  \caption[CDCS in July 2, 2010]{CDCS in July 2, 2010, virtually unchanged from today, courtesy of \emph{Archive.org}\cite{archive}}
   \label{fig:cdcs2010}
@@
 
-Moreover, since no major mobile browser currently supports browser extensions (and if one did, the extensions themselves would most likely need to be re-architected), CDCS on a mobile device loses all scheduling functionality, unlike Skedge.
+Moreover, since no major mobile browser currently supports browser extensions (and if one did, the extensions themselves would most likely need to be re-architected), CDCS on a mobile device loses all scheduling functionality, unlike Skedge, which supports it.
 
@@
   \begin{tabular}{c c}
-    \begin{subfigure}[h]{4.5cm}
+    \begin{subfigure}[h]{4.8cm}
       \centering
@@
     \end{subfigure}
-    \begin{subfigure}[h]{4.4cm}
+    \begin{subfigure}[h]{4.7cm}
       \centering
@@
   \end{tabular}
-  \caption{CDCS and Skedge running on a mobile device}
+  \caption{CDCS and Skedge on a mobile browser}
 \end{figure}
@@
 
-With the increasing number of attendees at University of Rochester's hackathons, it is clear that the University's ``hacker culture'' is growing---more students are collaborating to build side-projects that integrate resources and servies often benefitting the student community.
-Open-source and open-information services greatly help to foster such innovation, and having public APIs is essential toward this end \cite{Milberry}\cite{hackers}.
+With the increasing number of attendees at University of Rochester's hackathons, it is clear that the University's ``hacker culture'' is growing---more students are collaborating to build side-projects that integrate resources and services often benefitting the student community.
+Open-source and open-information services greatly help to foster such innovation, and having public APIs is essential toward this end \cite{Milberry} \cite{hackers}.
 
-Skedge provides a public JSON API at the root URL \url{http://www.skedgeur.com/api/}, made at the request of a student that was interested in using its course data, and the API has already been used in projects by several other student groups.
+Skedge provides a public JSON API at the root URL \url{http://www.skedgeur.com/api/}, originally made at the request of a student that was interested in using course data, and the API has since been used in projects by several other student groups.
 The endpoints included as of now are \url{/api/courses?q=query} (Skedge's query language---described in detail in section 2.3---is supported here) and \url{/api/departments} for a list of department names and codes.
@@
 
-As explained in the introduction, Skedge offers users a course schedule right in the page, unlike CDCS which requires the Better CDCS browser extension for this functionality.
-Having a schedule native in the application has several advantages:
+Skedge offers users a course schedule right in the page, unlike CDCS which requires the Better CDCS browser extension. Having a native schedule has several advantages:
 
 \begin{enumerate}
-\item Besides some CDCS users possibly not even knowing about Better CDCS, not requiring a browser extension provides for a faster and more seamless user onboarding, especially when building schedules on public computers when extensions can't always be installed.
+\item Besides some CDCS users possibly not even knowing about Better CDCS, not requiring a browser extension provides for a faster and more seamless user onboarding, especially when building schedules on public computers where extensions can't always be installed.
 
-\item Skedge was designed to have a schedule, whereas Better CDCS works within an interface that wasn't.
-CDCS suffers lower usability as a result, requiring toggle between results and schedules.
-Skedge, conversely, immediately projects a course into the user's schedule.
+\item Skedge was designed with a native schedule in mind, whereas Better CDCS works within an interface that wasn't.
+CDCS suffers from lower usability as a result, requiring the user to toggle between results and the schedule.
+Skedge, conversely, provides immediate visual feedback on how courses fit∫ into the user's schedule when the cursor is hovered over them.
 

### 2016-04-14 design/search.tex
submitted

@@
 From this came Skedge's unified, natural language based search method, which I call the Skedge Query Language, or \emph{SkedgeQL}.
-Figure \ref{fig:sk-search} is a list of example possible searches shown to users as they type.
+Figure \ref{fig:sk-search} is a list of example possible searches shown to users as they type into the searchbox.
 
@@
   \end{tabular}
-  \caption{Two philosophies of search method}
+  \caption{Two philosophies of search}
 \end{figure}
@@
     The main advantage of Skedge's search system is in its reduction of 15 fields to a single one.
-I hypothesize that since the vast majority of searches fit the department code / course number pattern (this hypothesis later supported in Chapter 4), the time required to force the user to select a term and then find the correct box to enter this query is wasted.
+I hypothesize that since the vast majority of searches fit the department code / course number pattern (e.g.
+``{\tt csc 171}''---this hypothesis later supported in Chapter 4), the time required to force the user to select a term and then find the correct box to enter this query is wasted.
 
@@
     Another notable disadvantage of SkedgeQL is having to know it, and it having a possibly steep learning curve for some users.
-However, the argument will be made in Chapter 4 that simply by using the site over time, users will self-learn Skedge.
+However, the argument will be made in Chapter 4 that simply by using the site over time, users will self-learn SkedgeQL.
 One of the ways this is possible is thanks to Skedge's search system being \emph{multi-purposed}.
@@
 I have identified \emph{three} use-cases for course searching (the existence of and distinction between these cases will be demonstrated with collected usage data in Chapter 4).
-The three cases are \textbf{requirements}, \textbf{electives}, and \textbf{peer recommendations}.
+The types of courses user look for are \textbf{requirements}, \textbf{electives}, and \textbf{peer recommendations} (associated with direct, exploratory, and peer-guided search, respectively).
 SkedgeQL and other application features offer substantial improvements over CDCS for each of these cases.
@@
 
-  These are courses that are required for a student to complete their degree, and are typically searched for directly.
+  This is for are courses that are required for a student to complete their degree, and are typically searched for directly.
 The functionality required here is simple and is mostly satisfied by CDCS, but Skedge offers the following improvements to the process:
@@
 
-  Elective courses can either be courses that are required for a major or minor, but are chosen by the student, or they can be courses of particular interest to students who have the flexibility to take them.
+  Elective courses can either be courses that are required for a major or minor, but are chosen by the student, or they can be courses of particular interest to students who have the flexibility to take them,.
 For this use-case, Skedge offers search, filter, and sorting features that substantially aid users in browsing through courses and discovering ones they might want to take.

### 2016-04-14 design/social.tex
Tech
submitted

@@
       \includegraphics[width=6.1cm]{images/skedge/facebook}
-      \caption{Sharing a Skedge schedule on Facebook, linked to the live and dynamic schedule share page on Skedge} \label{fig:sk-fb}
+      \caption{Sharing a Skedge schedule on Facebook links to the live and dynamic schedule share page} \label{fig:sk-fb}
     \end{subfigure}
@@
 
-    As mentioned above, the schedules generated with Better CDCS cannot be shared publicly, and therefore students can only post a static image of their schedule at the time of posting.
-This excludes the possibility of edits to the schedule being reflected in the same post to their friends.
+    As mentioned previously, the schedules generated with Better CDCS cannot be shared publicly, and therefore students can only post a static, exported image of their schedule.
+This excludes the possibility of edits to the schedule being reflected to their peers in the same post.
 Moreover, referencing courses (i.e.
-looking up more information on a course a friend is taking) is not very usable under this model as it requires a manual search.
+looking up more information on a course a peer is taking) is not very usable under this model as it requires a manual search.
 
@@
   Immediately after logging in, the user will see the ``schedule feed,'' the list of the user's peers' current schedules (see Figure \ref{fig:sk-feed}).
-This takes from a familiar paradigm of the \emph{feed} in social networks, which is routinely scrolled for updates.[1] For each schedule, the user can visualize their own schedule side-by-side (using with the same ``overlay'' mechanic explained earlier), access the Facebook profile of the schedule's owner, and see the list of courses that user likes.
+This takes from a familiar paradigm of the \emph{feed} in social networks, which is routinely scrolled for updates.
+For each schedule, the user can visualize their own schedule side-by-side (using with the same ``overlay'' mechanic explained earlier), access the Facebook profile of the schedule's owner, and see the list of courses that user likes.
 

### 2016-04-14 design/usability.tex
little updates
submitted

@@
 Collapsing subsections within courses can result in massive improvements in filtering the data most relevant to the user.
-For instance, the search {\tt csc} for Fall 2016 on CDCS results in 147 ``section boxes'', while Skedge only shows 45 ``\emph{course} boxes'', with subsections collapsed within their respective course.
+For instance, the search ``{\tt csc}'' for Fall 2016 on CDCS results in 147 ``section boxes'', while Skedge only shows 45 ``\emph{course} boxes'', with subsections collapsed within their respective course.
 This triage reduces the data (noise, more correctly) displayed by \textbf{$\sim$70\%}, and is even higher for departments with more abundant labs and workshops, such as Physics (Skedge: 35 vs.
@@
 
-Skedge can reduce the amount of data to scroll through---and thus the time taken to do so---by six- or seven-fold (and possibly more, counting the attention users otherwise have to pay to distinguish course from subsections), so this design decision has a large usability payoff.
+Skedge can reduce the amount of data to scroll through---and thus the time taken to do so---by six- or seven-fold (and possibly more, counting the attention users otherwise have to pay to distinguish courses from subsections), so this design decision has a large usability payoff.
 
@@
   \includegraphics[width=14cm]{images/skedge/export}
-  \caption{Exporting to Google Calendar (note the other export options on the top right)} \label{fig:sk-export}
+  \caption[Skedge exporting options]{Exporting to Google Calendar (note the other export options on the top right)} \label{fig:sk-export}
 \end{figure}

### 2016-04-14 introduction.tex
submitted

@@
 
-\vspace{-20pt}
+\onehalfspacing
+{\small
+I am proud to finally and officially introduce \emph{Skedge}, a website I have been developing since December 2013.
+Skedge began as a weekend project after far too much time was spent trying to find interesting elective classes, and it wasn't until I briefly presented it at a RocHack ``Hacker Night'' when I realized the potential it had to help others as well.
+I never expected the subject of my senior thesis to be something so close to my heart, so I am grateful for such a fun and motivating opportunity to continue work on Skedge, and, selfishly, for an outlet of recognition.
+}
 
-I am proud to finally, officially introduce the world to \emph{Skedge}, a website I have been developing since December 2013.
-It began as a weekend project after I spent too much time trying to find interesting elective classes, and it wasn't until I briefly presented it at a RocHack ``Hacker Night'' when I realized the potential it had to help others as well.
-I never expected the subject of my senior thesis to be something so close to my heart, so I am grateful for such a fun and motivating opportunity to continue work on Skedge, and, selfishly, for an outlet of recognition.
+\doublespacing
 
-\section{Overview of CDCS}
+\section{CDCS}
 
@@
 
-\section{Overview of Skedge}
+\section{Skedge}
 
@@
 \noindent
-This paper is organized in two parts, with a brief intermission in between.
-First, I will explain many of Skedge's design decisions in relation to what I call my ``grievances'' with CDCS in its current state.
-The second part will look at collected data, extrapolating how students use Skedge and using several novel metrics to measure its usability and effectiveness as a platform.
+This paper is organized into two parts, with a brief intermission in between.
+First, I will explain many of Skedge's design decisions in response to what I call ``grievances'' with CDCS.
+We will break for a technical overview, and the second part will look at live usage data, extrapolating how Skedge is used by students and measuring its efficacy as a platform using some novel metrics.

### 2016-04-14 tech.tex
start intro
submitted

@@
-\chapter{Technical overview}
+\chapter{Very brief technical overview}
 
@@
 Skedge is a \emph{Ruby on Rails} web application running on a webserver called \emph{unicorn}, and routes ports 80 to it using the \emph{nginx} webserver as a reverse-proxy.
-Skedge uses \emph{PostgreSQL} as database, and \emph{React.js} on the front-end, a framework developed by Facebook for making responsive user interfaces.
+Skedge uses \emph{PostgreSQL} for the database, and \emph{React.js} on the front-end, a framework developed by Facebook for responsive user interfaces.
 This stack is hosted on a paid (by me) virtual machine instance on DigitalOcean.
@@
 
-At the end of each day, Skedge runs an update script that updates its course data with any new course data that exists or has changed on CDCS, endearingly called ``scraping.'' There is minimal text processing to get the data to its final form as it is on Skedge, such as title formatting into proper capitalization.
-Highly sensitive data such as times, location, or virtually any other field remains completely unmodified.
-A substianal amount of logic goes into connecting courses to their subsections (workshops, labs, etc.), as these are not explicitly connected in CDCS and sometimes requires some basic natural language processing (finding CRN numbers, title similarity, etc.) Any ambiguities in this process are reported to me and they can easily be resolved (most arise from data entry errors in CDCS).
+At the end of each day, Skedge runs an update script that updates its course data with any new course data that exists or has changed on CDCS, endearingly called ``scraping.'' It accesses the ordinary CDCS search endpoint, but in XML format.
+There is minimal text processing to get the data to its final form as it is on Skedge, such as title formatting into proper capitalization.
+Highly important data such as times, location, or virtually any other field remains unmodified.
+
+A substantial amount works to connect courses to their subsections (workshops, labs, etc.), as these are not explicitly connected in CDCS and sometimes requires basic natural language processing (finding CRN references, title similarity, etc.) Any ambiguities in this process are reported to me and can easily be resolved (most arise from data entry errors in CDCS).
 
@@
 
-Users are purposefully light-weight on Skedge, because it aims for as little friction and overhead as possible before users can begin to search for courses and build a schedule.
+User accounts are purposefully light-weight on Skedge, because it aims for as little friction and overhead as possible before users can begin to search for courses and build a schedule.
 Thus there are no logins (besides now with Skedge Social, which is optional), and yet Skedge keeps users' schedules persistent.
@@
 Previously, I made it possible to retrieve and set the user secret from the user interface, allowing users to manually synchronize all of their devices' browser cookies to the same user ID and thus manage the same schedules from different devices, but this was replaced by the same feature offered through Skedge Social, as part of promoting the platform.
-A different way to solve the problem of better persistence while not sacrificing the trade-off of lightweight accounts, and one that doesn't require external services like Facebook will be looked into.
+A different way to solve the problem of better persistence while not sacrificing the trade-off of lightweight accounts (and one that doesn't require external services like Facebook) is being considered.
 A likely candidate at the moment is using University of Rochester NetID for schedule synchronization and improved persistence.
@@
 Skedge only stores an anonymous ID (uniquely generated for Skedge by Facebook) per account.
-Only a user’s name and list of friends also using Skedge are accessible via this ID, and Skedge doesn’t even store this in its database---both are used exclusively by the client, however---separate requests are made for if any given users (friends of the user) are taking certain classes, etc., and for when displaying the schedule feed.
+Only a user’s name and list of friends also using Skedge are accessible via this ID, and Skedge doesn't store this in its database---both are retrieved exclusively by the client, and making separate requests that ask whether any given users (i.e.
+peers) are taking certain classes, etc., and for when displaying the schedule feed.
 
@@
 
-As the saying goes, ``if it's not monitored, it's not in production'' (how else would we know that it was?) \cite{monitored}.
+As the saying goes, ``if it's not monitored, it's not in production'' (how else would we know that it's truly live?) \cite{monitored}.
 
-Skedge collects usage data that is anonymous with regards to who a specific user is (because that information is not even known), but that is not anonymous with respect to a trackable user ID.
+Skedge collects usage data that is anonymous with regards to who a specific user is (because that information is not even known or stored), but that is not anonymous with respect to a trackable user ID.
 Keeping track of certain properties across users is very important for the data analysis in Sections 4.2 and 4.3.

### 2016-04-14 analytics/dsl.tex
Tech
Small fixes

@@
 
-The \textbf{complexity} of a search query is the number of fields that aren't course number or department that it spans (the accepted fields are shown in Table \ref{fig:sc-fields}).
+The \textbf{complexity} of a search query is the number of fields that aren't \emph{course number} or \emph{department} that it spans. The accepted fields are shown in the following table:
 
@@
   \vspace{10pt}
-  \caption{``Complex'' search fields (omits only number and department)}
+  \caption{``Complex'' search fields (omits only \emph{number} and \emph{department})}
   \label{fig:sc-fields}
@@
 
-It is possible that user have neither, both, or only a \emph{first increase} metric.
+It is possible that a user has neither metric, both metrics, or only a \emph{first increase} metric.
 
@@
     & 0
-    & 
+    & \hspace{55pt} ---
     \\
@@
 
-  Other possibilities include linking the credits field of each course to a credits search, the CRN field of a section to a CRN search, the ``{\tt W}'' part of the course code to an upper-level writing search, having a crosslisted course link search the crosslisting of the two departments (instead of directly link to the crosslisted course), a ``feeling lucky?'' button for a search with ``{\tt random}''.}
+  Other possibilities include linking the credits field of each course to a credits search, the CRN field of a section to a CRN search, the ``{\tt W}'' part of the course code to an upper-level writing search, having a crosslisted course link search the crosslisting of the two departments (instead of directly link to the crosslisted course), or a ``feeling lucky?'' button for a search with ``{\tt random}''.}
 \end{enumerate}
@@
 This is purposeful---title searches only represent 8\% of all search field uses, as we saw in Figure \ref{fig:searchtypes} of the section before last.
-Moreover, a user's understanding and/or intuition that the Skedge search field is free-form and can be used for different search fields simultaneously is already a significant conceptualization of SkedgeQL, albeit a relatively basic field.
+Moreover, a user's understanding and/or intuition that the Skedge search field is free-form and can be used for different search fields simultaneously is already a significant conceptualization of SkedgeQL, albeit for a relatively basic search type.
 
@@
 
-  \caption{Users with a schedule and more than one search split by their number of increases in search complexity}
+  \caption[Users split by their number of increases in search complexity]{Users with a schedule and more than one search split by their number of increases in search complexity}
   \label{fig:sc-pie}

### 2016-04-14 analytics/npa.tex
Small fixes

@@
 This makes sense, as users may not be concerned with subsections if the goal of the browse is merely to mark interesting courses.
-This criterion saw more drastic drops in subsection amounts as well, as the number of navigationss increased: 5\% subsections for ${navs}=1$, and 4\% subsections for ${navs}=2$.
+This criterion saw more drastic drops in subsection amounts as well, as the number of navigations increased: 5\% subsections for ${navs}=1$, and 4\% subsections for ${navs}=2$.
 

### 2016-04-14 design/search.tex
Small fixes

@@
 
-  Elective courses can either be courses that are required for a major or minor, but are chosen by the student, or they can be courses of particular interest to students who have the flexibility to take them,.
+  Elective courses can either be courses that are required for a major or minor, but are chosen by the student, or they can be courses of particular interest to students who have the flexibility to take them.
 For this use-case, Skedge offers search, filter, and sorting features that substantially aid users in browsing through courses and discovering ones they might want to take.
@@
 
-    \item SkedgeQL includes the ``new'' keyword, which qualifies an accompanying search by only displaying the courses that were not offered the previous year for that term.
-For instance, searching ``new csc'' today (Fall 2016) would show Fall Computer Science courses that have been added to the catalog or changed since Fall 2015.
-    \item SkedgeQL could include an ``autofit'' keyword, which would take into account the user's current schedule, and only show sections that would not conflict with any already scheduled times.\footnote{This feature is under development and is not currently live.}
-    \item SkedgeQL includes the ``random'' keyword, which displays a single random course matching the accompanying query.
-For instance, ``random csc'' would show a random course in Computer Science for the current term.
-Used in conjunction with ``autofit'', it can a powerful way to find interesting new classes that are available for users to try.
+    \item SkedgeQL includes the ``{\tt new}'' keyword, which qualifies an accompanying search by only displaying the courses that were not offered the previous year for that term.
+For instance, searching ``{\tt new csc}'' today (Fall 2016) would show Fall Computer Science courses that have been added to the catalog or changed since Fall 2015.
+    \item SkedgeQL could include an ``{\tt autofit}'' keyword, which would take into account the user's current schedule, and only show sections that would not conflict with any already scheduled times.\footnote{This feature is under development and is not currently live.}
+    \item SkedgeQL includes the ``{\tt random}'' keyword, which displays a single random course matching the accompanying query.
+For instance, ``{\tt random csc}'' would show a random course in Computer Science for the current term.
+Used in conjunction with ``{\tt autofit}'', it can a powerful way to find interesting new classes that are available for users to try.
     \item SkedgeQL includes sorting by class size---useful for students trying to find smaller, discussion based classes---or by class starting time---for students with preferences of morning or afternoon/evening classes.

### 2016-04-14 tech.tex
Small fixes

@@
 
-Course \emph{likes} are also entirely locally stored and have nothing to do with Facbeook---Skedge merely uses the same concept and a similar icon for its own feature.
+Course \emph{likes} are also entirely locally stored and have nothing to do with Facebook---Skedge merely uses the same concept and a similar icon for its own feature.
 Comments on courses are not currently supported, although Facebook-integrated comments would be possible.
@@
 
-As the saying goes, ``if it's not monitored, it's not in production'' (how else would we know that it's truly live?) \cite{monitored}.
+As the saying goes, ``if it's not monitored, it's not in production'' (how else would we know if it's really live?) \cite{monitored}.
 

### 2016-04-14 analytics/usage.tex
small typo

@@
   \hline
-  141 users have linked Skedge to Facebook ($\sim$3.5 per day) \\
+  141 users on the platform ($\sim$3.5 per day) \\
   10,242 page views (256 page views/day)  \\
@@
   a 5\% bounce rate (users interact with the page before leaving) \\
-  67\% of visits being were returning visitors \\
+  67\% of visits being returning visitors \\
   284 overlays onto friends’ schedules (2 per user) \\

### 2016-04-14 paper.tex
Tech
done
Slight mods
Update abstract

@@
 
-In this paper I present Skedge, a web application for students to comfortably and effectively engage with the University's course catalog.
-Skedge matches and surpasses the capabilities of the existing University tool for this purpose, ``Course Description / Course Schedule'' (CDCS) and presents its information in a more visually appealing way.
-As a result, Skedge boasts strong user-retention rates, long session durations, and high student adoption despite having virtually no advertisement.
-Through collected usage data, I demonstrate that \textbf{a)} Skedge's differences from and additions to CDCS are usable and have real-world need, \textbf{b)} the three major use-cases associated with course scheduling---direct, exploratory, and peer-guided search---are effectively accommodated by Skedge, and \textbf{c)} Skedge's novel search mechanism is user-friendly and self-teaches to users over time.
+In this paper I present Skedge, a web application made for students to comfortably and effectively engage with the University’s course catalog.
+Skedge matches and surpasses the capabilities of the existing University tool for this purpose, ``Course Description / Course Schedule'' (CDCS), and boasts strong student adoption and retention rates despite having virtually no advertisement.
+Through collected usage data, I demonstrate that \textbf{a)} Skedge's differences from and additions to CDCS are usable and meet real-world needs, \textbf{b)} the three major use-cases associated with course scheduling---direct search, exploratory search, and peer-guided search---are effectively accommodated by Skedge, and \textbf{c)} Skedge's novel search mechanism is user-friendly and self-teaches to users over time.
 
@@
 % Bibtex commands go here
+
+\begin{spacing}{1.1}
+
 \bibliographystyle{ieeetr}
@@
 
+\end{spacing}
+
 \addcontentsline{toc}{chapter}{Bibliography}
 
-\clearpage
-\section*{Appendix}
-
 \end{document}

### 2016-04-27 analytics/npa.tex
Minor edits

@@
 
-\subsection{Direct search (requirement criteria)}
+\subsection{Direct search (required courses)}
 
@@
 
-  The difference between these results and those in \ref{fig:navs-combined} is suble but logical.
-Users searching for a specific course should have an increased proportion of 1-navigation adds, as the graph shows, because the use-case is \emph{search for course}, \emph{add course}, perhaps occasionally taking an extra try to account for typos, not knowing the title or course code exactly, etc.
+  The difference between these results and those in Figure \ref{fig:navs-combined} is subtle but logical.
+Users searching for a specific course should have an increased proportion of 1-navigation adds, as the graph shows, because the use-case is ``\emph{search for course}'' followed by ``\emph{add course}'', perhaps occasionally taking an extra try to account for typos, etc.
 
   But given this reasoning, why would 0-navigation adds still be so common in direct searches?
-Further analysis on only the sections that were added from direct search and with 0 navigations shows that 64\% of those sections are in fact \emph{subsections}.
-As happened in the illustrative example at the beginning of this section, this shows that a very common use-case is to add labs, workshops, etc.
-of a course directly after adding the main section.
+Further analysis on sections added from 0 navigations (from direct search) shows that 64\% of those sections were in fact \emph{subsections}.
+As was illustrated in the example at the beginning of this section, this demonstrates the very common use-case of adding labs, workshops, etc.
+of a course directly after adding the main section when directly searching.
 
-  This hypothesis is further confirmed by doing the same analysis for ${navs}=1$ and ${navs}=2$, when the inverse behavior would be expected (it makes sense for subsections which were added seperately from their main sections to be more rate).
-Only 11\% of sections that were added from a direct search after 1 navigation are subsections, and that number drops to 8\% for 2 navigations.
+  This hypothesis is further confirmed by doing the same analysis for ${navs}=1$ and ${navs}=2$, when the inverse behavior would be expected (it makes sense for subsections which were added seperately from their main sections to be more rare).
+Only 11\% of sections that were added from a direct search after 1 navigation were subsections, and that number drops to 8\% for 2 navigations.
 
-\subsection{Browse (elective criteria)}
+\subsection{Browse (elective courses)}
 
@@
 
-\subsection{Peer-guided (recommendation criteria)}
+\subsection{Peer-guided (recommended courses)}
 

### 2016-04-27 design/search.tex
Minor edits

@@
 I have identified \emph{three} use-cases for course searching (the existence of and distinction between these cases will be demonstrated with collected usage data in Chapter 4).
-The types of courses user look for are \textbf{requirements}, \textbf{electives}, and \textbf{peer recommendations} (associated with direct, exploratory, and peer-guided search, respectively).
+The types of courses users look for are \textbf{requirements}, \textbf{electives}, and \textbf{peer recommendations} (associated with direct, exploratory, and peer-guided search, respectively).
 SkedgeQL and other application features offer substantial improvements over CDCS for each of these cases.

### 2016-04-27 design/social.tex
Minor edits

@@
       \caption[Skedge Social landing page for a user who hasn't yet joined it]{Skedge Social landing page (\url{http://skedgeur.com/social}) for a user who hasn't yet joined it.
-Skedge Social appears in the context of the ordinary Skedge interface, and is accessible with the globe icon to the right of the searchbar.} \label{fig:sk-social-login}
+Skedge Social appears in the context of the ordinary Skedge interface, and is accessible from the globe icon to the right of the searchbar.} \label{fig:sk-social-login}
     \end{figure}
@@
 
-  This has the advantage of organizing and persisting peers' schedules, solving the first three issues of the ``Facebook post'' model listed above.
+  This has the advantage of organizing and persisting peers' schedules, solving the first two issues of the ``Facebook post'' model listed above.
 
@@
   Skedge Social also integrates into course search results, showing the user any other peers on the platform who are taking, have taken, or \emph{liked} courses in the results list (see Figure \ref{fig:sk-social-course}).
-This has several advantages, which solve the last ``Facebook post'' model issue listed above:
+This has several advantages, which solve the last two ``Facebook post model'' issues listed above:
 
